---
title: "Dave Andison's fire data preliminary analyses"
author: "Ceres Barros"
date: "March 28, 2018"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
## using - as of January 14th, 2019
# loading reproducible     0.2.5.9000
# loading quickPlot        0.1.6
# loading SpaDES.core      0.2.3.9003
# loading SpaDES.tools     0.3.0.9000
# loading SpaDES.addins    0.1.1
# devtools::install_github("PredictiveEcology/quickPlot@development")
# devtools::install_github("PredictiveEcology/reproducible@development")
# devtools::install_github("PredictiveEcology/SpaDES.core@development")
# devtools::install_github("PredictiveEcology/SpaDES.tools@development")


knitr::opts_knit$set(root.dir = "~/GitHub/LandscapesInMotion/")
library(ggplot2)
library(factoextra, lib.loc = "/home/cbarros/R/x86_64-pc-linux-gnu-library/3.5")
library(reshape2)
library(permute, lib.loc = "/home/cbarros/R/x86_64-pc-linux-gnu-library/3.5")
library(vegan, lib.loc = "/home/cbarros/R/x86_64-pc-linux-gnu-library/3.5")
library(sf)
library(data.table)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(quickPlot)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE) 
source("analyses/R_tools/summarizeABSK_AllData.R")
source("R/R_tools/prepCorrTable.R")
```


```{r dataprep, include = FALSE, echo = FALSE}
## source dataprep:
doCache <- TRUE ## control all caching
bindAllFires <- FALSE ##control whether all fire data binding should be repeated
source("analyses/DAfires_expAnalyses_dataPrep.R", echo = FALSE)

## exclude event perimeter patch type
memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) !any(!is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
rm(NAcols)
```

## NOTES - meeting with Dave Dec 6th
Carefull with no. of variables and sample size
Weather variables: how to look at it? Weather at the begining, middle, end? Only averages like Colin did? Averages, variation?

## COMMUNITY COMPOSITION/STRUCTURE PCAs
PCA of dominant species percentage cover and height across all fire event patch types. 
Blue arrow is severity in a continuous scale, fitted to the ordination *a posteriori*, using `env.fit`.
Severity is higher in "plots" dominated by White spruce (*Picea glauca*), Engelman's spruce (*P. engelmannii*), Aspen (*Populus tremuloides*) and where trees are taller.
It seems that canopy and understory levels usually highly correlated within species.

```{r PCA, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR|LAYER$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1+LAYER, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])

## PCA COLOURED BY FIRE_NAME
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by fire name")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$FIRE_NAME))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```

### PCA with canopy and understory levels collapsed for tree species (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but clearer.

```{r PCA2, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

# ## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```

### PCA with canopy and understory levels collapsed for tree species, but keeping dominant and subdominant spp separate (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but we now see some vegetation associations. Unlike before, crown closure now appears more correlated with severity. It seems higher severity levels coincide with forests where Aspen is the dominant species and White spruce is the subdominant. 

```{r PCA3, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2|3)$|SPEC(1|2|3)_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

spec3 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC3, value.var = "SPEC3_PER", fun.aggregate = sum, na.rm = TRUE)
setnames(spec3, sppNames, paste0(sppNames, "_3"), skip_absent = TRUE)

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
commCompStruc <- merge(commCompStruc, spec3, by = cols)
rm(spec1, spec2, spec3); amc::.gc()

commCompStruc$`NA` <- NULL

## check if any spp has >100 cover
# cols <- grep("_[[:digit:]]$", names(commCompStruc))
# summary(commCompStruc[, ..cols])

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```


### Remove third subdominant species, add age
Correlations between third most dominant species and axes are quite low
Results changed quite a bit, with Lodgepole pine and crown closure appearing both more related to severity, than before (and Aspen and White spruce losing importance).

```{r PCA4, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|ORIGIN_|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|ORIGIN|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

commCompStruc$`NA` <- NULL

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

### Remove origin - correlations with axes are quite low

```{r PCA5, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

commCompStruc$`NA` <- NULL

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

### Pure stands only 
Collapsing canopy and understory

```{r PCA6, echo = FALSE, fig.height=8, fig.show = "hold",  warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[SPEC1_PER == 100, ..cols]

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch|SPEC1", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## SPEC1 needs to be a factor for dudi
commCompStruc <- commCompStruc[, SPEC1 := as.factor(SPEC1)]

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompDudi <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB],
                              scannf = FALSE, nf = 4)
env.fit <- envfit(vegCompDudi, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompDudi, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

## COMMUNITY COMPOSITION/STRUCTURE PCAs - FOOTHILLS AND BOREAL SEPARATED

### PCA with canopy and understory levels collapsed for tree species, but keeping dominant and subdominant spp separate (summed and rescaled to 0-1 - maybe not the best way)
Only 1st and 2nd most dominant species included (same as 5th PCA)

Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r FootVsBoreal_PCA, echo = FALSE, fig.height=8, fig.show = "hold", warning=FALSE}
cols <- grep("NR1|SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("NR1|SEV|Patch|FIRE|pixID|ORIGIN|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2, sppNames); amc::.gc()

commCompStruc$`NA` <- NULL

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|NR1", colsA, invert = TRUE, value = TRUE)

## FOOTHILLS + ROCKY MOUNTAIN
vegCompPCA <- unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")

## BOREAL FOREST + CANADIAN SHIELD
vegCompPCA <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SEV_CONT %>%
  envfit(vegCompPCA, env = .)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")

## colour by SMR in Boreal
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
colsB <- grep("SEV_CONT|FIRE|NR1|SMR", colsA, invert = TRUE, value = TRUE)

vegCompPCA <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SEV_CONT %>%
  envfit(vegCompPCA, env = .)

PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by SMR\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SMR))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```


## SIMPLIFY ANALYSIS/DATA 
- Collapse climate data using summary statistics
- Convert species data to long format

```{r summaryData, echo = FALSE, }
## clean-up and free memory
rm(vegCompPCA, env.fit, commCompStruc, PCAplot); amc::.gc()

## summarise weather data
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             saveDir = "analyses/fireDataSummary",
                             overwrite = FALSE, ## ovewrite saved weather summary table?
                             useCache = doCache,
                             userTags = "summarizeABSK_AllData",
                             cacheRepo = "analyses/cache")
amc::.gc()
```

## COMMUNITY COMPOSITION/STRUCTURE PCAs - simplified data
Including elevation, slope, presence/absence of understory in PCA then fitting environmental vectors (mean and cv of temp, rain, rh, ffmc, fwi) and severity

```{r summarizedData_PCA, echo = FALSE, fig.height=8, fig.show = "hold", warning=FALSE}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR1|NR2", 
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix

## exclude also P. banksiana, because it has no data
cols2 <- grep("SEV_CONT|bank", cols, value = TRUE, invert = TRUE)
summDataPCA <- ade4::dudi.mix(as.data.frame(summaryABSK_AllData2[, ..cols2])
                              , nf = 3
                              , scannf = FALSE)

cols2 <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2[, ..cols2]))

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[["CROWN_CLOSURE_LOWER"]]))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = cols2)
```

### Now separating Foothills + Rocky Mountain from Boreal Forest + Canadian Shield
Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r FootVsBoreal_summarizedData_PCA, echo = FALSE, fig.height=8, fig.show = "hold", warning=FALSE}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR2", names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix

## FOOTHILLS + ROCKY MOUNTAIN
## exclude also P. banksiana, because it has no data
colsA <- grep("SEV_CONT|bank|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)


PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)


## BOREAL FOREST + CANADIAN SHIELD
## exclude also P. banksiana, P. engelmann and A. lasiocarpa because they have no data
colsA <- grep("SEV_CONT|bank|enge|lasi|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|enge|lasi|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)


PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
```

## FIRE WEATHER PCA - simplified data
Average and covariance of temp, rain, rh, ffmc, fwi
Also including elevation, slope, and then fitting everity

```{r summarizedData_PCA2, echo = FALSE}
rm(summaryABSK_AllData2); amc::.gc()

## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR1|NR2", 
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix

## exclude also P. banksiana, because it has no data
cols2 <- grep("mean|cv|Elev|Slope", cols, value = TRUE)
summDataPCA <- rda(as.data.frame(summaryABSK_AllData2[, ..cols2]),
                   scale = TRUE)
env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2$SEV_CONT))

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by mean FFMC")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2$meanFFMC))
text(PCAplot, "species", labels = rep("", length(cols2)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```

## CORRELATIONS
Pearson and spearman correlations between all covariates and each covariate and severity

### Pearson correlations Vegetation and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and ORIGIN_LOWER removed as they are redundant

Some "decent" correlation values (>0.15) between
* Picea glauca and -height, +age (origin)
* Pinus contorta and -Picea mariana, -Larix laricinea, -Populus tremuloides, -Pinus banksiana (wet vs dry)
* Pinus contorta and +closure, +height, -origin, +elevation
* Picea mariana and -Populus tremuloides, -Pinus banskiana,
* Picea mariana and -closure, -height, -origin, -slope
* Larix laricinea and -Populus tremuloides, -Pinus banskiana
* Larix laricinea and -closure, -height, -slope, +elevation
* Populus tremuloides and +closure, +height, -elevation
* Pinus banksiana  and +closure, +height, +origin, +elevation, +slope, -elevation
* Picea engelmannii and Abies lasiocarpa
* Picea engelmannii and -origin, +slope, +elevation 
* Abies lasiocarpa and +slope, +elevation

* Closure and +height, +origin
* Height and -origin
* Slope and +elevation (but not lots), ++SurfAspectRatio

```{r vegTopoCorrPearson, echo = FALSE, cache = TRUE}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
print(xtable::xtable(tabl), type = "html")
```

### Pearson correlations between fire weather and topo covariates only
SurfAspectRatio adn DegSlope removed due to high correlation with PctSlope

Some high correlation values (>0.5) between
* startJulDay and +meanRH, -cvRH, -maxFFMC, +meanDC, +minDC, +maxDC
* meanTemp and +minTemp, -cvTemp, -cvRH, +rangeDC, +meanBUI, +minBUI, +minFWI, -cvFWI
* minTemp and -cvtemp, -cvRH, +randeDC, +meanBUI
* maxTemp and cvTemp, rangeRH, -minRH, -minDMC, +rangeDMC, +meanISI, +maxISI, +cvISI, +rangeISI, +rangeBUI, +maxFWI
* cvTemp and rangeRH, -minDMC, -rangeDC, +maxISI, +cvISI
* rangeRH and -meanRH, maxRH, -minDMC, +cvDMC, +rangeDMC, +maxISI, +cvISI, +rangeISI, -minBUI, +rangeBUI, -minFWI
* meanRH and +minRH, +maxRH, -minDMC, +cvDMC, +rangeDMC, +maxISI, +cvISI, +rangeISI, -minBUI, +rangeBUI, -minFWI
* minRH and +meanRain, +maxRain, +rangeRain, -meanFFMC, -minFFMC, -maxFFMC, cvFFMC, -maxDMC, -rangeDMC, -meanISI, -maxISI, -cvISI, -rangeISI, -maxBUI, -rangeBUI, -meanFWI, -maxFWI, -rangeFWI
* maxRH and +meanRain, -meanFFMC, -minFFMC, -maxFFMC, cvFFMC, rangeFFMC, -meanDMC, -minDMC, -maxDMC, cvDMC, -meanISI, -minISI, -meanBUI, -minBUI, -maxBUI, +cvBUI, -meanFWI, -minFWi, -maxFWI, +cvFWI
* cvRH and +maxFFMC, -maxDC
* meanWS and +maxWS, +rangeWS
* maxWS and +cvWS, +rangeWS
* cvWS and +rangeWS
* rangeWS and +elevation
* meanRain and +maxRain, +rangeRain, -meanFFMC, -minFFMC, -maxFFMC, +cvFFMC, +rangeFFMC, -meanDMC, -maxDMC, +cvDMC, +cvDC, -meanISI, -maxISI, -rangeISI, -meanBUI, -minBUI, -maxBUI, +cvBUI, -meanFWI, -maxFWI, +cvFWI, -rangeFWI
* maxRain and +rangeRain, -meanFFMC, -minFFMC, -maxFFMC, +cvFFMC, +rangeFFMC, -maxDMC, +cvDMC, +cvDC, -meanISI, -meanBUI, -minBUI, -maxBUI, +cvBUI, -meanFWI, -maxFWI, +cvFWI,, -rangeFWI
* cvRain and +meanFFMC, +meanISI, +meanFWI
* rangeRain and -meanFFMC, -minFFMC, -maxFFMC, +cvFFMC, +rangeFFMC, -maxDMC, +cvDMC, +cvDC, -meanISI, -meanBUI, -minBUI, -maxBUI, +cvBUI, -meanFWI, -maxFWI, +cvFWi, -rangeFWI
* meanFFMC and +minFFMC, +maxFFMC, -cvFFMC, -rangeFFMC, +meanDMC, +maxDMC, -cvDMC, -cvDC, +meanISI, +meanBUI, +minBUI, +maxBUI, -cvBUI, +meanFWI, +maxFWI, -cvFWI, +rangeFWI
* minFFMC and -maxFFMC, -cvFFMC, +rangeFFMC, +meanDMC, +maxDMC, -cvDMC, +meanISI, +minISI, +meanBUI, +minBUI, -cvBUI, +meanFWI, +minFWI, +maxFWI, -cvFWI
* maxFFMC and +meanDMC, +maxDMC, +meanFWI, +maxFWI
* cvFFMC and +rangeFFMC, -meanDMC, -maxDMC, +cvDMC, +cvDC, -meanISI, -minISI, -meanBUI,-minBUI, -maxBUI, +cvBUI, -meanFWI, -maxFWI, +cvFWI, -rangeFWI
* meanDMC and +minDMC, +maxDMC, -cvDMC, +meanISI, +minISI, +meanBUI, +minBUI, +maxBUI, -cvBUI, +meanFWI, +minFWI, -cvFWI
* minDMC and -cvDMC, -rangeDMC, +minISI, -maxISI, -cvISI, -rangeISI, +minBUI, -cvBUI, -rangeBUI, +meanFWI, +minFWI, -cvFWI
* maxDMC and -cvDMC, +meanISI, +minISI, +meanBUI, +minBUI, +maxBUI, -cvBUI, +meanFWI, +minFWI, +maxFWI, -cvFMI
* cvDMC and -minISI, +cvISI, -meanBUI, -minBUI, +cvBUI, -meanFWI, -minFWI, +cvFWI
* rangeDMC and +meanISI, +maxISI, +cvISI, +rangeISI, +maxBUI, +rangeBUI
* meanDC and +minDC, +maxDC, -cvDC
* minDC and +maxDC, -cvDC
* maxDC and -cvDC
* cvDC and +rangeDC, -maxBUI, +cvBUI
* meanISI and +maxISI, +rangeISI, +meanBUI, +maxBUI, +rangeBUI, +meanFWI, +maxFWI, -cvFWI, +rangeFWI
* minISI and +meanBUI, +minBUI, -cvBUI, +meanFWI, +minFWI, -cvFWI
* maxISI and +cvISI, +rangeISI, +meanFWI
* cvISI and +rangeISI, +rangeBUI
* rangeISI and +rangeBUI
* meanBUI and +minBUI, +maxBUI, -cvBUI, +meanFWI, +minFWI, -cvFWI
* minBUI and +maxBUI, -cvBUI, +meanFWI, +minFWI, -cvFWI
* maxBUI and -cvBUI, +rangeBUI, +meanFWI, +maxFWI, -cvFWI
* cvBUI and -meanFWI, -minFWI, +cvFWI
* meanFWI and +minFWI, +maxFWI, -cvFWI, +rangeFWI
* minFWI and -cvFWI
* maxFWI and +rangeFWI

```{r weatherTopoCorrPearson, echo = FALSE, cache = TRUE}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|DegSlope|SurfAspectRatio|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|HEIGHT|CROWN|ORIGIN|pixID|Czone|^NR1|^NR2|Lari|Pice|Pinu|Abie|Popu|Betu",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
print(xtable::xtable(tabl), type = "html")
```


### Spearman correlations Vegetation and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and ORIGIN_LOWER removed as they are redundant

```{r vegTopoCorrSpearman, echo = FALSE, cache = TRUE}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
print(xtable::xtable(tabl), type = "html")
```

### Spearman correlations between fire weather and topo covariates only
SurfAspectRatio adn DegSlope removed due to high correlation with PctSlope

```{r weatherTopoCorrSpearman, echo = FALSE, cache = TRUE}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|DegSlope|SurfAspectRatio|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|HEIGHT|CROWN|ORIGIN|pixID|Czone|^NR1|^NR2|startJulDay|Lari|Pice|Pinu|Abie|Popu|Betu",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
print(xtable::xtable(tabl), type = "html")
```

### Pearson correlations between severity and vegetation and topo
Correlations are significant but very low

```{r sevVegTopoCorrPearson, echo = FALSE, cache = TRUE}
cols <- grep("SEV_CLASS|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))
tabl <- tabl[order(tabl[, "SEV_CONT"], decreasing = TRUE),, drop = FALSE]
print(xtable::xtable(tabl), type = "html")
```

### Pearson correlations between severity and weather
Correlations are significant but very low (only slightly higher than those with veg)

```{r sevWeatherTopoCorrPearson, echo = FALSE, cache = TRUE}
cols <- grep("SEV_CONT|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))
tabl <- tabl[order(tabl[, "SEV_CONT"], decreasing = TRUE),, drop = FALSE]
print(xtable::xtable(tabl), type = "html")
```

### Spearman correlations between severity and vegetation and topo
Correlations imporved in some cases and got worse in others.

```{r sevVegTopoCorrSpearman, echo = FALSE, cache = TRUE}
cols <- grep("SEV_CLASS|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))
tabl <- tabl[order(tabl[, "SEV_CONT"], decreasing = TRUE),, drop = FALSE]
print(xtable::xtable(tabl), type = "html")
```

### Spearman correlations between severity and weather
Correlations are significant but very low (only slightly higher than those with veg)

```{r sevWeatherTopoCorrSpearman, echo = FALSE, cache = TRUE}
cols <- grep("SEV_CONT|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))
tabl <- tabl[order(tabl[, "SEV_CONT"], decreasing = TRUE),, drop = FALSE]
print(xtable::xtable(tabl), type = "html")
```