---
title: "Dave Andison's fire data Severity models"
author: "Ceres Barros"
date: "02/04/2019"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/GitHub/LandscapesInMotion/")
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)

library(data.table)
library(ggplot2)
library(reshape2)
library(sf)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(gamlss.inf)
library(gamlss)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE,
        "reproducible.cachePath" = "analyses/cache")

source("analyses/R_tools/summarizeABSK_AllData.R")
source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R")
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("R/R_tools/summary.gamlssinf0to1_FIXED.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE
```


```{r dataprep, include = FALSE, cache = FALSE}
## source dataprep:
ABSK_AllData <- reproducible::Cache(ABSKfires_DataPrep,
                                    fireDataPath = "data/fires_Dave/fireSev",
                                    vegDataPath = "data/fires_Dave/prefireVeg",
                                    topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
                                    weatherDataPath = "data/fires_Dave/fireWeather",
                                    doCache = doCache, bindAllFires = FALSE, 
                                    cacheRepo = "analyses/cache",
                                    userTags = "ABSKfires_DataPrep")

## exclude event perimeter patch type
memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) all(is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
zeroCols <- sapply(ABSK_AllData, FUN = function(x) all(x == 0, na.rm = TRUE)) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!zeroCols))]

## calculate stand age at fire year by subtracting fire year - makes negative values
ABSK_AllData[!is.na(ORIGIN_UPPER), STAND_AGE := FIRE_YEAR - ORIGIN_UPPER]

## convert SMR to ordinal variable (moister as values increase)
ABSK_AllData[SMR == "D", SMR_ord := 1]
ABSK_AllData[SMR == "F", SMR_ord := 2]
ABSK_AllData[SMR == "M", SMR_ord := 3]
ABSK_AllData[SMR == "W", SMR_ord := 4]
ABSK_AllData[SMR == "A", SMR_ord := 5]

ABSK_AllData[, SMR_ord := as.ordered(SMR_ord)]

## summarise weather data - faster if not cached
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             dim = dim(ABSK_AllData),
                             saveDir = "analyses/fireDataSummary/",
                             overwrite = FALSE, omitArgs = c("DT"))

## collapse species into more generic categories
summaryABSK_AllData[, FlamConif := sum(`Pice glau`, `Pice mari`, `Pice enge`,
                                       `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Abie := sum(`Abie bals`, `Abie lasi`, na.rm = TRUE),
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 
dim(summaryABSK_AllData)

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")

collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if(all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]

summaryABSK_AllData[, isForest := 0]
summaryABSK_AllData[NonForestValue == "NA", isForest := 1]

## change lari laris to simpler name
setnames(summaryABSK_AllData, "Lari lari", "Lari")

## rescale severity from 0-1
summaryABSK_AllData[, SEV_PROP := SEV_CONT/100]

## change some columns to numeric
cols <- c("SEV_PROP", "meanFFMC", "cvDMC", "cvBUI", "minFFMC", "rangeFFMC", "cvFFMC", "minDMC", "minBUI", "FlamConif", "Abie",
          "Decid", "Lari", "UNDERSTOREY", "ORIGIN_UPPER", "PctSlope", "SurfAspectRatio", "Aspect", "CTI", "SurfReliefRatio", 
          "Elevation", "isForest", "FIRE_NAME")
cols <- grep("FIRE_NAME", cols, invert = TRUE, value = TRUE)
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## make fire name a factor
summaryABSK_AllData[, FIRE_NAME := as.factor(FIRE_NAME)]

## convert the highest fire severity to 1
summaryABSK_AllData[SEV_PROP == 0.9750, SEV_PROP := 1.0]

rm(NAcols, zeroCols, ABSK_AllData, cols)
amc::.gc()
```


Variables selected for models:
* Climate:
+ meanFFMC	-0.16***
+ cvDMC	0.16***
+ cvBUI	0.15***
+ minFFMC	-0.14***
+ rangeFFMC	0.14*** - highly correlated with minFFMC
+ cvFFMC	0.13*** - highly correlated with minFFMC
+ minDMC	-0.13*** - highly correlated with cvDMC
+ minBUI	-0.13*** - - highly correlated with min/cvDMC
* Topography
+ PctSlope	0.08***
+ SurfAspectRatio	0.08***
+ Aspect	0.07***
+ CTI	-0.07*** ([Compound Topographic Index (CTI) (Moore et al. 1991) or "soil wetness" transformation can be used to models aspects of hydrologic systems. CTI is strongly correlated with soil moisture](https://wikispaces.psu.edu/display/AnthSpace/Compound+Topographic+Index))
+ SurfReliefRatio	0.04*** ([Describes rugosity. Ratio between relief and length in a particular window](https://en.wikipedia.org/wiki/Relief_ratio))
+ Elevation	-0.03***
* Vegetation - Spearmand correlations with severity
+ Group flammable conifers 0.10 ***
+ Pinu bank	0.15***
+ Pice mari	0.06***
+ Pice glau	-0.04***
+ Pinu cont	-0.04***
+ Pice enge	0.01***
+ Group Larix
+ Lari lari	-0.05***
+ Group Abies 0.00***
+ Abie lasi	0.00
+ Abie bals	0.01***
+ Group deciduous 0.03***
+ Betu papy	-0.03***
+ Popu trem	0.03***
+ Popu balb	0.02***

+ UNDERSTOREY
+ ORIGIN_UPPER 0.10***
* Landcover type
+ isForest

```{r modelData, include = FALSE}
set.seed(123)
samp <- sample(1:nrow(na.omit(summaryABSK_AllData[, ..cols])), 
               size = round(nrow(na.omit(summaryABSK_AllData[, ..cols]))/3),
               replace = FALSE)
## variables for nanlysis
cols <- c("SEV_PROP", "meanFFMC", "cvDMC", "cvBUI", "minFFMC", "rangeFFMC", "cvFFMC", "minDMC", "minBUI", "FlamConif", "Abie",
          "Decid", "Lari", "UNDERSTOREY", "ORIGIN_UPPER", "PctSlope", "SurfAspectRatio", "Aspect", "CTI", "SurfReliefRatio", 
          "Elevation", "isForest", "FIRE_NAME")

modelData <- na.omit(summaryABSK_AllData[, ..cols])[samp]
modelData[, FIRE_NAME := as.factor(as.character(FIRE_NAME))]   ## reconvert to remove levels that were excluded by  sampling
```

```{r gamlssModels, results = "asis"}
modelNULL <- gamlss(SEV_PROP ~ 1 + re(random = ~ 1|FIRE_NAME),
                    family = BEINF, data = modelData)
# model1 <- stepGAIC(modelNULL,
#                    scope = ~ meanFFMC + cvDMC + cvBUI + minFFMC + rangeFFMC + cvFFMC + minDMC + minBUI + FlamConif +
#                      Abie + Decid + Lari + UNDERSTOREY + ORIGIN_UPPER + PctSlope + SurfAspectRatio + Aspect + CTI +
#                      SurfReliefRatio + Elevation + isForest + re(random = ~ 1|FIRE_NAME),
#                    direction = "both", k = log(length(samp)))
model1 <- gamlss(formula = SEV_PROP ~  Decid + CTI + Lari + UNDERSTOREY + re(random = ~1 | FIRE_NAME),
                 family = BEINF, data = modelData)C

# model2 <- stepGAIC(model1,
#                    what = "sigma",
#                    scope = ~ meanFFMC + cvDMC + cvBUI + minFFMC + rangeFFMC + cvFFMC + minDMC + minBUI + FlamConif +
#                      Abie + Decid + Lari + UNDERSTOREY + ORIGIN_UPPER + PctSlope + SurfAspectRatio + Aspect + CTI +
#                      SurfReliefRatio + Elevation + isForest + re(random = ~ 1|FIRE_NAME),
#                    direction = "both", k = log(length(samp)))

model2 <- gamlss(SEV_PROP ~ Decid + UNDERSTOREY + CTI + re(random = ~1 | FIRE_NAME),
                 sigma.formula = ~ Decid + Elevation + PctSlope + re(random = ~1 | FIRE_NAME),
                 family = BEINF, data = modelData)

# model3 <- stepGAIC(model2,
#                    what = "nu",
#                    scope = ~ meanFFMC + cvDMC + cvBUI + minFFMC + rangeFFMC + cvFFMC + minDMC + minBUI + FlamConif +
#                      Abie + Decid + Lari + UNDERSTOREY + ORIGIN_UPPER + PctSlope + SurfAspectRatio + Aspect + CTI +
#                      SurfReliefRatio + Elevation + isForest + re(random = ~ 1|FIRE_NAME),
#                    direction = "both", k = log(length(samp)))

model3 <- gamlss(SEV_PROP ~ Decid + UNDERSTOREY + CTI + re(random = ~1 | FIRE_NAME),
                 sigma.formula = ~ Decid + UNDERSTOREY + CTI + PctSlope + Abie + Aspect + re(random = ~1 | FIRE_NAME),
                 nu.formula = ~ FlamConif + ORIGIN_UPPER + Elevation + PctSlope + SurfAspectRatio + Aspect + Lari + CTI +
                   re(random = ~1 | FIRE_NAME),
                 family = BEINF, data = modelData)

# model4 <- stepGAIC(model3,
#                    what = "tau",
#                    scope = ~ meanFFMC + cvDMC + cvBUI + minFFMC + rangeFFMC + cvFFMC + minDMC + minBUI + FlamConif +
#                      Abie + Decid + Lari + UNDERSTOREY + ORIGIN_UPPER + PctSlope + SurfAspectRatio + Aspect + CTI +
#                      SurfReliefRatio + Elevation + isForest + re(random = ~ 1|FIRE_NAME),
#                    direction = "both", k = log(length(samp)))

## tau remained unchanged
model4 <- gamlss(SEV_PROP ~ Decid + UNDERSTOREY + CTI + re(random = ~1 | FIRE_NAME),
                 sigma.formula = ~ Decid + UNDERSTOREY + CTI + PctSlope + Abie + Aspect + re(random = ~1 | FIRE_NAME),
                 nu.formula = ~ FlamConif + ORIGIN_UPPER + Elevation + PctSlope + SurfAspectRatio + Aspect + Lari + CTI +
                   re(random = ~1 | FIRE_NAME),
                 tau.formula = ~  Decid + UNDERSTOREY + Aspect + PctSlope + SurfAspectRatio + CTI + Abie + Elevation + re(random = ~1 | FIRE_NAME),
                 family = BEINF, data = modelData)

GAIC(modelNULL,
     model1,
     model2,
     model3,  
     model4,  ## best
     k = log(nrow(modelData)))

plot(modelNULL)   ## pretty horrible
plot(model4)   ## pretty good!!!!

```


Let's try with a hurdle model, starting from the previous best model
Note that \code{gamlssInf0to1} doesn't seem to like \code{re} in this case.
`nu` and `tau` are also not used and the inflation at 0 and 1 is taken care of by fitting distinct distributions. SO I keep the same
co-variates but transfer them to `xi0` and `xi1`, respectively
```{r hurdleModels, results = "asis"}
modelHurdle <- gamlssInf0to1(y = SEV_PROP,
                             mu.formula = ~ Decid + CTI + Lari + UNDERSTOREY + random(FIRE_NAME),
                             sigma.formula = ~ Decid + Elevation + PctSlope + random(FIRE_NAME),
                             xi0.formula = ~ FlamConif + ORIGIN_UPPER + Elevation + PctSlope + SurfAspectRatio +
                               Aspect + Lari + CTI + random(FIRE_NAME),
                             xi1.formula = ~ Decid + UNDERSTOREY + Aspect + PctSlope + SurfAspectRatio +
                               CTI + Abie + Elevation + random(FIRE_NAME),
                             control = gamlss.control(n.cyc = 40),
                             data = modelData)
## inspect p-values and exclude non-significant effects/add interactions
summary.gamlssinf0to1_2(modelHurdle)
plot(modelHurdle)

modelHurdle2 <- gamlssInf0to1(y = SEV_PROP,
                              mu.formula = ~ Decid*UNDERSTOREY + CTI + Lari + random(FIRE_NAME),
                              sigma.formula = ~ Decid*Elevation + Elevation*PctSlope + random(FIRE_NAME),
                              xi0.formula = ~ FlamConif + ORIGIN_UPPER + Elevation*PctSlope + SurfAspectRatio*Aspect + 
                                Lari + CTI + random(FIRE_NAME),
                              xi1.formula = ~ Decid*UNDERSTOREY + SurfAspectRatio*Aspect + PctSlope +
                                CTI + Abie*UNDERSTOREY + Elevation + random(FIRE_NAME),
                              data = modelData,
                              control = gamlss.control(n.cyc = 30))
summary.gamlssinf0to1_2(modelHurdle2)

## add/ covariates for the 0 inflated component first
modelHurdle3 <- gamlssInf0to1(y = SEV_PROP,
                              mu.formula = ~ Decid*UNDERSTOREY + CTI + Lari*UNDERSTOREY + random(FIRE_NAME),
                              sigma.formula = ~ Decid*Elevation + Elevation*PctSlope + Decid*PctSlope + random(FIRE_NAME),
                              xi0.formula = ~ FlamConif*Elevation + FlamConif*ORIGIN_UPPER + FlamConif*PctSlope + SurfAspectRatio*Aspect + 
                                Lari + CTI + random(FIRE_NAME),
                              xi1.formula = ~ Decid*UNDERSTOREY + SurfAspectRatio*Aspect + PctSlope +
                                CTI + Abie*Elevation + random(FIRE_NAME),
                              data = modelData,
                              control = gamlss.control(n.cyc = 40))
summary(modelHurdle3)



GAIC(modelNULL,
     model4,  ## best
     modelHurdle,
     modelHurdle2,
     modelHurdle3,
     modelHurdle4,
     modelHurdle5,
     modelHurdle6,
     modelHurdle7,
     modelHurdle8,
     k = log(nrow(modelData)))

plot(model1)   ## pretty horrible
plot(model3)   ## as bad as the previous.
plot(modelHurdle) ## still pretty bad.

```

