---
title: "Dave Andison's fire data Severity models"
author: "Ceres Barros"
date: "02/04/2019"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/GitHub/LandscapesInMotion/")
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)

library(data.table)
library(reshape2)
library(sf)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(gamlss.inf)
library(gamlss)
library(ggplot2)
library(ggpubr)
library(ggrepel)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE,
        "reproducible.cachePath" = "analyses/cache")

source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R")
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("R/R_tools/summary.gamlssinf0to1_FIXED.R")
source("R/R_tools/Rsq_FIXED.R")
source("R/R_tools/crossValidFunction.R")
source("analyses/R_tools/summarizeABSK_AllData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE
```


```{r dataprep, include = FALSE, cache = FALSE}
## run dataprep:
ABSK_AllData <- reproducible::Cache(ABSKfires_DataPrep,
                                    fireDataPath = "data/fires_Dave/fireSev",
                                    vegDataPath = "data/fires_Dave/prefireVeg",
                                    topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
                                    weatherDataPath = "data/fires_Dave/fireWeather",
                                    resolution = 30,
                                    doCache = doCache, bindAllFires = TRUE, 
                                    cacheRepo = "analyses/cache", 
                                    userTags = "ABSKfires_DataPrep")

## exclude event perimeter patch type
# memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) all(is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
zeroCols <- sapply(ABSK_AllData, FUN = function(x) all(x == 0, na.rm = TRUE)) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!zeroCols))]

## calculate stand age at fire year by subtracting fire year - makes negative values
ABSK_AllData[!is.na(ORIGIN_UPPER), STAND_AGE := FIRE_YEAR - ORIGIN_UPPER]

## convert SMR to ordinal variable (moister as values increase)
ABSK_AllData[SMR == "D", SMR_ord := 1]
ABSK_AllData[SMR == "F", SMR_ord := 2]
ABSK_AllData[SMR == "M", SMR_ord := 3]
ABSK_AllData[SMR == "W", SMR_ord := 4]
ABSK_AllData[SMR == "A", SMR_ord := 5]

ABSK_AllData[, SMR_ord := as.ordered(SMR_ord)]

## summarise weather data - faster if not cached
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             dim = dim(ABSK_AllData),
                             days = 1:3,
                             omitArgs = c("DT"),
                             useCache = doCache)

## collapse species into more generic categories
summaryABSK_AllData[, FlamConif := sum(`Pice glau`, `Pice mari`, `Pice enge`,
                                       `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Abie := sum(`Abie bals`, `Abie lasi`, na.rm = TRUE),
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")

collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if(all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]

summaryABSK_AllData[, isForest := 0]
summaryABSK_AllData[NonForestValue == "NA", isForest := 1]

## change Lari lari to simpler name
setnames(summaryABSK_AllData, "Lari lari", "Lari")

## rescale severity from 0-1
summaryABSK_AllData[, SEV_PROP := SEV_CONT/100]
## convert the highest fire severity to 1
summaryABSK_AllData[SEV_PROP == 0.9750, SEV_PROP := 1.0]

## change some columns to numeric
cols <- grep("mean|max|min|cv|range|pixID|P_ID|Duration|ORIGIN|HEIGHT|isForest|SEV_PROP|Aspect|^CTI$|PctSlope|SurfReliefRatio|SurfAspectRatio|Elevation", names(summaryABSK_AllData), value = TRUE)
str(summaryABSK_AllData[, ..cols])
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## calculate SMR_wet
summaryABSK_AllData[, SMR_wet := as.integer(grepl("4|5", SMR_ord))]

## make some columns factors
cols <- c("FIRE_NAME", "Czone1", "NR1")
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.factor), .SDcols = cols]

rm(NAcols, zeroCols, ABSK_AllData, cols)
amc::.gc()
```

```{r addNgbSev, cache = FALSE, include = FALSE}
## make different buffer sizes around points
crsProj <- crs(shapefile("data/fires_Dave/fireSev/albertafires1_postfire.shp"))

SEV_PROPpoints <- summaryABSK_AllData[, .(pixID, SEV_PROP, Lat, Long)]
SEV_PROPpoints <- st_as_sf(SEV_PROPpoints, coords = c("Long", "Lat"),
                           agr = "constant", crs = crsProj)

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i-1] * 2
}

nbgSev <- reproducible::Cache(calculateNgbSevWrapper,
                              dists = bufferDists,
                              sevPoints = SEV_PROPpoints, 
                              sevColID = "SEV_PROP",
                              useCache = doCache,
                              cacheRepo = "analyses/cache",
                              userTags = "nbgSev") 
summaryABSK_AllData <- nbgSev[summaryABSK_AllData, on = .(pixID)]
```


```{r addNgbNoBurns, cache = FALSE, include = FALSE}
## make different buffer sizes around points
crsProj <- crs(shapefile("data/fires_Dave/fireSev/albertafires1_postfire.shp"))

SEV_PROPpoints <- summaryABSK_AllData[, .(pixID, FIRE_NAME, SEV_PROP, Lat, Long)]
SEV_PROPpoints <- st_as_sf(SEV_PROPpoints, coords = c("Long", "Lat"),
                           agr = "constant", crs = crsProj)

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i-1] * 2
}

nbgNoBurns <- reproducible::Cache(calculateNgbBurnsWrapper,
                                  dists = bufferDists,
                                  sevPoints = SEV_PROPpoints, 
                                  sevColID = "SEV_PROP",
                                  fireColID = "FIRE_NAME",
                                  resolution = 30,
                                  parallel = FALSE,
                                  useCache = doCache,
                                  cacheRepo = "analyses/cache",
                                  userTags = "nbgSev") 
summaryABSK_AllData <- nbgSev[summaryABSK_AllData, on = .(pixID)]
```

## SUMMARY OF PRELIMINARY ANALYSES

Variables selected for models (when using all fire weather days):
* Climate:
+ meanFFMC	-0.16***
+ cvDMC	0.16***
+ cvBUI	0.15***
+ minFFMC	-0.14***
+ rangeFFMC	0.14*** - highly correlated with minFFMC
+ cvFFMC	0.13*** - highly correlated with minFFMC
+ minDMC	-0.13*** - highly correlated with cvDMC
+ minBUI	-0.13*** - - highly correlated with min/cvDMC
* Topography
+ PctSlope	0.08***
+ SurfAspectRatio	0.08*** - Ratio of the centerline-to-centerline spacing between dimples to the height of the dimple (https://arxiv.org/ftp/arxiv/papers/1305/1305.4705.pdf) / Surface Relief Ratio - describes rugosity in an continuous raster surface within a specified window (https://evansmurphy.wixsite.com/evansspatial/arcgis-gradient-metrics-toolbox)
+ Aspect	0.07***
+ CTI	-0.07*** ([Compound Topographic Index (CTI) (Moore et al. 1991) or "soil wetness" transformation can be used to models aspects of hydrologic systems. CTI is strongly correlated with soil moisture](https://wikispaces.psu.edu/display/AnthSpace/Compound+Topographic+Index))
+ SurfReliefRatio	0.04*** ([Describes rugosity. Ratio between relief and length in a particular window](https://en.wikipedia.org/wiki/Relief_ratio))
+ Elevation	-0.03***
* Vegetation - Spearmand correlations with severity
+ Group flammable conifers 0.10 ***
+ Pinu bank	0.15***
+ Pice mari	0.06***
+ Pice glau	-0.04***
+ Pinu cont	-0.04***
+ Pice enge	0.01***
+ Group Larix
+ Lari lari	-0.05***
+ Group Abies 0.00***
+ Abie lasi	0.00
+ Abie bals	0.01***
+ Group deciduous 0.03***
+ Betu papy	-0.03***
+ Popu trem	0.03***
+ Popu balb	0.02***

+ UNDERSTOREY
+ ORIGIN_UPPER 0.10***
* Landcover type
+ isForest

```{r modelData, include = FALSE}
## variables for analysis when considering all weather (first 3 days) - dropping range
## due to lack of data in some fires, dropping:
## crown closure, SMR (which has a proxy- CTI), cvRain, ORIGIN_*
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "Czone1|NR1|FIRE_NAME|pixID", sep = "|"), names(summaryABSK_AllData),
             value = TRUE)

## sample a forth of the pixels in each fire
sampDT <- na.omit(summaryABSK_AllData[, ..cols])[,.(FIRE_NAME, pixID)]
sampDT[, noPixels :=  length(pixID), by = FIRE_NAME]
set.seed(123)
samp <- sampDT[, sample(pixID, size = noPixels/4, replace = FALSE), by = FIRE_NAME]$V1
rm(sampDT)

modelData <- na.omit(summaryABSK_AllData[, ..cols])[pixID %in% samp]

## check we didn't lose fires
if (length(setdiff(unique(summaryABSK_AllData$FIRE_NAME), 
                   unique(modelData$FIRE_NAME))))
  stop("Fires lost in sampling!")
```

## GAMLSS models
Selection of variables going "in".
Start with a high covergence criterion to speed up model fitting for now
```{r gamlssModels, results = "asis", message=FALSE}
modelNULL <- gamlss(SEV_PROP ~ 1 + re(random = ~ 1|Czone1/NR1/FIRE_NAME),
                    family = BEINF, control = gamlss.control(c.crit = 0.1),
                    data = modelData)

# model1 <- stepGAIC(modelNULL,
#                    what = "mu",
#                    scope = list(lower = ~ re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                                 upper = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC +
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI +
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME)),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 20,
#                    direction = "both", k = log(length(samp)))

## 20m res
# model1 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)

## 30m res
model1 <- gamlss(formula = SEV_PROP ~ Decid + CTI + Lari + HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model2 <- stepGAIC(model1,
#                    what = "sigma",
#                    scope = list(lower = ~re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                                 upper = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC +
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI +
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME)),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 20,
#                    direction = "both", k = log(length(samp)))

## 20m res
# model2 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)

## 30m res
model2 <- gamlss(formula = SEV_PROP ~ Decid + CTI + Lari + HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model3 <- stepGAIC(model2,
#                    what = "nu",
#                    scope = list(lower = ~re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                                 upper = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC +
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI +
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME)),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 20,
#                    direction = "both", k = log(length(samp)))

## 20m res
# model3 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  nu.formula = ~ Decid + FlamConif + SurfAspectRatio + UNDERSTOREY + SurfReliefRatio + PctSlope +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)

## 30m res
model3 <- gamlss(formula = SEV_PROP ~ Decid + CTI + Lari + HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ FlamConif + UNDERSTOREY + SurfAspectRatio + Decid + PctSlope + SurfReliefRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model4 <- stepGAIC(model3,
#                    what = "tau",
#                    scope = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC +
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI +
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 18,
#                    direction = "both", k = log(length(samp)))

## 20m res
# model4 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  nu.formula = ~ Decid + FlamConif + SurfAspectRatio + UNDERSTOREY + SurfReliefRatio + PctSlope +
#                    re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
#                  tau.formula = ~ FlamConif + Lari + HEIGHT_LOWER + UNDERSTOREY + CTI + Abie + SurfAspectRatio +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)

## 30m res
model4 <- gamlss(formula = SEV_PROP ~ Decid + CTI + Lari + HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ FlamConif + UNDERSTOREY + SurfAspectRatio + Decid + PctSlope + SurfReliefRatio +
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ Decid + HEIGHT_LOWER + UNDERSTOREY + CTI + isForest + SurfReliefRatio + Abie + Lari +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

GAIC(modelNULL,
     model1,
     model2,
     model3,  
     model4,  ## best
     k = log(length(samp)))

plot(modelNULL)   ## pretty horrible (with bad sampling) | pretty good  (good sampling and more REs)
plot(model4)   ## pretty good!!!!  (with bad sampling) | looks even better (good sampling and more REs)
```

## GAMLSS INF models
Let's try with the GAMLSS INF (probabilities at 0s and 1s have more direct interpretation), starting from the previous best model
Note that \code{gamlssInf0to1} doesn't seem to like \code{re} in this case.
`nu` and `tau` are replaced by `xi0` and `xi1`, respectively, which in the case of 0 and 1 inflation are the same between `gamlss` and `gamlssInf0to1`. GAMLSS Inf fits a log-linear multinomial model to the 0s 1s and everything in between  (0 becomes 1, 1 becomes 2 and everything else becomes 3).
Note that when the sampling was corrected and more random effects added, using gamlss.inf offered no improvement
```{r gamlssinfModels, results = "asis", message=FALSE}
modelInf <- gamlssInf0to1(y = SEV_PROP,
                          mu.formula = ~ Decid + CTI + Lari + HEIGHT_LOWER +
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          sigma.formula = ~ Decid + UNDERSTOREY + 
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          xi0.formula = ~ FlamConif + UNDERSTOREY + SurfAspectRatio + Decid + PctSlope + SurfReliefRatio +
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          xi1.formula = ~ Decid + HEIGHT_LOWER + UNDERSTOREY + CTI + isForest + SurfReliefRatio + Abie + Lari +
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          family = BE, control = gamlss.control(c.crit = 0.1),
                          data = modelData)

## inspect p-values and exclude non-significant effects/add interactions
summary.gamlssinf0to1_2(modelInf)
plot(modelInf)

GAIC(modelNULL,
     model4,  
     modelInf, ## this is actually worse.
     k = log(length(samp)))  
```

## Adding/removing covariates and interactions 
Go back to "standard" GAMLSS, since GAMLSS Inf did not improve models. The advantage of using GAMLSS is that more methods are available. Also, because we are using a beta inflates distribution there's not much gain in using `gamlssInf0to1`.
We can use re() again.

The coefficients are pratically the same (between GAMLSS and GAMLSSInf)

We can get the fitted probabilities at 0 (p0) and 1 (p1) by reversing the following equations:
xi0 = nu = p0 / (1 + p0 + p1)
xi1 = tau = p1 / (1 + p0 + p1)

hence,
p0 = xi0 / (1 + xi0 + xi1) = nu / (1 + nu + tau)
p1 = xi1 / (1 + xi0 + xi1) = tau / (1 + nu + tau)

Start by adding interactions within covariate groups (vegetation, topography, climate).
Not all possible 2-way interations were added because that made the model not converge.

```{r addInteractions, message=FALSE}
model5 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari*HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid*UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ FlamConif*UNDERSTOREY + Decid*UNDERSTOREY +
                   PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ isForest*HEIGHT_LOWER + isForest*UNDERSTOREY +
                   Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                   Abie*HEIGHT_LOWER + Abie*UNDERSTOREY + 
                   Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                   CTI+SurfReliefRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)
summary(model5)

## remove non-significant main effects and interactions, and add others 
## tried adding other interactions to mu/sigma/nu/tau, but it creates a singular model

# removed Abie from tau since neither interactions nor main eeffect were significant
model6 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid*UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                   PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ isForest + isForest*UNDERSTOREY +
                   Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                   Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                   CTI+SurfReliefRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)
summary(model6)

GAIC(model4, ## best
     # modelInf,
     model5, 
     model6, ## best model
     k = log(length(samp))) 
Rsq(model4)
Rsq(model5)  ## best R2 - but only by 0.0003
Rsq(model6)  ## will keep this model, for parsimony
```

## Try adding soil moisture from veg. inventories and drought code
```{r reDO_modelData, include = FALSE, cache = FALSE}
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "Czone1|NR1|FIRE_NAME|pixID|SMR_ord", sep = "|"), names(summaryABSK_AllData),
             value = TRUE)

## sample a forth of the pixels in each fire
modelData <- na.omit(summaryABSK_AllData[, ..cols])[pixID %in% samp]

## make a binary soil moisture variable of wet vs everything else (suggested by Dave from past analyses)
modelData[, SMR_wet := as.integer(grepl("4|5", SMR_ord))]

## check we didn't lose fires
if (length(setdiff(unique(summaryABSK_AllData$FIRE_NAME), 
                   unique(modelData$FIRE_NAME))))
  stop("Fires lost in sampling!")
```

```{r addSM_DC, message=FALSE}
## add full SMR variable first - this was not possible at 30m res
## 20 res
# model7 <- gamlss(SEV_PROP ~ CTI + isForest + Lari*HEIGHT_LOWER +   ## Lari*Height no longer significant
#                    FlamConif*HEIGHT_LOWER + SMR_ord +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  sigma.formula = ~ Elevation + Decid*UNDERSTOREY + SMR_ord +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
#                    SMR_ord +  PctSlope + SurfAspectRatio + SurfReliefRatio + 
#                    re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
#                  tau.formula = ~ FlamConif + Lari + Abie + 
#                    UNDERSTOREY + HEIGHT_LOWER + SMR_ord + SurfAspectRatio +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)
# summary(model7)

## try removing CTI - this was not possible at 30m res
## 20 res
# model7.2 <- gamlss(SEV_PROP ~ isForest + Lari*HEIGHT_LOWER +     ## Lari*Height no longer significant
#                      FlamConif*HEIGHT_LOWER + SMR_ord +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    sigma.formula = ~ Elevation + Decid*UNDERSTOREY + SMR_ord +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
#                      SMR_ord +  PctSlope + SurfAspectRatio + SurfReliefRatio + 
#                      re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
#                    tau.formula = ~ FlamConif + Lari + Abie + 
#                      UNDERSTOREY + HEIGHT_LOWER + SMR_ord + SurfAspectRatio +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    family = BEINF, control = gamlss.control(c.crit = 0.1),
#                    data = modelData)

# summary(model7.2)

## add binary SMR variable
## 20 res
# model8 <- gamlss(SEV_PROP ~ CTI + isForest + Lari*HEIGHT_LOWER +   ## Lari*Height no longer significant
#                    FlamConif*HEIGHT_LOWER + SMR_wet +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  sigma.formula = ~ Elevation + Decid*UNDERSTOREY + SMR_wet +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
#                    SMR_ord +  PctSlope + SurfAspectRatio + SurfReliefRatio + 
#                    re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
#                  tau.formula = ~ FlamConif + Lari + Abie + 
#                    UNDERSTOREY + HEIGHT_LOWER + SMR_ord + SurfAspectRatio +
#                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                  family = BEINF, control = gamlss.control(c.crit = 0.1),
#                  data = modelData)

model8 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                   SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Decid*UNDERSTOREY +
                   SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                   PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                   SMR_wet + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ isForest + isForest*UNDERSTOREY +
                   Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                   Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                   CTI + SurfReliefRatio + SMR_wet +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)
summary(model8)

## try removing CTI
## 20 res - this was not possible at 30m res
# model8.2 <- gamlss(SEV_PROP ~ isForest + Lari*HEIGHT_LOWER + 
#                      FlamConif*HEIGHT_LOWER + SMR_wet +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    sigma.formula = ~ Elevation + Decid*UNDERSTOREY + SMR_wet +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
#                      SMR_ord +  PctSlope + SurfAspectRatio + SurfReliefRatio + 
#                      re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
#                    tau.formula = ~ FlamConif + Lari + Abie + 
#                      UNDERSTOREY + HEIGHT_LOWER + SMR_ord + SurfAspectRatio +
#                      re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                    family = BEINF, control = gamlss.control(c.crit = 0.1),
#                    data = modelData)

GAIC(model5,
     # model7, ## best for res 20m
     # model7.2,
     model8,  ## best for res 30m
     # model8.2,
     k = log(nrow(modelData)))

## AIC is worse in all cases if DC is added
noCores <- detectCores()
## 20 res
# upperFormula <- as.formula(paste(
#   paste(as.character(model7$mu.formula)[c(1,3)], collapse = " "), 
#   "+",
#   paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
## of all DC variabvles the least bad are cvDC and maxDC (equal AIC)
# model9 <- stepGAIC(model7, what = "mu",
#                    scope = list(upper = upperFormula,
#                                 lower = model7$mu.formula),
#                    direction = "both", k = log(length(samp)), 
#                    parallel = "multicore", ncpus = noCores - 20)

## 30 res
upperFormula <- as.formula(paste(
  paste(as.character(model8$mu.formula)[c(1,3)], collapse = " "), 
  "+",
  paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
model9 <- stepGAIC(model8, what = "mu",
                   scope = list(upper = upperFormula,
                                lower = model8$mu.formula),
                   direction = "both", k = log(length(samp)),
                   parallel = "multicore", ncpus = 10)

## of all DC variables the least bad are minDC, meanDC and maxDC (equal AIC)
## 20 res
# upperFormula <- as.formula(paste(
#   paste(as.character(model7$sigma.formula), collapse = " "), 
#   "+",
#   paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
# model10 <- stepGAIC(model9, what = "sigma",
#                     scope = list(upper = upperFormula,
#                                  lower = model7$sigma.formula),
#                     direction = "both", k = log(length(samp)),
#                     parallel = "multicore", ncpus = noCores - 20)

## 30 res
upperFormula <- as.formula(paste(
  paste(as.character(model8$sigma.formula), collapse = " "), 
  "+",
  paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
model10 <- stepGAIC(model9, what = "sigma",
                    scope = list(upper = upperFormula,
                                 lower = model8$sigma.formula),
                    direction = "both", k = log(length(samp)),
                    parallel = "multicore", ncpus = 10)


## of all DC variabvles the least bad is cvDC (equal AIC)
## 20 res
# upperFormula <- as.formula(paste(
#   paste(as.character(model7$nu.formula), collapse = " "), 
#   "+",
#   paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
# model11 <- stepGAIC(model10, what = "nu",
#                     scope = list(upper = upperFormula,
#                                  lower = model7$nu.formula),
#                     direction = "both", k = log(length(samp)), 
#                     parallel = "multicore", ncpus = noCores - 20)  


## 30 res
upperFormula <- as.formula(paste(
  paste(as.character(model8$nu.formula), collapse = " "), 
  "+",
  paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
model11 <- stepGAIC(model10, what = "nu",
                    scope = list(upper = upperFormula,
                                 lower = model8$nu.formula),
                    direction = "both", k = log(length(samp)),
                    parallel = "multicore", ncpus = 10)


## of all DC variabvles the least bad is rangeDC (equal AIC)
## 20 res
# upperFormula <- as.formula(paste(
#   paste(as.character(model7$tau.formula), collapse = " "), 
#   "+",
#   paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
# model12 <- stepGAIC(model11, what = "tau",
#                     scope = list(upper = upperFormula,
#                                  lower = model7$tau.formula),
#                     direction = "both", k = log(length(samp)), 
#                     parallel = "multicore", ncpus = noCores - 20)    

## 30 res
upperFormula <- as.formula(paste(
  paste(as.character(model8$tau.formula), collapse = " "), 
  "+",
  paste(grep("DC$", names(modelData), value = TRUE), collapse = " + ")))
model12 <- stepGAIC(model11, what = "tau",
                    scope = list(upper = upperFormula,
                                 lower = model8$tau.formula),
                    direction = "both", k = log(length(samp)),
                    parallel = "multicore", ncpus = 10)

## force best DC variables in - stepAIC didn't manage to add any
## best variables decided based on which had lowest in AIC trials
## singularity found when maxDC added to mu, meanDC added to sigma, cvDC added to nu and rangeDV added to tau
## these variables were removed progressively, from tau to mu.
## 20 res
# model13 <- gamlss(SEV_PROP ~ CTI + isForest + Lari * HEIGHT_LOWER + FlamConif * HEIGHT_LOWER + 
#                     SMR_ord + maxDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   sigma.formula = ~ Elevation + Decid * UNDERSTOREY + 
#                     SMR_ord + meanDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   nu.formula = ~ UNDERSTOREY * FlamConif + UNDERSTOREY * Decid + 
#                     SMR_ord + PctSlope + cvDC +
#                     SurfAspectRatio + SurfReliefRatio + 
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   tau.formula = ~ FlamConif + Lari + Abie + UNDERSTOREY + HEIGHT_LOWER + 
#                     SMR_ord + SurfAspectRatio + #rangeDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),  
#                   family = BEINF, data = modelData, 
#                   control = gamlss.control(c.crit = 0.1)) 

## res30
model13 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                    SMR_wet + rangeDC + 
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  sigma.formula = ~ Decid*UNDERSTOREY + SMR_wet +
                    meanDC + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                    PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                    SMR_wet + meanDC + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                  tau.formula = ~ isForest + isForest*UNDERSTOREY +
                    Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                    Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                    CTI + SurfReliefRatio + SMR_wet + minDC +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  family = BEINF, control = gamlss.control(c.crit = 0.1),
                  data = modelData)
summary(model13)

## res 30m - remove non-significant variables
model14 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                    SMR_wet +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  sigma.formula = ~ Decid*UNDERSTOREY + SMR_wet +
                    meanDC + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  nu.formula = ~ FlamConif + UNDERSTOREY +
                    PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                    SMR_wet + meanDC + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                  tau.formula = ~ isForest + isForest*UNDERSTOREY +
                    Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                    Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                    CTI + SurfReliefRatio + SMR_wet + minDC +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  family = BEINF, control = gamlss.control(c.crit = 0.1),
                  data = modelData)

GAIC(
  # model5, 
  # model7, ## 20 res still best
  model8,  ## 30 res still best 
  model9,
  model10,
  model11,
  model12,
  model13, # 20res worse than model7 by 32 AIC points (only worse by 5 points, if k = 2 is used)
  # 30res, worse than model8
  model14, # 30res better than model13, still worse than model8
  k = log(length(samp)))   
## residuals of models 9-12 are the same and have the same stats.
plot(model13) ## almost the same as model7 (res20) or model8 (30m)

# Rsq(model7)
Rsq(model8)
Rsq(model13) ## tinsy bit worse (res 20m)/ identical (res 30m)
Rsq(model14) ## (res30m) identical
```

### 20 m res
After adding SMR and DC some other variables lost significance
Will try removing them - going from `model7`
I see no reason to keep DC variables in the models. They make AIC worse and lower the r2.
Keeping SMR_ord and CTI, however may be a good idea.

```{r addSM_DC2, message=FALSE}
## RES 20M -------------------------------------
## hit singularity when non significant variables in model13 were removed, 
## so I removed DC vars one at a time, from tau->mu
# model14 <- gamlss(SEV_PROP ~ CTI + isForest + Lari + FlamConif * HEIGHT_LOWER +
#                     SMR_ord + #maxDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   sigma.formula = ~ Elevation + Decid * UNDERSTOREY +
#                     SMR_ord + #meanDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   nu.formula = ~ UNDERSTOREY * FlamConif + UNDERSTOREY * Decid +
#                     SMR_ord + PctSlope + #cvDC +
#                     SurfAspectRatio + SurfReliefRatio +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   tau.formula = ~ FlamConif + Lari + Abie + UNDERSTOREY + HEIGHT_LOWER +
#                     SMR_ord + SurfAspectRatio + #rangeDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   family = BEINF, data = modelData,
#                   control = gamlss.control(c.crit = 0.1))

## try removing CTI 
# model15 <- gamlss(SEV_PROP ~ isForest + FlamConif * HEIGHT_LOWER + 
#                     SMR_ord + #maxDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   sigma.formula = ~ Elevation + Decid * UNDERSTOREY + 
#                     SMR_ord + #meanDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   nu.formula = ~ UNDERSTOREY * FlamConif + UNDERSTOREY * Decid + 
#                     SMR_ord + PctSlope + #cvDC +
#                     SurfAspectRatio + SurfReliefRatio + 
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   tau.formula = ~ FlamConif + Lari + Abie + UNDERSTOREY + HEIGHT_LOWER + 
#                     SMR_ord + SurfAspectRatio + #rangeDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),  
#                   family = BEINF, data = modelData, 
#                   control = gamlss.control(c.crit = 0.1)) 

## try adding interactions
# model16 <- gamlss(SEV_PROP ~ CTI + (isForest + FlamConif + SMR_ord) * maxDC + 
#                     FlamConif * HEIGHT_LOWER +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                   sigma.formula = ~ Elevation + Decid * UNDERSTOREY +
#                     (Decid + SMR_ord + UNDERSTOREY) * meanDC +
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   nu.formula = ~ UNDERSTOREY * FlamConif + UNDERSTOREY * Decid +
#                     (FlamConif + Decid + UNDERSTOREY + SMR_ord) + # * cvDC +
#                     PctSlope + SurfAspectRatio + SurfReliefRatio + 
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME), 
#                   tau.formula = ~ (FlamConif + Lari + Abie + 
#                                      UNDERSTOREY + HEIGHT_LOWER + SMR_ord) + #* rangeDC  + 
#                     SurfAspectRatio + 
#                     re(random = ~1 | Czone1/NR1/FIRE_NAME),  
#                   family = BEINF, data = modelData, 
#                   control = gamlss.control(c.crit = 0.1)) 

# summary(model16)   ## adding effects with DC are significant 

# GAIC(model5, ## only CTI
#      model7, ## CTI and SMR_ord, best
#      model13,
#      model14,  # best
#      model15,
#      model16, ## worse
#      k = log(nrow(modelData)))

# plot(model14)   # residuals look virtually the same (metrics are similar too)
# plot(model13)
# plot(model7)

## models 14, 7 and 13 have v similar R2
# Rsq(model14)
# Rsq(model7)
# Rsq(model13) 
# Rsq(model16)   # r^2 is also worse

```

### 30 m res
Adding the DC variables didn't change the r2, but lowered AIC.
I will now try to remove one of the roughtness variables, then CTI and see.

```{r rmSAR_CTI, message = FALSE}
## RES 30M -------------------------------------
## removing SAR - this model fails
model15 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                    SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  sigma.formula = ~ Decid*UNDERSTOREY + SMR_wet +
                    meanDC + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  nu.formula = ~ FlamConif + UNDERSTOREY +
                    PctSlope*SurfReliefRatio +
                    SMR_wet + meanDC + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                  tau.formula = ~ isForest + isForest*UNDERSTOREY +
                    Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                    Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                    CTI + SurfReliefRatio + SMR_wet + minDC +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  family = BEINF, control = gamlss.control(c.crit = 0.1),
                  data = modelData)

## removing SAR and CTI
model16 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + Lari +
                    SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  sigma.formula = ~ Decid*UNDERSTOREY + SMR_wet +
                    meanDC + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  nu.formula = ~ FlamConif + UNDERSTOREY +
                    PctSlope*SurfReliefRatio +
                    SMR_wet + meanDC + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                  tau.formula = ~ isForest + isForest*UNDERSTOREY +
                    Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                    Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                    SurfReliefRatio + SMR_wet + minDC +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  family = BEINF, control = gamlss.control(c.crit = 0.1),
                  data = modelData)

## removing SAR from model8
model8.2 <-gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                    SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  sigma.formula = ~ Decid*UNDERSTOREY +
                    SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                    PctSlope*SurfReliefRatio +
                    SMR_wet + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                  tau.formula = ~ isForest + isForest*UNDERSTOREY +
                    Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                    Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                    CTI + SurfReliefRatio + SMR_wet +
                    re(random = ~1 | Czone1/NR1/FIRE_NAME),
                  family = BEINF, control = gamlss.control(c.crit = 0.1),
                  data = modelData)

## removing SAR and CTI from model8 - fails
model8.3 <- gamlss(formula = SEV_PROP ~ Decid*HEIGHT_LOWER + Lari +
                     SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                   sigma.formula = ~ Decid*UNDERSTOREY +
                     SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                   nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                     PctSlope*SurfReliefRatio +
                     SMR_wet + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                   tau.formula = ~ isForest + isForest*UNDERSTOREY +
                     Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                     Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                     SurfReliefRatio + SMR_wet +
                     re(random = ~1 | Czone1/NR1/FIRE_NAME),
                   family = BEINF, control = gamlss.control(c.crit = 0.1),
                   data = modelData)

GAIC(
  model8,  ## 30 res still best 
  model8.2,
  # model8.3,
  model14, 
  # model15,
  model16,  #worse
  k = log(length(samp)))   
Rsq(model8)
Rsq(model14)
Rsq(model16)  ## worse
```


## Final model
### 20 res
The final model is `model14`.

### 30 res
The model didn't improve by removing SAR/CTI, and adding DC variables decreased AIC while not improving R2.
So, for this resolution `model8` was the best model.
**However**, `model8` failed when using the whole dataset, or different subsets of data. So, I fit simpler models to see if I could get a more generic one. In the end, the next best model was one with a simpler random effects structure, using only "FIRE_NAME", while keeping SMR_wet and CTI - `model8.4`. This model will be submitted to validation. 

```{r fullDataModel, cache = FALSE, message=FALSE}
## not being able to fit model to another subset
# crossValidFunction(fullDT = summaryABSK_AllData, origData = modelData,
#                    statsModel = model8, k = 4, idCol = "pixID")


## 20m res
# model14_allData <- gamlss(SEV_PROP ~ CTI + isForest + Lari + FlamConif * HEIGHT_LOWER +
#                             SMR_ord + #maxDC +
#                             re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                           sigma.formula = ~ Elevation + Decid * UNDERSTOREY +
#                             SMR_ord + #meanDC +
#                             re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                           nu.formula = ~ UNDERSTOREY * FlamConif + UNDERSTOREY * Decid +
#                             SMR_ord + PctSlope + #cvDC +
#                             SurfAspectRatio + SurfReliefRatio +
#                             re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                           tau.formula = ~ FlamConif + Lari + Abie + UNDERSTOREY + HEIGHT_LOWER +
#                             SMR_ord + SurfAspectRatio + #rangeDC +
#                             re(random = ~1 | Czone1/NR1/FIRE_NAME),
#                           family = BEINF,
#                           control = gamlss.control(c.crit = 0.1),
#                           data = na.omit(summaryABSK_AllData[, c("SEV_PROP", "CTI", "isForest", "Lari", "Abie", "HEIGHT_LOWER",
#                                                                  "FlamConif", "SMR_ord", "Elevation", "Decid", "UNDERSTOREY",
#                                                                  "PctSlope", "SurfAspectRatio", "SurfReliefRatio",
#                                                                  "Czone1",  "NR1", "FIRE_NAME"), with = FALSE])) ## need to subset cols for na.omit

# Rsq(model14)
# Rsq(model14_allData)  ## a bit worse with all the data.
# plot(model14_allData)  ## marginally better
# summary(model14_allData)

cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet", "Czone1", "NR1", "FIRE_NAME",
          "UNDERSTOREY", "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio",
          "isForest")
## model doesn't fit with full data
model8FULL <- gamlss(SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                       SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                     sigma.formula = ~ Decid*UNDERSTOREY +
                       SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                     nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                       PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                       SMR_wet + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                     tau.formula = ~ isForest + isForest*UNDERSTOREY +
                       Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                       Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                       CTI + SurfReliefRatio + SMR_wet +
                       re(random = ~1 | Czone1/NR1/FIRE_NAME),
                     family = BEINF, control = gamlss.control(c.crit = 0.1),
                     data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))

## try to remove SAR (hit singularity again) and CTI (converged)
model8.2FULL <- gamlss(SEV_PROP ~ Decid*HEIGHT_LOWER + Lari +
                         SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       sigma.formula = ~ Decid*UNDERSTOREY +
                         SMR_wet + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                         + PctSlope*SurfReliefRatio +
                         SMR_wet + re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                       tau.formula = ~ isForest + isForest*UNDERSTOREY +
                         Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                         Lari*HEIGHT_LOWER + Lari*UNDERSTOREY + SurfReliefRatio +
                         SMR_wet +
                         re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       family = BEINF, control = gamlss.control(c.crit = 0.1),
                       data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))

## try to remove SAR (hit singularity again) and SMR (hit singularity again)
model8.3FULL <- gamlss(SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                         re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       sigma.formula = ~ Decid*UNDERSTOREY +
                         re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                         PctSlope*SurfReliefRatio +
                         re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                       tau.formula = ~ isForest + isForest*UNDERSTOREY +
                         Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                         Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                         CTI + SurfReliefRatio + 
                         re(random = ~1 | Czone1/NR1/FIRE_NAME),
                       family = BEINF, control = gamlss.control(c.crit = 0.1),
                       data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))

## simplify REs, but keep SMR_wet and CTI - now using the datasubset so that we can compare with the more complex model
model8.4 <- gamlss(SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                     SMR_wet + re(random = ~1 | FIRE_NAME),
                   sigma.formula = ~ Decid*UNDERSTOREY +
                     SMR_wet + re(random = ~1 | FIRE_NAME),
                   nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                     PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                     SMR_wet + re(random = ~ 1 | FIRE_NAME),
                   tau.formula = ~ isForest + isForest*UNDERSTOREY +
                     Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                     Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                     CTI + SurfReliefRatio + SMR_wet +
                     re(random = ~1 | FIRE_NAME),
                   family = BEINF, control = gamlss.control(c.crit = 0.1),
                   data = modelData)

model8.4FULL <- gamlss(SEV_PROP ~ Decid*HEIGHT_LOWER + CTI + Lari +
                         SMR_wet + re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~ Decid*UNDERSTOREY +
                         SMR_wet + re(random = ~1 | FIRE_NAME),
                       nu.formula = ~ FlamConif + Decid + UNDERSTOREY +
                         PctSlope*SurfAspectRatio + PctSlope*SurfReliefRatio +
                         SMR_wet + re(random = ~ 1 | FIRE_NAME),
                       tau.formula = ~ isForest + isForest*UNDERSTOREY +
                         Decid*HEIGHT_LOWER + Decid*UNDERSTOREY +
                         Lari*HEIGHT_LOWER + Lari*UNDERSTOREY +
                         CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       family = BEINF, control = gamlss.control(c.crit = 0.1),
                       data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))


GAIC(model8,
     model8.4,  ## almost the same.
     k = log(nrow(modelData)))

Rsq(model8)
Rsq(model8.4) ## also almost the same
Rsq(model8.4FULL)  ## a bit lower with all the data
```

## Model validation
Using k-fold cross validation on `model8.4` (which was fitted on a 1/4 of the data), with the data partioned in 4 chunks (so the size of the validation chunk is the same as the one used to fit the model).
Note that simplifying the random effects makes no differenc, with regards to the more complex nesting.
```{r crossValidation, cache = FALSE, message=FALSE}
crossValidResults <- Cache(crossValidFunction,
                           fullDT = summaryABSK_AllData, 
                           origData = modelData, 
                           statsModel = model8.4,
                           k = 4, idCol = "pixID",
                           cacheObj1 = summaryABSK_AllData$pixID, 
                           cacheObj2 = modelData$pixID,
                           omitArgs = c("fullDT", "origData"), 
                           cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics <- rbindlist(lapply(crossValidResults, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics[, sampID := 1:length(crossValidResults)]

allConfMatrices <- rbindlist(lapply(1:length(crossValidResults), FUN = function(x) {
  DT <- as.data.table(crossValidResults[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

allRsquared <- rbindlist(lapply(crossValidResults, FUN = function(x) as.data.table(t(x$Rsquared))))
allRsquared[, sampID := 1:length(crossValidResults)]

## add identifier
allConfMatrices[, compareID := paste(Prediction, Reference, sep = "_")]

## average confusion matrix 
avgAllConfMatrices <- allConfMatrices[, list(avgN = mean(N)), by = compareID]
avgAllConfMatrices <- unique(allConfMatrices[, .(Prediction, Reference, compareID)])[avgAllConfMatrices, on = "compareID"]
avgAllConfMatrices[, compareID := NULL]
avgAllConfMatrices <- dcast(avgAllConfMatrices, Prediction ~ Reference)

## ------------------------------------------------------------------------------------------------
## !!! RESULTS ARE EXACTLY THE SAME BECAUSE FIRE_NAME IS UNIQUE IN THE NESTING
crossValidResults2 <- Cache(crossValidFunction,
                            fullDT = summaryABSK_AllData, 
                            origData = modelData, 
                            statsModel = model8,
                            k = 4, idCol = "pixID",
                            cacheObj1 = summaryABSK_AllData$pixID, 
                            cacheObj2 = modelData$pixID,
                            omitArgs = c("fullDT", "origData"), 
                            cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics2 <- rbindlist(lapply(crossValidResults2, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics2[, sampID := 1:length(crossValidResults2)]

allConfMatrices2 <- rbindlist(lapply(1:length(crossValidResults2), FUN = function(x) {
  DT <- as.data.table(crossValidResults2[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

## add identifier
allConfMatrices2[, compareID := paste(Prediction, Reference, sep = "_")]

## average and confusion matrix 
avgallConfMatrices2 <- allConfMatrices2[, list(avgN = mean(N)), by = compareID]
avgallConfMatrices2 <- unique(allConfMatrices2[, .(Prediction, Reference, compareID)])[avgallConfMatrices2, on = "compareID"]
avgallConfMatrices2[, compareID := NULL]
avgallConfMatrices2 <- dcast(avgallConfMatrices2, Prediction ~ Reference)


summary(allValidMetrics)
summary(allValidMetrics2)
```

## Model overall performance
Checking the overall performance of the model, when fitted used the full dataset. 
```{r fullModelValidation, cache = FALSE, message=FALSE}
predictionsDT <- predictAll(model8.4FULL,
                            data = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            type = "response", output = "matrix")
predictionsDT <- as.data.table(predictionsDT)
## change name
setnames(predictionsDT, "y", "SEV_PROP")

## add severity classes
cols <- unique(c("pixID", cols))
predictionsDT[, pixID := na.omit(summaryABSK_AllData[, ..cols])$pixID]
cols <- unique(c("pixID", "SEV_CLASS", cols))
predictionsDT <- na.omit(summaryABSK_AllData[, ..cols])[, .(pixID, SEV_CLASS)][predictionsDT, on = "pixID"]

# predict using rBEINF
predictionsDT[, predSEV_PROP := mean(rBEINF(10, mu, sigma, nu, tau)), by = pixID]

## convert to classes, using the quantiles corresponding to the observed class proportions
## accumulate proportions to get probabilities
quantProbs <- cumsum(table(predictionsDT$SEV_CLASS)/nrow(predictionsDT))
classRanges <- c(0, quantile(predictionsDT$predSEV_PROP, probs = quantProbs))
predictionsDT[, predSEV_CLASS := cut(predSEV_PROP, breaks = classRanges,
                                     include.lowest = TRUE, right = FALSE)]  ## classify as with intervals as ],]
predictionsDT[, predSEV_CLASS := as.numeric(predSEV_CLASS)-1]

classes <- as.character(sort(unique(predictionsDT$SEV_CLASS)))
predictionsDT[, `:=`(SEV_CLASS = factor(SEV_CLASS, levels = classes),
                     predSEV_CLASS = factor(predSEV_CLASS, levels = classes))]

## calculate overall statistics
validMetrics <- caret::multiClassSummary(predictionsDT[, list(obs = SEV_CLASS,
                                                              pred = predSEV_CLASS)],
                                         lev = classes)

## calculate confusion matrix
confMatrix <- caret::confusionMatrix(data = predictionsDT$predSEV_CLASS, reference = predictionsDT$SEV_CLASS)

```


## Include neighbours
### Average severity of burnt neighbours
Start from last best model, and find the best buffer sizes to add to each parameter.
```{r NgbSevModels, cache = FALSE}
upperFormula <- as.formula(paste(
  paste(as.character(model8.4$mu.formula)[c(1,3)], collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))

modelNgbSevMu <- add1(model8.4, what = "mu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

modelNgbSev1 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev1$sigma.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevSigma <- add1(modelNgbSev1, what = "sigma",
                         scope = upperFormula,
                         k = log(length(samp)),
                         parallel = "multicore", ncpus = 10)

modelNgbSev2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev2$nu.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevNu <- add1(modelNgbSev2, what = "nu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

modelNgbSev3 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev3$tau.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevTau <- add1(modelNgbSev3, what = "tau",
                       scope = upperFormula,
                       k = log(length(samp)),
                       parallel = "multicore", ncpus = 10)
modelNgbSev4 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                         SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))
```

#### fit model to full dataset:
```{r FullNgbSevModel, cache = FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")

modelNgbSev4FULL <- update(modelNgbSev4, data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))
Rsq(modelNgbSev4FULL)
summary(modelNgbSev4FULL)

## remove some interactions that are not significant:
modelNgbSev4FULL2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                              CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
                              re(random = ~1 | FIRE_NAME),
                            sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
                              re(random = ~1 | FIRE_NAME),
                            nu.formula = ~FlamConif +
                              Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                              PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
                              re(random = ~1 | FIRE_NAME),
                            tau.formula = ~isForest + isForest *
                              UNDERSTOREY + Decid * HEIGHT_LOWER +
                              Lari * HEIGHT_LOWER + CTI +
                              SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
                              re(random = ~1 | FIRE_NAME), family = BEINF,
                            data = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            control = gamlss.control(c.crit = 0.1))

summary(modelNgbSev4FULL2)
```

#### Model validation
Using k-fold cross validation on `modelNgbSev4` (which was fitted on a 1/4 of the data), with the data partioned in 4 chunks (so the size of the validation chunk is the same as the one used to fit the model).
Note that simplifying the random effects makes no difference, with regards to the more complex nesting.
```{r crossValidationNgbSev, message=FALSE}
crossValidResults3 <- Cache(crossValidFunction,
                            fullDT = summaryABSK_AllData,
                            origData = na.omit(modelData),
                            statsModel = modelNgbSev4,
                            k = 4, idCol = "pixID",
                            cacheObj1 = summaryABSK_AllData$pixID,
                            cacheObj2 = na.omit(modelData)$pixID,
                            # useCache = "overwrite",
                            userTags = c("validation", "modelNgbSev4"),
                            omitArgs = c("fullDT", "origData", "userTags"),
                            cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics3 <- rbindlist(lapply(crossValidResults3, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics3[, sampID := 1:length(crossValidResults3)]

allConfMatrices3 <- rbindlist(lapply(1:length(crossValidResults3), FUN = function(x) {
  DT <- as.data.table(crossValidResults3[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

allRsquared3 <- rbindlist(lapply(crossValidResults3, FUN = function(x) as.data.table(t(x$Rsquared))))
allRsquared3[, sampID := 1:length(crossValidResults3)]

## add identifier
allConfMatrices3[, compareID := paste(Prediction, Reference, sep = "_")]

## average confusion matrix 
avgAllConfMatrices3 <- allConfMatrices3[, list(avgN = mean(N)), by = compareID]
avgAllConfMatrices3 <- unique(allConfMatrices3[, .(Prediction, Reference, compareID)])[avgAllConfMatrices3, on = "compareID"]
avgAllConfMatrices3[, compareID := NULL]
avgAllConfMatrices3 <- dcast(avgAllConfMatrices3, Prediction ~ Reference)

## ------------------------------------------------------------------------------------------------
## !!! RESULTS ARE EXACTLY THE SAME BECAUSE FIRE_NAME IS UNIQUE IN THE NESTING
cols <- c("SEV_PROP", "SEV_CLASS", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest", "pixID")

crossValidResults4 <- Cache(crossValidFunction,
                            fullDT = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            origData = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            statsModel = modelNgbSev4FULL2,
                            k = 4, idCol = "pixID",
                            cacheObj1 = na.omit(summaryABSK_AllData[, cols, with = FALSE])[, pixID],
                            cacheObj2 = na.omit(summaryABSK_AllData[, cols, with = FALSE])[, pixID],
                            # useCache = "overwrite",
                            userTags = c("validation", "modelNgbSev4FULL"),
                            omitArgs = c("fullDT", "origData", "userTags"),
                            cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics4 <- rbindlist(lapply(crossValidResults4, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics4[, sampID := 1:length(crossValidResults4)]

allConfMatrices4 <- rbindlist(lapply(1:length(crossValidResults4), FUN = function(x) {
  DT <- as.data.table(crossValidResults4[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

## Metrics by class
allValidMetricsbyClass4 <- rbindlist(lapply(1:length(crossValidResults4), FUN = function(x) {
  DT <- as.data.table(crossValidResults4[[x]]$confMatrix$byClass)
  DT[, sampID := x]
  DT[, class := rownames(crossValidResults4[[x]]$confMatrix$byClass)]
  DT
}))

## add identifier
allConfMatrices4[, compareID := paste(Predictioqn, Reference, sep = "_")]

## average confusion matrix 
avgallConfMatrices4 <- allConfMatrices4[, list(avgN = mean(N)), by = compareID]
avgallConfMatrices4 <- unique(allConfMatrices4[, .(Prediction, Reference, compareID)])[avgallConfMatrices4, on = "compareID"]
avgallConfMatrices4[, compareID := NULL]
avgallConfMatrices4 <- dcast(avgallConfMatrices4, Prediction ~ Reference)


summary(allValidMetrics3)
summary(allValidMetrics4)
cols <- grep("sampID|class", names(allValidMetricsbyClass4), invert = TRUE, value = TRUE)
allValidMetricsbyClass4[, lapply(.SD, mean), .SDcols = cols, by = class]
```

#### Model overall performance
Checking the overall performance of the model, when fitted used the full dataset. 
```{r fullNgbSevModelValidation, message=FALSE}
predictionsDT2 <- predictAll(modelNgbSev4FULL2,
                             data = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                             type = "response", output = "matrix")
predictionsDT2 <- as.data.table(predictionsDT2)
## change name
setnames(predictionsDT2, "y", "SEV_PROP")

## add severity classes
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
cols <- unique(c("pixID", cols))
pixIDs <- na.omit(summaryABSK_AllData[, ..cols])$pixID
predictionsDT2[, pixID := pixIDs]
cols <- unique(c("pixID", "SEV_CLASS", cols))
predictionsDT2 <- na.omit(summaryABSK_AllData[, ..cols])[, .(pixID, SEV_CLASS)][predictionsDT2, on = "pixID"]

## predict using rBEINF
predictionsDT2[, predSEV_PROP := mean(rBEINF(10, mu, sigma, nu, tau)), by = pixID]

## convert to classes, using the quantiles corresponding to the observed class proportions
## accumulate proportions to get probabilities
quantProbs <- cumsum(table(predictionsDT2$SEV_CLASS)/nrow(predictionsDT2))
classRanges <- c(0, quantile(predictionsDT2$predSEV_PROP, probs = quantProbs))
predictionsDT2[, predSEV_CLASS := cut(predSEV_PROP, breaks = classRanges,
                                      include.lowest = TRUE, right = FALSE)]  ## classify as with intervals as ],]
predictionsDT2[, predSEV_CLASS := as.numeric(predSEV_CLASS)-1]

classes <- as.character(sort(unique(predictionsDT2$SEV_CLASS)))
predictionsDT2[, `:=`(SEV_CLASS = factor(SEV_CLASS, levels = classes),
                      predSEV_CLASS = factor(predSEV_CLASS, levels = classes))]

## calculate overall statistics
validMetrics2 <- caret::multiClassSummary(predictionsDT2[, list(obs = SEV_CLASS,
                                                                pred = predSEV_CLASS)],
                                          lev = classes)

## calculate confusion matrix
confMatrix2 <- caret::confusionMatrix(data = predictionsDT2$predSEV_CLASS, reference = predictionsDT2$SEV_CLASS)
```

### Proportion of burnt neighbours
Start from last best model, and find the best buffer sizes to add to each parameter.
```{r burnNgbPropModels, cache = FALSE}
upperFormula <- as.formula(paste(
  paste(as.character(model8.4$mu.formula)[c(1,3)], collapse = " "),
  "+",
  paste(grep("ngbPropBurnsSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))

modelNgbSevMu <- add1(model8.4, what = "mu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

# modelNgbSev1 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
#                          CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        sigma.formula = ~Decid * UNDERSTOREY + SMR_wet +
#                          re(random = ~1 | FIRE_NAME),
#                        nu.formula = ~FlamConif +
#                          Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
#                          PctSlope * SurfReliefRatio + SMR_wet +
#                          re(random = ~1 | FIRE_NAME),
#                        tau.formula = ~isForest + isForest *
#                          UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
#                          Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
#                          re(random = ~1 | FIRE_NAME), family = BEINF,
#                        data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev1$sigma.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevSigma <- add1(modelNgbSev1, what = "sigma",
                         scope = upperFormula,
                         k = log(length(samp)),
                         parallel = "multicore", ncpus = 10)

# modelNgbSev2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
#                          CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
#                          re(random = ~1 | FIRE_NAME),
#                        nu.formula = ~FlamConif +
#                          Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
#                          PctSlope * SurfReliefRatio + SMR_wet +
#                          re(random = ~1 | FIRE_NAME),
#                        tau.formula = ~isForest + isForest *
#                          UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
#                          Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
#                          re(random = ~1 | FIRE_NAME), family = BEINF,
#                        data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev2$nu.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevNu <- add1(modelNgbSev2, what = "nu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

# modelNgbSev3 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
#                          CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
#                          re(random = ~1 | FIRE_NAME),
#                        nu.formula = ~FlamConif +
#                          Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
#                          PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        tau.formula = ~isForest + isForest *
#                          UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
#                          Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
#                          re(random = ~1 | FIRE_NAME), family = BEINF,
#                        data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev3$tau.formula), collapse = " "),
  "+",
  paste(grep("meanngbSEV_PROP_", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevTau <- add1(modelNgbSev3, what = "tau",
                       scope = upperFormula,
                       k = log(length(samp)),
                       parallel = "multicore", ncpus = 10)
# modelNgbSev4 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
#                          CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
#                          re(random = ~1 | FIRE_NAME),
#                        nu.formula = ~FlamConif +
#                          Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
#                          PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME),
#                        tau.formula = ~isForest + isForest *
#                          UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
#                          Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
#                          SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
#                          re(random = ~1 | FIRE_NAME), family = BEINF,
#                        data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))
```

#### fit model to full dataset:
```{r FullburnNgbPropModel, cache = FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")

modelNgbSev4FULL <- update(modelNgbSev4, data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))
Rsq(modelNgbSev4FULL)
summary(modelNgbSev4FULL)

## remove some interactions that are not significant:
# modelNgbSev4FULL2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
#                               CTI + Lari + SMR_wet + meanngbSEV_PROP_30m +
#                               re(random = ~1 | FIRE_NAME),
#                             sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + meanngbSEV_PROP_120m +
#                               re(random = ~1 | FIRE_NAME),
#                             nu.formula = ~FlamConif +
#                               Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
#                               PctSlope * SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
#                               re(random = ~1 | FIRE_NAME),
#                             tau.formula = ~isForest + isForest *
#                               UNDERSTOREY + Decid * HEIGHT_LOWER +
#                               Lari * HEIGHT_LOWER + CTI +
#                               SurfReliefRatio + SMR_wet + meanngbSEV_PROP_30m +
#                               re(random = ~1 | FIRE_NAME), family = BEINF,
#                             data = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
#                             control = gamlss.control(c.crit = 0.1))

summary(modelNgbSev4FULL2)
```

#### Model validation
Using k-fold cross validation on `modelNgbSev4` (which was fitted on a 1/4 of the data), with the data partioned in 4 chunks (so the size of the validation chunk is the same as the one used to fit the model).
Note that simplifying the random effects makes no difference, with regards to the more complex nesting.
```{r crossValidationBurnNgbProp, message=FALSE}
crossValidResults3 <- Cache(crossValidFunction,
                            fullDT = summaryABSK_AllData,
                            origData = na.omit(modelData),
                            statsModel = modelNgbSev4,
                            k = 4, idCol = "pixID",
                            cacheObj1 = summaryABSK_AllData$pixID,
                            cacheObj2 = na.omit(modelData)$pixID,
                            # useCache = "overwrite",
                            userTags = c("validation", "modelNgbSev4"),
                            omitArgs = c("fullDT", "origData", "userTags"),
                            cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics3 <- rbindlist(lapply(crossValidResults3, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics3[, sampID := 1:length(crossValidResults3)]

allConfMatrices3 <- rbindlist(lapply(1:length(crossValidResults3), FUN = function(x) {
  DT <- as.data.table(crossValidResults3[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

allRsquared3 <- rbindlist(lapply(crossValidResults3, FUN = function(x) as.data.table(t(x$Rsquared))))
allRsquared3[, sampID := 1:length(crossValidResults3)]

## add identifier
allConfMatrices3[, compareID := paste(Prediction, Reference, sep = "_")]

## average confusion matrix 
avgAllConfMatrices3 <- allConfMatrices3[, list(avgN = mean(N)), by = compareID]
avgAllConfMatrices3 <- unique(allConfMatrices3[, .(Prediction, Reference, compareID)])[avgAllConfMatrices3, on = "compareID"]
avgAllConfMatrices3[, compareID := NULL]
avgAllConfMatrices3 <- dcast(avgAllConfMatrices3, Prediction ~ Reference)

## ------------------------------------------------------------------------------------------------
## !!! RESULTS ARE EXACTLY THE SAME BECAUSE FIRE_NAME IS UNIQUE IN THE NESTING
cols <- c("SEV_PROP", "SEV_CLASS", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest", "pixID")

crossValidResults4 <- Cache(crossValidFunction,
                            fullDT = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            origData = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            statsModel = modelNgbSev4FULL2,
                            k = 4, idCol = "pixID",
                            cacheObj1 = na.omit(summaryABSK_AllData[, cols, with = FALSE])[, pixID],
                            cacheObj2 = na.omit(summaryABSK_AllData[, cols, with = FALSE])[, pixID],
                            # useCache = "overwrite",
                            userTags = c("validation", "modelNgbSev4FULL"),
                            omitArgs = c("fullDT", "origData", "userTags"),
                            cacheRepo = "analyses/cache")

## collapse results into data.tables
allValidMetrics4 <- rbindlist(lapply(crossValidResults4, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics4[, sampID := 1:length(crossValidResults4)]

allConfMatrices4 <- rbindlist(lapply(1:length(crossValidResults4), FUN = function(x) {
  DT <- as.data.table(crossValidResults4[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))

## Metrics by class
allValidMetricsbyClass4 <- rbindlist(lapply(1:length(crossValidResults4), FUN = function(x) {
  DT <- as.data.table(crossValidResults4[[x]]$confMatrix$byClass)
  DT[, sampID := x]
  DT[, class := rownames(crossValidResults4[[x]]$confMatrix$byClass)]
  DT
}))

## add identifier
allConfMatrices4[, compareID := paste(Predictioqn, Reference, sep = "_")]

## average confusion matrix 
avgallConfMatrices4 <- allConfMatrices4[, list(avgN = mean(N)), by = compareID]
avgallConfMatrices4 <- unique(allConfMatrices4[, .(Prediction, Reference, compareID)])[avgallConfMatrices4, on = "compareID"]
avgallConfMatrices4[, compareID := NULL]
avgallConfMatrices4 <- dcast(avgallConfMatrices4, Prediction ~ Reference)


summary(allValidMetrics3)
summary(allValidMetrics4)
cols <- grep("sampID|class", names(allValidMetricsbyClass4), invert = TRUE, value = TRUE)
allValidMetricsbyClass4[, lapply(.SD, mean), .SDcols = cols, by = class]
```

#### Model overall performance
Checking the overall performance of the model, when fitted used the full dataset. 
```{r fullburnNgbPropModelValidation, message=FALSE}
predictionsDT2 <- predictAll(modelNgbSev4FULL2,
                             data = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                             type = "response", output = "matrix")
predictionsDT2 <- as.data.table(predictionsDT2)
## change name
setnames(predictionsDT2, "y", "SEV_PROP")

## add severity classes
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
cols <- unique(c("pixID", cols))
pixIDs <- na.omit(summaryABSK_AllData[, ..cols])$pixID
predictionsDT2[, pixID := pixIDs]
cols <- unique(c("pixID", "SEV_CLASS", cols))
predictionsDT2 <- na.omit(summaryABSK_AllData[, ..cols])[, .(pixID, SEV_CLASS)][predictionsDT2, on = "pixID"]

## predict using rBEINF
predictionsDT2[, predSEV_PROP := mean(rBEINF(10, mu, sigma, nu, tau)), by = pixID]

## convert to classes, using the quantiles corresponding to the observed class proportions
## accumulate proportions to get probabilities
quantProbs <- cumsum(table(predictionsDT2$SEV_CLASS)/nrow(predictionsDT2))
classRanges <- c(0, quantile(predictionsDT2$predSEV_PROP, probs = quantProbs))
predictionsDT2[, predSEV_CLASS := cut(predSEV_PROP, breaks = classRanges,
                                      include.lowest = TRUE, right = FALSE)]  ## classify as with intervals as ],]
predictionsDT2[, predSEV_CLASS := as.numeric(predSEV_CLASS)-1]

classes <- as.character(sort(unique(predictionsDT2$SEV_CLASS)))
predictionsDT2[, `:=`(SEV_CLASS = factor(SEV_CLASS, levels = classes),
                      predSEV_CLASS = factor(predSEV_CLASS, levels = classes))]

## calculate overall statistics
validMetrics2 <- caret::multiClassSummary(predictionsDT2[, list(obs = SEV_CLASS,
                                                                pred = predSEV_CLASS)],
                                          lev = classes)

## calculate confusion matrix
confMatrix2 <- caret::confusionMatrix(data = predictionsDT2$predSEV_CLASS, reference = predictionsDT2$SEV_CLASS)
```


## Inspecting the final model and figures
Probabilities of 0s and 1s in function of covariates:
xi0 = nu = p0 / (1 + p0 + p1)
xi1 = tau = p1 / (1 + p0 + p1)

Only significant covariates are plotted
```{r modelResults, include = FALSE}
## plot the probabilities at 0 and 1
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "meanngbSEV_PROP_30m", "FIRE_NAME", "UNDERSTOREY", "meanngbSEV_PROP_120m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
plotData <- na.omit(summaryABSK_AllData[, ..cols, with = FALSE])
plotData$p0 <- fitted(modelNgbSev4FULL2, "nu") / (1 + fitted(modelNgbSev4FULL2, "nu") + fitted(modelNgbSev4FULL2, "tau"))
plotData$p1 <- fitted(modelNgbSev4FULL2, "tau") / (1 + fitted(modelNgbSev4FULL2, "nu") + fitted(modelNgbSev4FULL2, "tau"))
```

Plots of probabilities at 0 and 1 vs. covariates
```{r p0p1Plots}
## 0 probs
plot1 <- ggplot(plotData, aes(x = FlamConif, y = p0)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() +
  labs(title = "p0 ~ %FlamConifer\n GAM smoother",
       x = "% cover")

plot2 <- ggplot(plotData, aes(x = Decid, y = p0)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p0 ~ %Deciduous\n GAM smoother",
       x = "% cover")

plot3 <- ggplot(plotData, aes(x = as.integer(UNDERSTOREY), y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "p0 ~ hasUnderstorey\n(GLM logit smoother)",
       x = "understorey")

plot4 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "p0 ~ SoilMoisture\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

plot5 <- ggplot(plotData, 
                aes(x = PctSlope, y = p0,
                    colour = cut(SurfAspectRatio, 
                                 breaks = quantile(plotData$SurfAspectRatio, 
                                                   probs = c(0, 0.10, 0.25, 0.5, 0.75, 0.90, 1))))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p0 ~ Slope * Roughness(SAR)\n GAM smoother",
       colour = "% roughness")

plot6 <- ggplot(plotData, 
                aes(x = meanngbSEV_PROP_30m, y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p0 ~ meanNgbSeverity (30m)\n GAM smoother",
       x = "avg. severity")

ggarrange(plot1,plot2, 
          plot6, plot3, plot4, 
          # plot5, # get_legend(plot1),
          nrow = 3, ncol = 2)

## 1 probs,
plot7 <- ggplot(plotData, aes(x = as.integer(isForest), y = p1,
                              col = UNDERSTOREY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "p1 ~ isForest\n(GLM logit smoother)",
       x = "", y = "understorey")

plot8 <- ggplot(plotData, aes(x = Decid, y = p1,
                              colour = cut(HEIGHT_LOWER, 
                                           breaks = c(0, 5, 10, 20, max(plotData$HEIGHT_LOWER))))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p1 ~ %Deciduous * Height\n GAM smoother",
       x = "% cover", colour = "height")

plot9 <- ggplot(plotData, aes(x = Lari, y = p1,
                              colour = cut(HEIGHT_LOWER, 
                                           breaks = c(0, 5, 10, 20, max(plotData$HEIGHT_LOWER))))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p1 ~ %Larch\n GAM smoother",
       x = "% cover", colour = "height")

plot10 <- ggplot(plotData, aes(x = CTI, y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  theme_pubr() + 
  labs(title = "p1 ~ Wetness\n GAM smoother",
       x = "CTI", y = "mean")

plot11 <- ggplot(plotData, aes(x = SurfReliefRatio, y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p1 ~ Roughness (SRR)\n GAM smoother",
       x = "% roughness")

plot12 <- ggplot(plotData, 
                 aes(x = meanngbSEV_PROP_30m, y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "p1 ~ meanNgbSeverity (30m)\n GAM smoother",
       x = "avg. severity")

plot13 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "p1 ~ SoilMoisture\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

ggarrange(plot7, plot8,
          plot9, plot10, 
          plot11, plot12, 
          plot13,
          nrow = 3, ncol = 3)
```

Plot of the fitted mean vs covariates
```{r fittedMeanPlots}
plotData[, fittedVals := fitted(modelNgbSev4FULL2)]


plot14 <- ggplot(plotData, aes(x = Decid, y = fittedVals,
                               colour = cut(HEIGHT_LOWER, 
                                            breaks = c(0, 5, 10, 20, max(plotData$HEIGHT_LOWER))))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "Severity ~ %Deciduous * Height \n GAM smoother",
       x = "% cover", colour = "height")

plot15 <- ggplot(plotData, aes(x = CTI, y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  theme_pubr() + 
  labs(title = "Severity ~ Wetness\n GAM smoother",
       x = "CTI")

plot16 <- ggplot(plotData, aes(x = Lari, y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "Severity ~ %Larch\n GAM smoother",
       x = "% cover")

plot17 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "Severity ~ SoilMoisture\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

plot18 <- ggplot(plotData, 
                 aes(x = meanngbSEV_PROP_30m, y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "Severity ~ meanNgbSeverity (30m)\n GAM smoother",
       x = "avg. severity")

ggarrange(plot14, plot15,
          plot16,plot17, plot18,
          nrow = 3, ncol = 2, 
          font.label = list(size = 16))

```

Plot of the fitted scale vs covariates
```{r fittedScalePlots}
plotData[, fittedValsScale := fitted(modelNgbSev4FULL2, what = "sigma")]

plot18 <- ggplot(plotData, aes(x = Decid, y = fittedValsScale,
                               col = UNDERSTOREY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "Severity (scale) ~ %Deciduous\n GAM smoother",
       x = "% cover", colour = "understorey")

plot19 <-  ggplot(plotData, aes(x = as.integer(SMR_wet), y = fittedValsScale)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  theme_pubr() + 
  labs(title = "Severity (scale) ~ SoilMoisture\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

plot20 <- ggplot(plotData, 
                 aes(x = meanngbSEV_PROP_30m, y = fittedValsScale)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  theme_pubr() + 
  labs(title = "Severity (scale) ~ meanNgbSeverity (30m)\n GAM smoother",
       x = "avg. severity")

ggarrange(plot18, plot19, plot20,
          nrow = 1, ncol = 3,
          font.label = list(size = 16))

```


```{r saveLoadRdata, include = FALSE, eval = FALSE}
save(list = c("plotData", grep("model", ls(), value = TRUE)),
     # file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_20m_goodSample_Nov1.RData")
     # file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_30m_goodSample_Nov1.RData")
     file = "/media/data/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_30m_goodSample_NgbSevApr21.RData")

## save to GDrive
# library(googledrive)
# save.image(file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/LIM.RData")
# drive_auth(use_oob = TRUE)
# drive_upload("~/Downloads/LIM.RData", path = as_id("1jSrwbU83NVBosWFTxnv2JB2FVkiR5OLD"))

# load("analyses/modelsGAMLSS_0-3Days_goodSample_Nov1.RData")
```

