---
title: "Dave Andison's fire data Severity models"
author: "Ceres Barros"
date: "02/04/2019"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/GitHub/LandscapesInMotion/")
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)

library(data.table)
library(ggplot2)
library(reshape2)
library(sf)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(zoib)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE,
        "reproducible.cachePath" = "analyses/cache")

source("analyses/R_tools/summarizeABSK_AllData.R")
source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R")
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE
```


```{r dataprep, include = FALSE, cache = FALSE}
## source dataprep:
ABSK_AllData <- reproducible::Cache(ABSKfires_DataPrep,
                                    fireDataPath = "data/fires_Dave/fireSev",
                                    vegDataPath = "data/fires_Dave/prefireVeg",
                                    topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
                                    weatherDataPath = "data/fires_Dave/fireWeather",
                                    doCache = doCache, bindAllFires = FALSE, 
                                    cacheRepo = "analyses/cache",
                                    userTags = "ABSKfires_DataPrep")

## exclude event perimeter patch type
memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) all(is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
zeroCols <- sapply(ABSK_AllData, FUN = function(x) all(x == 0, na.rm = TRUE)) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!zeroCols))]

## calculate stand age at fire year by subtracting fire year - makes negative values
ABSK_AllData[!is.na(ORIGIN_UPPER), STAND_AGE := FIRE_YEAR - ORIGIN_UPPER]

## convert SMR to ordinal variable (moister as values increase)
ABSK_AllData[SMR == "D", SMR_ord := 1]
ABSK_AllData[SMR == "F", SMR_ord := 2]
ABSK_AllData[SMR == "M", SMR_ord := 3]
ABSK_AllData[SMR == "W", SMR_ord := 4]
ABSK_AllData[SMR == "A", SMR_ord := 5]

ABSK_AllData[, SMR_ord := as.ordered(SMR_ord)]

## summarise weather data - faster if not cached
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             dim = dim(ABSK_AllData),
                             saveDir = "analyses/fireDataSummary/",
                             overwrite = FALSE, omitArgs = c("DT"))

## collapse species into more generic categories
summaryABSK_AllData[, FlamConif := sum(`Pice glau`, `Pice mari`, `Pice enge`,
                                        `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Abie := sum(`Abie bals`, `Abie lasi`, na.rm = TRUE),
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 
dim(summaryABSK_AllData)

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")
  
collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if(all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]

summaryABSK_AllData[, isForest := 0]
summaryABSK_AllData[is.na(NonForestValue), isForest := 1]

## change lari laris to simpler name
setnames(summaryABSK_AllData, "Lari lari", "Lari")

## change understorey to numeric
summaryABSK_AllData[, UNDERSTOREY := as.numeric(UNDERSTOREY)]

## rescale severity from 0-1
summaryABSK_AllData[, SEV_PROP := SEV_CONT/100]

rm(NAcols, zeroCols, ABSK_AllData)
amc::.gc()
```


Variables selected for models:
* Climate:
  + meanFFMC	-0.16***
  + cvDMC	0.16***
  + cvBUI	0.15***
  + minFFMC	-0.14***
  + rangeFFMC	0.14*** - highly correlated with minFFMC
  + cvFFMC	0.13*** - highly correlated with minFFMC
  + minDMC	-0.13*** - highly correlated with cvDMC
  + minBUI	-0.13*** - - highly correlated with min/cvDMC
* Topography
  + PctSlope	0.08***
  + SurfAspectRatio	0.08***
  + Aspect	0.07***
  + CTI	-0.07*** 
  + SurfReliefRatio	0.04***
  + Elevation	-0.03***
* Vegetation - Spearmand correlations with severity
  + Group flammable conifers 0.10 ***
    + Pinu bank	0.15***
    + Pice mari	0.06***
    + Pice glau	-0.04***
    + Pinu cont	-0.04***
    + Pice enge	0.01***
  + Group Larix
    + Lari lari	-0.05***
  + Group Abies 0.00***
    + Abie lasi	0.00
    + Abie bals	0.01***
  + Group deciduous 0.03***
    + Betu papy	-0.03***
    + Popu trem	0.03***
    + Popu balb	0.02***
 
  + UNDERSTOREY
  + ORIGIN_UPPER 0.10***
* Landcover type
  + isForest
  
```{r severityHist, fig.height=8, fig.show = "hold"}
## subset data
modelData <- summaryABSK_AllData[, .(SEV_PROP, meanFFMC, cvDMC, cvBUI, minFFMC, rangeFFMC, cvFFMC, minDMC, minBUI, FlamConif, Abie, Decid, Lari, UNDERSTOREY, ORIGIN_UPPER, PctSlope, SurfAspectRatio, Aspect, CTI, SurfReliefRatio, Elevation, isForest, FIRE_NAME)]
cols <- grep("FIRE_NAME", names(modelData), invert = TRUE, value = TRUE)
modelData <- modelData[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

# form <- SEV_PROP ~ meanFFMC + cvDMC + cvBUI | 1 | FlamConif + isForest | meanFFMC + cvDMC| 1
# Error in jags.model(op, data = dataIn, n.adapt = 0, inits = inits.internal,  : 
#   Error in node (log(p1[1,1]))
# Invalid parent values

# form <- SEV_PROP ~ meanFFMC + cvDMC + cvBUI | 1 | FlamConif | meanFFMC + cvDMC| 1
# Error in setParameters(init.values[[i]], i) : RUNTIME ERROR:
# Dimension mismatch in values supplied for sigmab0.L1

form <- SEV_PROP ~ meanFFMC + cvDMC + cvBUI | 1 | FlamConif | meanFFMC + cvDMC| 1
form <- Formula(form)

zoib1 <- zoib(model = form, zero.inflation = TRUE, one.inflation = TRUE, random = 1, EUID = modelData$FIRE_NAME, data = na.omit(modelData), joint = FALSE)


```


