---
title: "Dave Andison's fire data Severity models - using Gradient-Boosted Trees"
author: "Ceres Barros"
date: "09/09/2025"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(getwd()))
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)
```

```{r setup2, include=FALSE}
if (!exists("pkgDir")) {
  pkgDir <- file.path(
    if (Sys.info()[["user"]] == "rstudio") "packages_docker" else "packages",
    version$platform,
    paste0(version$major, ".", strsplit(version$minor, "[.]")[[1]][1])
  )
  
  if (!dir.exists(pkgDir)) {
    dir.create(pkgDir, recursive = TRUE)
  }
  .libPaths(pkgDir)
}

install.packages("Require")

Require::Require("xgboost (==3.0.5.1)", repos = "https://dmlc.r-universe.dev") ## not working "SSL connect error", same with install.packages('xgboost', repos = c('https://dmlc.r-universe.dev', 'https://cloud.r-project.org'))
# if (!"xgboost" %in% rownames(installed.packages()) && packageVersion("xgboost") != "3.0.5.1") {
#   install.packages("C:/Users/cbarros/Downloads/xgboost_3.0.5.1.tar.gz")  ## needed to manually download
# }
Require::Install("party")  ## for gpboost

Require::Require(
  c(
    # "caret",
    "crayon",
    "data.table",
    "dplyr",
    "foreign",
    "fasterize",
    "future",
    "future.apply",
    "gamlss",
    "gamlss.inf", 
    "gpboost",
    "CeresBarros/gpboost.utils (>=0.1.0.9001)",
    "ggplot2",
    "ggpubr",
    "ggrepel",
    "Hmisc",
    "LandR.CS",
    "lightgbm",
    "prioritizr",
    "purrr",
    "pROC",
    "raster",
    "CeresBarros/reproducible@devCeres",
    "reshape2",
    "sf",
    "shapr",  
    "PredictiveEcology/SHAPforxgboost (>=0.1.3.9001)",
    "terra",
    "tidyterra",
    "CeresBarros/ToolsCB (>=0.0.9001)",
    # "xgboost (>=3.0.5.1)",  ## above
    "xlsx"
  )
)

devtools::load_all("C:/Users/cbarros/GitHub/caret/pkg/caret/")

options("reproducible.cachePath" = "analyses/cache",
        "reproducible.useDBI" = TRUE,
        "future.globals.maxSize" = +Inf)

source("R/R_tools/Useful_functions.R") ## may replace some functions in ToolsCB, this is ok
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("analyses/R_tools/summarizeABSK_AllData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE

figDir <- "analyses/Figs"
dir.create(figDir, showWarnings = FALSE)
```

```{r dataprep, include = FALSE, cache = FALSE}
resolution = 30
summaryABSK_AllData <- Cache(
  dataPrepWrapper,
  resolution = resolution, 
  fireDataPath = "data/fires_Dave/fireSev",
  vegDataPath = "data/fires_Dave/prefireVeg",
  topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
  weatherDataPath = "data/fires_Dave/fireWeather",
  doCache = doCache,
  useCache = doCache,
  userTags = "summaryABSK_AllData",
  cacheId = "0a70f8b25dc39895",   ## overwride. oct 22 2025, Cache is being silly about "new" userTags
  omitArgs = c("useCache", "doCache", "userTags"))
```

```{r addNgbNoBurns, cache = FALSE, include = FALSE}
## make different buffer sizes around points
crsProj <- crs(vect("data/fires_Dave/fireSev/albertafires1_postfire.shp"), proj = TRUE)

SEV_PROPpoints <- summaryABSK_AllData[, .(pixID, FIRE_NAME, SEV_PROP, Lat, Long)]
SEV_PROPpoints <- st_as_sf(SEV_PROPpoints, coords = c("Long", "Lat"),
                           agr = "constant", crs = st_crs(crsProj))

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i - 1] * 2
}

ngbNoBurns <- reproducible::Cache(calculateNgbBurnsWrapper,
                                  dists = bufferDists,
                                  sevPoints = SEV_PROPpoints, 
                                  sevColID = "SEV_PROP",
                                  fireColID = "FIRE_NAME",
                                  resolution = resolution,
                                  parallel = TRUE,
                                  cores = 3,
                                  useCache = doCache,
                                  cacheId = "9a366f60217632b1",
                                  userTags = "ngbBurns",
                                  omitArgs = c("useCache", "userTags", "cores")) 
ngbNoBurns <- dcast.data.table(ngbNoBurns, ... ~ bufferSize, value.var = "ngbPropBurns")

cols <- grep("pixID|FIRE_NAME", names(ngbNoBurns), invert = TRUE, value = TRUE)
setnames(ngbNoBurns, old = cols, new = paste0("buffer_", cols, "m"))

summaryABSK_AllData <- ngbNoBurns[summaryABSK_AllData, on = .(pixID, FIRE_NAME)]
```

## Maps of fire perimeters, ecozones and nat. regions
```{r}
firesABSK <- Cache(cleanAndBindFireData,
                   files = c("albertafires1_postfire", "albertafires2_postfire", "saskatchewanfires_postfire"),
                   fireDataPath = "data/fires_Dave/fireSev",
                   cachePath = "analyses/cache",
                   userTags = "allFireData",
                   cacheId = "allFireData",
                   omitArgs = c("cachePath", "userTags"))
firesABSK <- vect(firesABSK)

NRAlberta <- prepInputs(targetFile = "AB_Natural_Sub_Regions.shp",
                        archive = "AB_Natural_Sub_Regions.zip",
                        alsoExtract = "similar",
                        url = "https://drive.google.com/file/d/1hW6zy0CpUBdk-K2IAjzW4INjVl1J4aLJ",
                        fun = "sf::st_read",
                        destinationPath = "data/maps",
                        cachePath = "analyses/cache",
                        userTags = "NRAlberta")

NRAlberta <- Cache(prepInputs,
                   targetFile = "Natural_Regions_Subregions_of_Alberta.shp",
                   archive = asPath("natural_regions_subregions_of_alberta.zip"),
                   url = "https://www.albertaparks.ca/media/429607/natural_regions_subregions_of_alberta.zip",
                   alsoExtract = "similar",
                   fun = "sf::st_read",
                   destinationPath = "R/SpaDES/inputs/",
                   cachePath = "R/SpaDES/data/cache",
                   userTags = c("prepInputsNatSubRegionsAB_SA"),
                   cacheId = "652465e92ab1f7e2",
                   omitArgs = c("cachePath", "userTags", "destinationPath", "alsoExtract"))

NRAlberta <- vect(NRAlberta)

## dissolve by region
NRAlberta <- aggregate(NRAlberta[, "NRNAME"], by = "NRNAME", dissolve = TRUE)
NRAlberta <- project(NRAlberta, crs(firesABSK, proj = TRUE))

EcozonesCan <- prepInputs(targetFile = "ecozones.shp",
                          url = "https://sis.agr.gc.ca/cansis/nsdb/ecostrat/zone/ecozone_shp.zip",
                          archive = "ecozone_shp.zip",
                          fun = "sf::st_read",
                          destinationPath = "data/maps",
                          cachePath = "analyses/cache",
                          userTags = "EcozonesCan")
EcozonesCan <- vect(EcozonesCan)
EcozonesCan[EcozonesCan$ZONE_NAME == "Boreal PLain"]$ZONE_NAME <- "Boreal Plain"
EcozonesCan <- project(EcozonesCan, crs(firesABSK, proj = TRUE))

CA_admin <- vect("data/CA_admin/gpr_000a11a_e.shp")
CA_admin <- project(CA_admin, crs(firesABSK, proj = TRUE))

## crop layers to a slightly larger area
extentFires <- ext(rbind(firesABSK, NRAlberta))
extentFires <- extend(extentFires, 50000)
# NRAlberta <- crop(NRAlberta, extentFires)
EcozonesCan <- crop(EcozonesCan, extentFires)
CA_admin <- crop(CA_admin, extentFires)


## TODO: move legend to outside of plot
if (FALSE) {
  png(file.path(figDir, "firesEcozonesNatRAB.png"), height = 7, width = 8,
      units = "in", res = 300)
  plot(EcozonesCan, y = "ZONE_NAME", col = RColorBrewer::brewer.pal(name = "Pastel2", n = 7),
       border = NA, plg = list(x = "topright", inset = c(-0.255, 0)),
       mar = c(3.1, 3.1, 2.1, 7.5), main = "")
  plot(CA_admin, legend = FALSE, add = TRUE)
  plot(NRAlberta, y = "NRNAME", col = RColorBrewer::brewer.pal(name = "Set1", n = 5),
       border = RColorBrewer::brewer.pal(name = "Set1", n = 6), lwd = 2,
       alpha = 0.3, plg = list(x = "bottomright", inset = c(-0.24, 0)),
       mar = c(3.1, 3.1, 2.1, 7.5), add = TRUE)
  text(CA_admin[CA_admin$PRENAME %in% c("Alberta", "Saskatchewan")],
       labels = "PRENAME", adj = c(0.5, 6))
  plot(firesABSK[firesABSK$FIRE_NAME %in% summaryABSK_AllData$FIRE_NAME], legend = FALSE, add = TRUE)
  dev.off()
}
```


## SUMMARY OF PRELIMINARY ANALYSES

Variables selected for models (when using all fire weather days):
* Climate:
+ meanFFMC	-0.16***
+ cvDMC	0.16***
+ cvBUI	0.15***
+ minFFMC	-0.14***
+ rangeFFMC	0.14*** - highly correlated with minFFMC
+ cvFFMC	0.13*** - highly correlated with minFFMC
+ minDMC	-0.13*** - highly correlated with cvDMC
+ minBUI	-0.13*** - - highly correlated with min/cvDMC

* Topography
+ PctSlope	0.08***
+ SurfAspectRatio	0.08*** - **SHOULD BE CALLED SURFACE AREA RATIO** - 
+ Aspect	0.07***
+ CTI	-0.07*** ([Compound Topographic Index (CTI) (Moore et al. 1991) or "soil wetness" transformation can be used to models aspects of hydrologic systems. CTI is strongly correlated with soil moisture](https://wikispaces.psu.edu/display/AnthSpace/Compound+Topographic+Index))
+ SurfReliefRatio	0.04*** ([Describes rugosity. Ratio between relief and length in a particular window](https://en.wikipedia.org/wiki/Relief_ratio))
+ Elevation	-0.03***

* Vegetation - Spearmand correlations with severity
+ Group flammable conifers 0.10 ***
+ Pinu bank	0.15***
+ Pice mari	0.06***
+ Pice glau	-0.04***
+ Pinu cont	-0.04***
+ Pice enge	0.01***
+ Group Larix
+ Lari lari	-0.05***
+ Group Abies 0.00***
+ Abie lasi	0.00
+ Abie bals	0.01***
+ Group deciduous 0.03***
+ Betu papy	-0.03***
+ Popu trem	0.03***
+ Popu balb	0.02***

+ UNDERSTOREY
+ ORIGIN_UPPER 0.10***

* Landcover type
+ isForest

## Gradient boosted trees with XGBoost

### Fit and test model with k-fold cross-validation

*Note: we don't need to scale data for XGBoost*
See [here](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.1365-2656.2008.01390.x)

**Oct 30th 2025**: I tried using a Tweedie distribution. Residuals were just as bad 
and there were actually decreases in model fit.
```{r}
## variables for analysis when considering all weather (first 3 days) - dropping range
## due to lack of data in some fires, dropping:
## crown closure, SMR (which has a proxy- CTI), cvRain, ORIGIN_*
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "Czone1|NR1|FIRE_NAME|pixID|SMR_ord|SMR_wet|buffer", sep = "|"), 
             names(summaryABSK_AllData),
             value = TRUE)
cols <- grep("^range", cols, invert = TRUE, value = TRUE)

## TODO: WE MAY NOT NEED TO SCALE DATA for XGBoost
## https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.1365-2656.2008.01390.x
# modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
# cols2scale <- setdiff(cols, 
#                       c("pixID", "FIRE_NAME", "Czone1", "NR1", "SMR_ord", 
#                         "UNDERSTOREY", "isForest", "SEV_PROP", "SMR_wet"))
# modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]
```


```{r}
cols <- setdiff(cols, c("FIRE_NAME", "Czone1", "NR1"))
digestOfData <- .robustDigest(summaryABSK_AllData[, ..cols])  ## TODO: clear cache before final runs.

## drop FIRENAME for now, as I'm not sure how to include it as a random effesct
runXGBOOST(summaryABSK_AllData[, ..cols], digestOfData, nFolds = 5)
```



Compare with zero-one-inflated beta regression
```{r}
## these were the covariates
# cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
#           "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
#           "FlamConif", "SurfReliefRatio", "isForest")
# modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
# cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
#                 "buffer_1920m", "buffer_30m", "FlamConif",
#                 "SurfReliefRatio")
# modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]
modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")


```

