---
title: "Dave Andison's fire data Severity models"
author: "Ceres Barros"
date: "02/04/2019"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/GitHub/LandscapesInMotion/")
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)

library(data.table)
library(ggplot2)
library(reshape2)
library(sf)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(gamlss.inf)
library(gamlss)
library(ggpubr)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE,
        "reproducible.cachePath" = "analyses/cache")

source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R")
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("R/R_tools/summary.gamlssinf0to1_FIXED.R")
source("R/R_tools/Rsq_FIXED.R")
source("analyses/R_tools/summarizeABSK_AllData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE
```


```{r dataprep, include = FALSE, cache = FALSE}
## source dataprep:
ABSK_AllData <- reproducible::Cache(ABSKfires_DataPrep,
                                    fireDataPath = "data/fires_Dave/fireSev",
                                    vegDataPath = "data/fires_Dave/prefireVeg",
                                    topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
                                    weatherDataPath = "data/fires_Dave/fireWeather",
                                    doCache = doCache, bindAllFires = FALSE, 
                                    cacheRepo = "analyses/cache",
                                    userTags = "ABSKfires_DataPrep")

## exclude event perimeter patch type
memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) all(is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
zeroCols <- sapply(ABSK_AllData, FUN = function(x) all(x == 0, na.rm = TRUE)) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!zeroCols))]

## calculate stand age at fire year by subtracting fire year - makes negative values
ABSK_AllData[!is.na(ORIGIN_UPPER), STAND_AGE := FIRE_YEAR - ORIGIN_UPPER]

## convert SMR to ordinal variable (moister as values increase)
ABSK_AllData[SMR == "D", SMR_ord := 1]
ABSK_AllData[SMR == "F", SMR_ord := 2]
ABSK_AllData[SMR == "M", SMR_ord := 3]
ABSK_AllData[SMR == "W", SMR_ord := 4]
ABSK_AllData[SMR == "A", SMR_ord := 5]

ABSK_AllData[, SMR_ord := as.ordered(SMR_ord)]

## summarise weather data - faster if not cached
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             dim = dim(ABSK_AllData),
                             days = 1:3,
                             saveDir = "analyses/fireDataSummary/",
                             overwrite = FALSE, omitArgs = c("DT"))

## collapse species into more generic categories
summaryABSK_AllData[, FlamConif := sum(`Pice glau`, `Pice mari`, `Pice enge`,
                                       `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Abie := sum(`Abie bals`, `Abie lasi`, na.rm = TRUE),
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")

collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if(all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]

summaryABSK_AllData[, isForest := 0]
summaryABSK_AllData[NonForestValue == "NA", isForest := 1]

## change Lari lari to simpler name
setnames(summaryABSK_AllData, "Lari lari", "Lari")

## rescale severity from 0-1
summaryABSK_AllData[, SEV_PROP := SEV_CONT/100]
## convert the highest fire severity to 1
summaryABSK_AllData[SEV_PROP == 0.9750, SEV_PROP := 1.0]

## change some columns to numeric
cols <- grep("mean|max|min|cv|range|pixID|P_ID|Duration|ORIGIN|HEIGHT|isForest|SEV_PROP|Aspect|^CTI$|PctSlope|SurfReliefRatio|SurfAspectRatio|Elevation", names(summaryABSK_AllData), value = TRUE)
str(summaryABSK_AllData[, ..cols])
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## make some columns factors
cols <- c("FIRE_NAME", "Czone1", "NR1")
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.factor), .SDcols = cols]

rm(NAcols, zeroCols, ABSK_AllData, cols)
amc::.gc()
```


Variables selected for models (when using all fire weather days):
* Climate:
+ meanFFMC	-0.16***
+ cvDMC	0.16***
+ cvBUI	0.15***
+ minFFMC	-0.14***
+ rangeFFMC	0.14*** - highly correlated with minFFMC
+ cvFFMC	0.13*** - highly correlated with minFFMC
+ minDMC	-0.13*** - highly correlated with cvDMC
+ minBUI	-0.13*** - - highly correlated with min/cvDMC
* Topography
+ PctSlope	0.08***
+ SurfAspectRatio	0.08*** - Ratio of the centerline-to-centerline spacing between dimples to the height of the dimple (https://arxiv.org/ftp/arxiv/papers/1305/1305.4705.pdf) / Surface Relief Ratio - describes rugosity in an continuous raster surface within a specified window (https://evansmurphy.wixsite.com/evansspatial/arcgis-gradient-metrics-toolbox)
+ Aspect	0.07***
+ CTI	-0.07*** ([Compound Topographic Index (CTI) (Moore et al. 1991) or "soil wetness" transformation can be used to models aspects of hydrologic systems. CTI is strongly correlated with soil moisture](https://wikispaces.psu.edu/display/AnthSpace/Compound+Topographic+Index))
+ SurfReliefRatio	0.04*** ([Describes rugosity. Ratio between relief and length in a particular window](https://en.wikipedia.org/wiki/Relief_ratio))
+ Elevation	-0.03***
* Vegetation - Spearmand correlations with severity
+ Group flammable conifers 0.10 ***
+ Pinu bank	0.15***
+ Pice mari	0.06***
+ Pice glau	-0.04***
+ Pinu cont	-0.04***
+ Pice enge	0.01***
+ Group Larix
+ Lari lari	-0.05***
+ Group Abies 0.00***
+ Abie lasi	0.00
+ Abie bals	0.01***
+ Group deciduous 0.03***
+ Betu papy	-0.03***
+ Popu trem	0.03***
+ Popu balb	0.02***

+ UNDERSTOREY
+ ORIGIN_UPPER 0.10***
* Landcover type
+ isForest

```{r modelData, include = FALSE}
## variables for analysis when considering all weather (first 3 days) - dropping range
## due to lack of data in some fires, dropping:
## crown closure, SMR (which has a proxy- CTI), cvRain, ORIGIN_*
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "Czone1|NR1|FIRE_NAME|pixID", sep = "|"), names(summaryABSK_AllData),
             value = TRUE)
cols <- cols[!grepl("cvRain", cols)]

## sample a forth of the pixels in each fire
sampDT <- na.omit(summaryABSK_AllData[, ..cols])[,.(FIRE_NAME, pixID)]
sampDT[, noPixels :=  length(pixID), by = FIRE_NAME]
set.seed(123)
samp <- sampDT[, sample(pixID, size = noPixels/4, replace = FALSE), by = FIRE_NAME]$V1
rm(sampDT)

modelData <- na.omit(summaryABSK_AllData[, ..cols])[pixID %in% samp]

## check we didn't lose fires
if (length(setdiff(unique(summaryABSK_AllData$FIRE_NAME), 
                   unique(modelData$FIRE_NAME))))
  stop("Fires lost in sampling!")
```

## GAMLSS models
Selection of variables going "in".
Start with a high covergence criterion to speed up model fitting for now
```{r gamlssModels, results = "asis"}
modelNULL <- gamlss(SEV_PROP ~ 1 + re(random = ~ 1|Czone1/NR1/FIRE_NAME),
                    family = BEINF, control = gamlss.control(c.crit = 0.1),
                    data = modelData)

model1 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model2 <- stepGAIC(model1,
#                    what = "sigma",
#                    scope = list(lower = ~re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                                 upper = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC + 
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI + 
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME)),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 20,
#                    direction = "both", k = log(length(samp)))

model2 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model3 <- stepGAIC(model2,
#                    what = "nu",
#                    scope = list(lower = ~re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                                 upper = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC + 
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI + 
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME)),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 20,
#                    direction = "both", k = log(length(samp)))

model3 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ Decid + FlamConif + SurfAspectRatio + UNDERSTOREY + SurfReliefRatio + PctSlope +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

amc::.gc()
# model4 <- stepGAIC(model3,
#                    what = "tau",
#                    scope = ~ meanFFMC + cvFFMC + meanDMC + cvDMC + meanDC + cvDC + 
#                                   meanISI + cvISI + meanBUI + cvBUI + meanFWI + cvFWI + 
#                                   Aspect + CTI + PctSlope + Elevation + SurfAspectRatio + SurfReliefRatio +
#                                   Lari + FlamConif + Abie + Decid +
#                                   UNDERSTOREY + HEIGHT_LOWER + isForest +
#                                   re(random = ~ 1|Czone1/NR1/FIRE_NAME),
#                    control = gamlss.control(c.crit = 0.1),
#                    parallel = "multicore", ncpus = 18,
#                    direction = "both", k = log(length(samp)))

model4 <- gamlss(formula = SEV_PROP ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Elevation + Decid + UNDERSTOREY + re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ Decid + FlamConif + SurfAspectRatio + UNDERSTOREY + SurfReliefRatio + PctSlope +
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ FlamConif + Lari + HEIGHT_LOWER + UNDERSTOREY + CTI + Abie + SurfAspectRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)

GAIC(modelNULL,
     model1,
     model2,
     model3,  
     model4,  ## best
     k = log(nrow(modelData)))

plot(modelNULL)   ## pretty horrible (with bad sampling) | pretty good  (good sampling and more REs)
plot(model4)   ## pretty good!!!!  (with bad sampling) | looks even better (good sampling and more REs)
```


Let's try with the GAMLSS INF (probabilities at 0s and 1s have more direct interpretation), starting from the previous best model
Note that \code{gamlssInf0to1} doesn't seem to like \code{re} in this case.
`nu` and `tau` are replaced by `xi0` and `xi1`, respectively, which in the case of 0 and 1 inflation are the same between `gamlss` and `gamlssInf0to1`. GAMLSS Inf fits a log-linear multinomial model to the 0s 1s and everything in between  (0 becomes 1, 1 becomes 2 and everything else becomes 3).
Note that when the sampling was corrected and more random effects added, using gamlss.inf offered no improvement
```{r gamlssinfModels, results = "asis"}
modelInf <- gamlssInf0to1(y = SEV_PROP,
                          mu.formula = ~ CTI + isForest + Lari + HEIGHT_LOWER + FlamConif + 
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          sigma.formula = ~ Elevation + Decid + UNDERSTOREY + random(Czone1) + random(NR1) + random(FIRE_NAME),
                          xi0.formula = ~ Decid + FlamConif + SurfAspectRatio + UNDERSTOREY + SurfReliefRatio + PctSlope +
                            random(Czone1) + random(NR1) + random(FIRE_NAME),
                          xi1.formula = ~ FlamConif + Lari + HEIGHT_LOWER + UNDERSTOREY + CTI + Abie + SurfAspectRatio +
                            random(Czone1) + random(NR1) + random(FIRE_NAME), 
                          control = gamlss.control(c.crit = 0.1),
                          data = modelData)

## inspect p-values and exclude non-significant effects/add interactions
summary.gamlssinf0to1_2(modelInf)
plot(modelInf)

GAIC(modelNULL,
     model4,  
     modelInf)  ## this is actually worse.
```

Go back to "standard" GAMLSS, since GAMLSS Inf did not improve models. The advantage of using GAMLSS is that more methods are available. Also, because we are using a beta inflates distribution there's not much gain in using `gamlssInf0to1`.
We can use re() again.

The coefficients are pratically the same (between GAMLSS and GAMLSSInf)

We can get the fitted probabilities at 0 (p0) and 1 (p1) as by reversing the following equations:
xi0 = nu = p0 / (1 - p0 - p1)
xi1 = tau = p1 / (1 - p0 - p1)

hence,
p0 = xi0 / (1 + xi0 + xi1) = nu / (1 + nu + tau)
p1 = xi1 / (1 + xi0 + xi1) = tau / (1 + nu + tau)

Start by adding interactions within covariate groups (vegetation, topography, climate)

```{r addInteractions}
model5 <- gamlss(SEV_PROP ~ CTI + isForest + Lari*HEIGHT_LOWER + FlamConif*HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Elevation + Decid*UNDERSTOREY + 
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
                   PctSlope + SurfReliefRatio + SurfAspectRatio + 
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ FlamConif + Lari + Abie +
                   HEIGHT_LOWER + UNDERSTOREY + CTI*SurfAspectRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)
summary(model5)

## remove non-significant main effects and interactions, and add others 
## tried adding other interactions to mu, but it creates a singular model
## tried adding other interactions to sigma, but it creates a singular model
## tried adding other interactions to nu, but it creates a singular model
## tried adding other interactions to tau, but it creates a singular model
model6 <- gamlss(SEV_PROP ~ CTI + isForest + Lari*HEIGHT_LOWER + FlamConif*HEIGHT_LOWER +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 sigma.formula = ~ Elevation + Decid*UNDERSTOREY +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 nu.formula = ~ UNDERSTOREY*FlamConif + UNDERSTOREY*Decid + 
                   PctSlope + SurfAspectRatio + SurfReliefRatio + 
                   re(random = ~ 1 | Czone1/NR1/FIRE_NAME),
                 tau.formula = ~ FlamConif + Lari + Abie +
                   UNDERSTOREY + HEIGHT_LOWER + SurfAspectRatio +
                   re(random = ~1 | Czone1/NR1/FIRE_NAME),
                 family = BEINF, control = gamlss.control(c.crit = 0.1),
                 data = modelData)
summary(model6)

GAIC(model4,
     modelInf3,
     model5, 
     model6)   ## worse, but I think its worth it for parsimony's sake
Rsq(model5)
Rsq(model6)
```


##### Inspecting final model
Probabilities of 0s and 1s in function of covariates:
xi0 = nu = p0 / (1 - p0 - p1)
xi1 = tau = p1 / (1 - p0 - p1)

Only significant covariates are plotted
```{r back2gamlss}
Rsq(model6)

## plot the probabilities at 0 and 1
modelData$p0 <- fitted(model6, "nu") / (1 + fitted(model6, "nu") + fitted(model6, "tau"))
modelData$p1 <- fitted(model6, "tau") / (1 + fitted(model6, "nu") + fitted(model6, "tau"))

## 0 probs
plot1 <- ggplot(modelData, aes(x = as.integer(UNDERSTOREY), y = p0)) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  labs(title = "p0 ~ presence of understory\n(GLM logit smoother)")

plot2 <- ggplot(modelData, aes(x = FlamConif, y = p0)) +
  geom_smooth() +
  labs(title = "p0 ~ Flamm. conifers % cover\n GAM smoother")

plot3 <- ggplot(modelData, aes(x = Decid, y = p0)) +
  geom_smooth() +
  labs(title = "p0 ~ Deciduous % cover\n GAM smoother")

plot4 <- ggplot(modelData, aes(x = PctSlope, y = p0)) +
  geom_smooth() +
  labs(title = "p0 ~ % slope\n GAM smoother")

plot5 <- ggplot(modelData, aes(x = SurfAspectRatio, y = p0)) +
  geom_smooth() +
  labs(title = "p0 ~ Roughness (SAR)\n GAM smoother")

plot6 <- ggplot(modelData, aes(x = SurfReliefRatio, y = p0)) +
  geom_smooth() +
  labs(title = "p0 ~ Roughness (SRR)\n GAM smoother")

ggarrange(plot1, plot2,
          plot3, plot4, 
          plot5, plot6,
          nrow = 3, ncol = 2)


## 1 probs
plot7 <- ggplot(modelData, aes(x = as.integer(UNDERSTOREY), y = p1)) +
 geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  labs(title = "p1 ~ presence of understory\n(GLM logit smoother)")

plot8 <- ggplot(modelData, aes(x = FlamConif, y = p1)) +
 geom_smooth() +
  labs(title = "p1 ~ Flamm. conifers % cover\n GAM smoother")

plot9 <- ggplot(modelData, aes(x = Lari, y = p1)) +
 geom_smooth() +
  labs(title = "p1 ~ Larch % cover\n GAM smoother")

plot10 <- ggplot(modelData, aes(x = Abie, y = p1)) +
 geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  labs(title = "p1 ~ Fir % cover\n(GLM logit smoother)")

plot11 <- ggplot(modelData, aes(x = SurfAspectRatio, y = p1)) +
 geom_smooth() +
  labs(title = "p1 ~ Roughness (SAR)\n GAM smoother")

plot12 <- ggplot(modelData, aes(x = HEIGHT_LOWER, y = p1)) +
 geom_smooth() +
  labs(title = "p1 ~ Stand height\n GAM smoother")

ggarrange(plot7, plot8,
          plot9, plot10, 
          plot11, plot12,
          nrow = 3, ncol = 2)

```

```{r saveRdata}
save(list = grep("model", ls(), value = TRUE), file = "modelsGAMLSS_0-3Days_goodSample_Nov1.RData")
```



