---
title: "Dave Andison's fire data preliminary analyses"
author: "Ceres Barros"
date: "March 28, 2018"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
## using - as of January 14th, 2019
# loading reproducible     0.2.5.9000
# loading quickPlot        0.1.6
# loading SpaDES.core      0.2.3.9003
# loading SpaDES.tools     0.3.0.9000
# loading SpaDES.addins    0.1.1
# devtools::install_github("PredictiveEcology/quickPlot@development")
# devtools::install_github("PredictiveEcology/reproducible@development")
# devtools::install_github("PredictiveEcology/SpaDES.core@development")
# devtools::install_github("PredictiveEcology/SpaDES.tools@development")


knitr::opts_knit$set(root.dir = "E:/GitHub/LandscapesInMotion/")
library(factoextra)
library(reshape2)
library(vegan)
library(sf)
library(ggplot2)
library(data.table)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(velox)
library(reproducible)
library(quickPlot)

options('reproducible.useNewDigestAlgorithm' = FALSE) 
```


```{r dataprep, include = FALSE}
## source dataprep:
doCache <- TRUE ## control all caching
bindAllFires <- FALSE ##control whether all fire data binding should be repeated
source("analyses/DAfires_expAnalyses_dataPrep.R", echo = FALSE)

## exclude event perimeter patch type
memory.limit(size = 64000)
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) !any(!is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]

rm(NAcols)
amc::.gc()
```

## NOTES - meeting with Dave Dec 6th
Carefull with no. of variables and sample size
Weather variables: how to look at it? Weather at the begining, middle, end? Only averages like Colin did? Averages, variation?

## COMMUNITY COMPOSITION/STRUCTURE PCAs

### PCA of dominant species percentage cover and height across all fire event patch types. 
Blue arrow is severity in a continuous scale, fitted to the ordination *a posteriori*, using `env.fit`.
Severity is higher in "plots" dominated by White spruce (*Picea glauca*), Engelman's spruce (*P. engelmannii*), Aspen (*Populus tremuloides*) and where trees are taller.
It seems that canopy and understory levels usually highly correlated within species.

Also the points pattern doesn't seem related to fire identity, but to levels of crown closure.

```{r PCA, echo = FALSE, fig.height=8, warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR|LAYER$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1+LAYER, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])

## PCA COLOURED BY FIRE_NAME
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by fire name")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$FIRE_NAME))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```

### PCA with canopy and understory levels collapsed for tree species (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but clearer.

```{r PCA2, echo = FALSE, fig.height=8, warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

# ## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```

### PCA with canopy and understory levels collapsed for tree species, but keeping dominant and subdominant spp separate (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but we now see some vegetation associations. Unlike before, crown closure now appears more correlated with severity. It seems higher severity levels coincide with forests where Aspen is the dominant species and White spruce is the subdominant. 

```{r PCA3, echo = FALSE, fig.height=8, warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2|3)$|SPEC(1|2|3)_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

spec3 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC3, value.var = "SPEC3_PER", fun.aggregate = sum, na.rm = TRUE)
setnames(spec3, sppNames, paste0(sppNames, "_3"), skip_absent = TRUE)

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
commCompStruc <- merge(commCompStruc, spec3, by = cols)
rm(spec1, spec2, spec3); amc::.gc()

commCompStruc$`NA` <- NULL

## check if any spp has >100 cover
# cols <- grep("_[[:digit:]]$", names(commCompStruc))
# summary(commCompStruc[, ..cols])

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")

```


### Remove third subdominant species, add age
Correlations between third most dominant species and axes are quite low
Results changed quite a bit, with Lodgepole pine and crown closure appearing both more related to severity, than before (and Aspen and White spruce losing importance).

```{r PCA4, echo = FALSE, fig.height=8, warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|ORIGIN_|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|ORIGIN|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

commCompStruc$`NA` <- NULL

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

### Remove origin - correlations with axes are quite low

```{r PCA5, echo = FALSE, fig.height=8, warning=FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

commCompStruc$`NA` <- NULL

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

Consider pure stands only collapsing canopy and understory

```{r PCA6, echo = FALSE, fig.height=8, warning=FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[SPEC1_PER == 100, ..cols]

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch|SPEC1", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## SPEC1 needs to be a factor for dudi
commCompStruc <- commCompStruc[, SPEC1 := as.factor(SPEC1)]

## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompDudi <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB],
                              scannf = FALSE, nf = 4)
env.fit <- envfit(vegCompDudi, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

dev()
PCAplot <- ordiplot(vegCompDudi, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```


## SIMPLIFY ANALYSIS/DATA 
- Collapse climate data using summary statistics
- Convert species data to long format

```{r summaryData, eval=FALSE, include=FALSE}
## clean-up and free memory
rm(vegCompPCA, env.fit, commCompStruc); amc::.gc()

## sumamrise weather data
summaryABSK_AllData <- ABSK_AllData %>%
  group_by(pixID) %>%
  summarise(startJulDay = min(julDay),
            meanTemp = mean(temp),
            minTemp = min(temp),
            maxTemp = max(temp),
            cvTemp = sd(temp)/mean(temp),
            rangeRH = max(rh) - min(rh),
            meanRH = mean(rh),
            minRH = min(rh),
            maxRH = max(rh),
            cvRH = sd(rh)/mean(rh),
            rangeRH = max(rh) - min(rh),
            meanWS = mean(ws),
            minWS = min(ws),
            maxWS = max(ws),
            cvWS = sd(ws)/mean(ws),
            rangeWS = max(ws) - min(ws),
            meanRain = mean(rain),
            minRain = min(rain),
            maxRain = max(rain),
            cvRain = sd(rain)/mean(rain),
            rangeRain = max(rain) - min(rain),
            meanFFMC = mean(ffmc),
            minFFMC = min(ffmc),
            maxFFMC = max(ffmc),
            cvFFMC = sd(ffmc)/mean(ffmc),
            rangeFFMC = max(ffmc) - min(ffmc),
            meanDMC = mean(dmc),
            minDMC = min(dmc),
            maxDMC = max(dmc),
            cvDMC = sd(dmc)/mean(dmc),
            rangeDMC = max(dmc) - min(dmc),
            meanDC = mean(dc),
            minDC = min(dc),
            maxDC = max(dc),
            cvDC = sd(dc)/mean(dc),
            rangeDC = max(dc) - min(dc),
            meanISI = mean(isi),
            minISI = min(isi),
            maxISI = max(isi),
            cvISI = sd(isi)/mean(isi),
            rangeISI = max(isi) - min(isi),
            meanBUI = mean(bui),
            minBUI = min(bui),
            maxBUI = max(bui),
            cvBUI = sd(bui)/mean(bui),
            rangeBUI = max(bui) - min(bui),
            meanFWI = mean(fwi),
            minFWI = min(fwi),
            maxFWI = max(fwi),
            cvFWI = sd(fwi)/mean(fwi),
            rangeFWI = max(fwi) - min(fwi)) %>%
  data.table(.)

## merge with remaining data
setkey(summaryABSK_AllData, pixID)
setkey(ABSK_AllData, pixID)
cols <- grep("julDay|temp|rh|ws|rain|ffmc|dmc|dc|isi|bui|fwi", names(ABSK_AllData),
             invert = TRUE, value = TRUE)
summaryABSK_AllData <- summaryABSK_AllData[unique(ABSK_AllData[, ..cols])]

## make column with presence/absence of understory and remove the understorey layer
## as it can't really be trusted (Andison pers. comm. January 29th 2019)
## even considering its presence might be risky as some times aerial photos do not detect the presence of the understorey, even if it is present.
cols <- grep("SPEC._PER", names(summaryABSK_AllData), value = TRUE)
summaryABSK_AllData[LAYER == 2, UNDERSTOREY := rowSums(.SD, na.rm = TRUE) > 0, .SDcols = cols]
summaryABSK_AllData[, UNDERSTOREY2 := any(UNDERSTOREY), by = "pixID"]
summaryABSK_AllData[, UNDERSTOREY := UNDERSTOREY2]
summaryABSK_AllData[, UNDERSTOREY2 := NULL]
summaryABSK_AllData[is.na(UNDERSTOREY), UNDERSTOREY := FALSE]
summaryABSK_AllData <- summaryABSK_AllData[LAYER == 1]

## change NA cover to 0 and melt species and cover
## don't exclude pixels without vegetation
colA = grep("SPEC.$", names(summaryABSK_AllData), value = TRUE)
colB = grep("SPEC._PER", names(summaryABSK_AllData), value = TRUE)

for (j in which(names(summaryABSK_AllData) %in% colB))
  set(summaryABSK_AllData, which(is.na(summaryABSK_AllData[[j]])), j, 0)

summaryABSK_AllData <- melt(summaryABSK_AllData, measure = list(colA, colB), 
                           value.name = c("SPEC", "SPEC_PER"), variable.name = "SPEC_dominance")
rm(colA, colB); amc::.gc()

## and now re-cast to have a column per species
summaryABSK_AllData <- dcast(summaryABSK_AllData, ... ~ SPEC, value.var = "SPEC_PER", fill = 0)

## remove "NA" columns
cols <- grep("^NA$", names(summaryABSK_AllData))
summaryABSK_AllData[, (cols) := NULL]

## check for duplicates
summaryABSK_AllData <- unique(summaryABSK_AllData)
```

## COMMUNITY COMPOSITION/STRUCTURE PCAs - simplified data
TODO

```{r summarizedData_PCA, eval=FALSE, include=FALSE}
## HERE DO pca
## check if any spp has >100 cover
cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS", 
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)
summary(commCompStruc[, ..cols])

cols <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS", 
             names(commCompStruc), invert = TRUE, value = TRUE)
vegCompPCA <- rda(unique(na.omit(commCompStruc[, ..cols]))[,-grep("SEV_CONT", cols), with = FALSE], scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..cols]))$SEV_CONT)

## CHECK CROWN CLOSURE, lower and upper inversely correlated?! 
dev()
biplot(vegCompPCA, scaling = 2, type = c("text","points"))
plot(env.fit, labels = "SEVERITY")
```





