---
title: "Dave Andison's fire data preliminary analyses"
author: "Ceres Barros"
date: "March 28, 2018"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_knit$set(root.dir = "~/LandscapesInMotion")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, cache = 2)
library(ggplot2)
library(factoextra)
library(ggpubr)
library(permute)
library(vegan)
library(reshape2)
library(sf)
library(data.table)
library(dplyr)
library(xlsx)
library(foreign)
library(fasterize)
library(prioritizr)
library(reproducible)
library(quickPlot)
library(xtable)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.useMemoise" = FALSE,
        "reproducible.cachePath" = "analyses/cache",
        "future.globals.maxSize" = +Inf)

source("analyses/R_tools/summarizeABSK_AllData.R")
source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R")
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE

figDir <- "analyses/Figs"
if (!dir.exists(figDir))
  dir.create(figDir)
```

```{r dataprep, include = FALSE, cache = FALSE}
## source dataprep:
ABSK_AllData <- Cache(ABSKfires_DataPrep,
                      fireDataPath = "data/fires_Dave/fireSev",
                      vegDataPath = "data/fires_Dave/prefireVeg",
                      topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
                      weatherDataPath = "data/fires_Dave/fireWeather",
                      resolution = 30, ## cacheId = "468b1d728b735352"
                      doCache = doCache, bindAllFires = FALSE,
                      # useCache = FALSE,
                      userTags = "ABSKfires_DataPrep",
                      omitArgs = c("bindAllFires", "doCache", "userTags"))

## exclude event perimeter patch type
ABSK_AllData <- subset(x = ABSK_AllData, PatchType != "eventPerim")
amc::.gc()

## remove columns with no data
NAcols <- sapply(ABSK_AllData, FUN = function(x) all(is.na(x))) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!NAcols))]
zeroCols <- sapply(ABSK_AllData, FUN = function(x) all(x == 0, na.rm = TRUE)) 
ABSK_AllData <- ABSK_AllData[, .SD, .SDcols = names(which(!zeroCols))]

## calculate stand age at fire year by subtracting fire year.
ABSK_AllData[!is.na(ORIGIN_UPPER), STAND_AGE := FIRE_YEAR - ORIGIN_UPPER]

## convert SMR to ordinal variable (moister as values increase)
ABSK_AllData[SMR == "D", SMR_ord := 1]
ABSK_AllData[SMR == "F", SMR_ord := 2]
ABSK_AllData[SMR == "M", SMR_ord := 3]
ABSK_AllData[SMR == "W", SMR_ord := 4]
ABSK_AllData[SMR == "A", SMR_ord := 5]

ABSK_AllData[, SMR_ord := as.ordered(SMR_ord)]
```

## NOTES - meeting with Dave Dec 6th 2018
Careful with no. of variables and sample size
Weather variables: how to look at it? Weather at the beginning, middle, end? Only averages like Colin did? Averages, variation?

## COMMUNITY COMPOSITION/STRUCTURE PCAs
PCA of dominant species percentage cover and height across all fire event patch types. 
Blue arrow is severity in a continuous scale, fitted to the ordination *a posteriori*, using `env.fit`.
Severity is higher in "plots" dominated by White spruce (*Picea glauca*), Engelman's spruce (*P. engelmannii*), Aspen (*Populus tremuloides*) and where trees are taller.
It seems that canopy and understory levels usually highly correlated within species.

```{r PCA_prep, cache = FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR_ord|LAYER$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1+LAYER, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])

```

```{r PCA_plotFireName, fig.height=8, fig.show = "hold"}
## PCA COLOURED BY FIRE_NAME
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|Pice mari_3|Popu balb_1", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by fire name")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$FIRE_NAME))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
```

```{r PCA_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|Pice mari_3|Popu balb_1", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### PCA with canopy and understory levels collapsed for tree species (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but clearer.

```{r PCA2_prep, cache = FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|SPEC1_PER|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]
commCompStruc <- dcast(commCompStruc, ... ~ SPEC1, value.var = "SPEC1_PER", 
                       fun.aggregate = sum, na.rm = TRUE)

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR|Patch", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

# ## check if any spp has >100 cover
# cols <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
#              names(commCompStruc), invert = TRUE, value = TRUE)
# summary(commCompStruc[, ..cols])
```

```{r PCA2_plotClosure, fig.height=8, fig.show = "hold",  warning=FALSE}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|SMR$|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|balb", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### PCA with canopy and understory levels collapsed for tree species, but keeping dominant and subdominant spp separate (summed and rescaled to 0-1 - maybe not the best way)
Results are similar, but we now see some vegetation associations. Unlike before, crown closure now appears more correlated with severity. It seems higher severity levels coincide with forests where Aspen is the dominant species and White spruce is the subdominant. 

```{r PCA3_prep, cache = FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2|3)$|SPEC(1|2|3)_PER|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC3", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

spec3 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1|SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC3, value.var = "SPEC3_PER", fun.aggregate = sum, na.rm = TRUE)
setnames(spec3, sppNames, paste0(sppNames, "_3"), skip_absent = TRUE)

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
commCompStruc <- merge(commCompStruc, spec3, by = cols)
rm(spec1, spec2, spec3); amc::.gc()

cols <- grep("^NA", names(commCompStruc), value = TRUE)
set(commCompStruc, j = cols, value = NULL)

## check if any spp has >100 cover
# cols <- grep("_[[:digit:]]$", names(commCompStruc))
# summary(commCompStruc[, ..cols])
```

```{r PCA3_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|balb_1", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### Remove third subdominant species, add age
Correlations between third most dominant species and axes are quite low
Results changed quite a bit, with Lodgepole pine and crown closure appearing both more related to severity, than before (and Aspen and White spruce losing importance).

```{r PCA4_prep, cache = FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|STAND_AGE|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|ORIGIN|STAND_AGE|HEIGHT|SMR|CROWN_CLOSURE|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

cols <- grep("^NA", names(commCompStruc), value = TRUE)
set(commCompStruc, j = cols, value = NULL)
```

```{r, PCA4_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|balb_1", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### Remove stand age - correlations with axes are quite low

```{r PCA5_prep, cache = FALSE}
cols <- grep("SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData))
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("SEV|Patch|FIRE|pixID|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2); amc::.gc()

cols <- grep("^NA", names(commCompStruc), value = TRUE)
set(commCompStruc, j = cols, value = NULL)
```

```{r PCA5_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|bank|balb_1", colsA, invert = TRUE, value = TRUE)
vegCompPCA <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB]
                             , nf = 3
                             , scannf = FALSE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### Pure stands only (>80% cover)
Collapsing canopy and understory

```{r PCA6_prep, cache = FALSE}
## extend dominant species data
cols <- grep("SEV|Patch|FIRE|pixID|SPEC1$|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
             names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[SPEC1_PER >= 80, ..cols]

## some columns should be numeric
cols <- grep("FIRE_NAME|SMR_ord|Patch|SPEC1", names(commCompStruc), invert = TRUE, value = TRUE)
commCompStruc <- commCompStruc[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## SPEC1 needs to be a factor for dudi
commCompStruc <- commCompStruc[, SPEC1 := as.factor(SPEC1)]
```

```{r PCA6_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE", colsA, invert = TRUE, value = TRUE)
vegCompDudi <- ade4::dudi.mix(unique(na.omit(commCompStruc[, ..colsA]))[, ..colsB],
                              scannf = FALSE, nf = 3)
env.fit <- envfit(vegCompDudi, env = unique(na.omit(commCompStruc[, ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompDudi, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[, ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


## COMMUNITY COMPOSITION/STRUCTURE PCAs - FOOTHILLS AND BOREAL SEPARATED

### PCA with canopy and understory levels collapsed for tree species, but keeping dominant and subdominant spp separate (summed and rescaled to 0-1 - maybe not the best way)
Only 1st and 2nd most dominant species included (same as 5th PCA)

Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r FootVsBoreal_PCA_prep, cache = FALSE}
cols <- grep("NR1|SEV|Patch|FIRE|pixID|SPEC(1|2)$|SPEC(1|2)_PER|HEIGHT|CROWN_CLOSURE|SMR_ord|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(ABSK_AllData), value = TRUE)
commCompStruc <- ABSK_AllData[!SPEC1 %in% c("XXXX MISS", "NA", NA),
                              ..cols]

spec1 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC2", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC1 , value.var = "SPEC1_PER", fun.aggregate = sum, na.rm = TRUE)
sppNames <- grep("NR1|SEV|Patch|FIRE|pixID|ORIGIN|STAND_AGE|HEIGHT|CROWN_CLOSURE|SMR|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", names(spec1), invert = TRUE, value = TRUE)
setnames(spec1, sppNames, paste0(sppNames, "_1"))

spec2 <- dcast(commCompStruc[, .SD, .SDcols = grep("SPEC1", names(commCompStruc), invert = TRUE)],
               ... ~ SPEC2 , value.var = "SPEC2_PER", fun.aggregate = sum, na.rm = TRUE) 
setnames(spec2, sppNames, paste0(sppNames, "_2"))

cols <- grep(paste(sppNames, collapse = "|"), names(spec1), invert = TRUE, value = TRUE)
rm(commCompStruc); amc::.gc()
commCompStruc <- merge(spec1, spec2, by = cols)
rm(spec1, spec2, sppNames); amc::.gc()

commCompStruc$`NA` <- NULL
```

```{r Foot_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## COLOURED BY CROWN CLOSURE
## columns to keep in table
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
## columns to use for PCA
colsB <- grep("SEV_CONT|FIRE|NR1", colsA, invert = TRUE, value = TRUE)

commCompStruc[, SMR_ord := as.integer(SMR_ord)]

## FOOTHILLS + ROCKY MOUNTAIN
vegCompPCA <- unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- envfit(vegCompPCA, env = unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))$SEV_CONT)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[grep("Foothills|Rocky", NR1), ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## BOREAL FOREST + CANADIAN SHIELD
vegCompPCA <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SEV_CONT %>%
  envfit(vegCompPCA, env = .)

# dev()
PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_PCA_plotSMR, fig.height=8, fig.show = "hold"}
## colour by SMR in Boreal
colsA <- grep("SEV_CLASS|Patch|FIRE|pixID|Elevation|Aspect|temp|rh|rain|dmc|dc|bui|fwi|ws", 
              names(commCompStruc), invert = TRUE, value = TRUE)
colsB <- grep("SEV_CONT|FIRE|NR1|SMR", colsA, invert = TRUE, value = TRUE)

vegCompPCA <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))[, ..colsB] %>%
  rda(., scale = TRUE)
env.fit <- unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SEV_CONT %>%
  envfit(vegCompPCA, env = .)

PCAplot <- ordiplot(vegCompPCA, type = "none", main = "Coloured by SMR\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SMR))
text(PCAplot, "species", labels = rep("", length(colsB)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(unique(na.omit(commCompStruc[grep("Boreal|Canadian", NR1), ..colsA]))$SMR)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


## SIMPLIFY ANALYSIS/DATA 
- Collapse climate data using summary statistics
- Convert species data to long format

```{r summaryData, cache = FALSE}
## clean-up and free memory
rm(vegCompPCA, vegCompDudi, env.fit, commCompStruc, PCAplot)
amc::.gc()

# summarise weather data - faster if not cached
summaryABSK_AllData <- Cache(summarizeABSK_AllData,
                             DT = ABSK_AllData,
                             dim = dim(ABSK_AllData),
                             days = 1:3,
                             omitArgs = c("DT"),
                             useCache = doCache) ## cacheId = "7f5e7f0ae8fe1943"
## collapse species into more generic categories
summaryABSK_AllData[, FlamConif := sum(`Pice glau`, `Pice mari`, `Pice enge`,
                                       `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Abie := sum(`Abie bals`, `Abie lasi`, na.rm = TRUE),
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")

collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if (all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]

summaryABSK_AllData[, isForest := 0]
summaryABSK_AllData[NonForestValue == "NA", isForest := 1]
summaryABSK_AllData[!is.na(WETLAND_CLASS) & !WETLAND_VEG_MOD %in% c("F", "T"), isForest := 0]   ## some wetlands are not forests

## change Lari lari to simpler name
setnames(summaryABSK_AllData, "Lari lari", "Lari")

## rescale severity from 0-1
summaryABSK_AllData[, SEV_PROP := SEV_CONT/100]
## convert the highest fire severity to 1
summaryABSK_AllData[SEV_PROP == 0.9750, SEV_PROP := 1.0]

## change some columns to numeric
cols <- grep("mean|max|min|cv|range|pixID|P_ID|Duration|ORIGIN|HEIGHT|isForest|SEV_PROP|Aspect|^CTI$|PctSlope|SurfReliefRatio|SurfAspectRatio|Elevation", names(summaryABSK_AllData), value = TRUE)
str(summaryABSK_AllData[, ..cols])
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]

## calculate SMR_wet
summaryABSK_AllData[, SMR_wet := as.integer(grepl("4|5", SMR_ord))]

## make some columns factors
cols <- c("FIRE_NAME", "Czone1", "NR1", "NonForestValue")
summaryABSK_AllData <- summaryABSK_AllData[, (cols) := lapply(.SD, as.factor), .SDcols = cols]
amc::.gc()
```

## COMMUNITY COMPOSITION/STRUCTURE PCAs - simplified data
Including elevation, slope, presence/absence of understorey and CTI, crown closure and height in PCA then fitting environmental vectors (mean and cv of fire weather indices) and severity. Species are still separate

Coloured by FIRE_NAME
```{r summaryData_PCA_plotFIRE_NAME, cache = FALSE, fig.height=8, fig.show = "hold"}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- c(grep("^(Popu|Pinu|Lari|Pice|Abie |Betu)", names(summaryABSK_AllData), value = TRUE),
          grep("isForest|NonForestValue|pixID", names(summaryABSK_AllData), value = TRUE),
          grep("SMR_ord|CROWN.*UPPER|HEIGHT.*UPPER|UNDERSTOREY", names(summaryABSK_AllData), value = TRUE),
          grep("SEV_CONT|FIRE_NAME|Czone|NR1", names(summaryABSK_AllData), value = TRUE),
          grep("^mean(DMC|DC|ISI|FFMC|BUI|FWI)", names(summaryABSK_AllData), value = TRUE),
          grep("^cv(DMC|DC|ISI|FFMC|BUI|FWI)", names(summaryABSK_AllData), value = TRUE),
          grep("PctSlope|Elevation|^CTI$", names(summaryABSK_AllData), value = TRUE))


summaryABSK_AllData2 <- na.omit(summaryABSK_AllData[, ..cols])
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix
summaryABSK_AllData2[, isForest := as.factor(as.character(isForest))] ## convert to factor for dudi.mix
summaryABSK_AllData2[, NonForestValue := as.factor(as.character(NonForestValue))] ## convert to factor for dudi.mix

## make a PCA without climate/severity/pixID
## exclude also P. banksiana, because it has no data
cols2 <- grep("SEV_CONT|bank|mean|cv|pixID|NonForestValue|isForest|FIRE_NAME|Czone|NR1", 
              cols, value = TRUE, invert = TRUE)
summDataPCA <- ade4::dudi.mix(as.data.frame(summaryABSK_AllData2[, ..cols2])
                              , nf = 3
                              , scannf = FALSE)

## superimpose severity and climate
cols3 <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2[, ..cols3]))

plotDataPCApoints <- data.table(summDataPCA$li)[, 1:2]
setnames(plotDataPCApoints, new = c("x", "y"))
plotDataPCApoints$isForest <- summaryABSK_AllData2$isForest
plotDataPCApoints$NonForestValue <- summaryABSK_AllData2$NonForestValue
plotDataPCApoints$FIRE_NAME <- summaryABSK_AllData2$FIRE_NAME
plotDataPCApoints$Czone1 <- summaryABSK_AllData2$Czone1
plotDataPCApoints$NR1 <- summaryABSK_AllData2$NR1

plotDataPCAvect <- summDataPCA$co[, 1:2]
plotDataPCAvect$labels <- rownames(plotDataPCAvect)
plotDataPCAvect <- data.table(plotDataPCAvect)
setnames(plotDataPCAvect, new = c("x", "y", "labels"))

plotDataEnvFitvect <- data.frame(env.fit$vectors$arrows)[, 1:2]
plotDataEnvFitvect$labels <- rownames(plotDataEnvFitvect)
plotDataEnvFitvect <- data.table(plotDataEnvFitvect)
setnames(plotDataEnvFitvect, new = c("x", "y", "labels"))

colLabs <- sort(unique(plotDataPCApoints$FIRE_NAME))
names(colLabs) <- as.numeric(as.factor(colLabs))

plot1 <- ggplot() +
  geom_point(data = plotDataPCApoints[isForest == 1], 
             aes(x = x, y = y, color = as.numeric(as.factor(FIRE_NAME))),
             alpha = 0.5, size = 1) +
  geom_segment(data = plotDataPCAvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1) +
  geom_text(data = plotDataPCAvect,
            aes(x = x, y = y, label = labels), 
            size = 5, vjust = 1.2, color = "black") +
  geom_segment(data = plotDataEnvFitvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1, linetype = "dashed") +
  geom_text(data = plotDataEnvFitvect,
            aes(x = x, y = y, label = labels),
            size = 5, vjust = 1.2, color = "black") +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) +
  scale_colour_distiller(palette = "Spectral", 
                         breaks = as.numeric(names(colLabs)), 
                         labels = colLabs, guide = "legend") +
  theme_pubr(base_size = 12) + 
  theme(legend.direction = "vertical") +
  labs(x = "PC 1", y = "PC 2", colour = "") +
  guides(colour = guide_legend(override.aes = list(size = 2)))

plot1Zoom <- plot1 +
  scale_x_continuous(limits = c(-1.1,1.1)) +
  scale_y_continuous(limits = c(-1.1,1.1))

ggarrange(plot1, plot1Zoom, common.legend = TRUE, legend = "right")

```

Coloured by Czone1 and NR1
```{r summaryData_PCA_plotCzone1NR1}
plot2 <- ggplot() +
  geom_point(data = plotDataPCApoints[isForest == 1], 
             aes(x = x, y = y, color = paste(Czone1, NR1, sep = " - ")),
             alpha = 0.5, size = 1) +
  geom_segment(data = plotDataPCAvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", 
               size = 1) +
  geom_text(data = plotDataPCAvect,
            aes(x = x, y = y, label = labels), 
            size = 5, vjust = 1.2, 
            color =  "black") +
  geom_segment(data = plotDataEnvFitvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color =  "black", linetype = "dashed", size = 1) +
  geom_text(data = plotDataEnvFitvect,
            aes(x = x, y = y, label = labels),
            size = 5, vjust = 1.2, 
            color =  "black") +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) +
  scale_colour_brewer(palette = "Dark2") +
  theme_pubr(base_size = 12) + 
  theme(legend.direction = "vertical") +
  labs(x = "PC 1", y = "PC 2", colour = "") +
  guides(colour = guide_legend(override.aes = list(size = 2)))

plot2Zoom <- plot2 +
  scale_x_continuous(limits = c(-1.1,1.1)) +
  scale_y_continuous(limits = c(-1.1,1.1))

ggarrange(plot2, plot2Zoom, common.legend = TRUE, legend = "right")
```

```{r savePlots2, include = FALSE}
png(filename = file.path(figDir, "PCAforestComp_climate_byFire.png"), res = 300, 
    width = 14, height = 6, units = "in")
ggarrange(plot1, plot1Zoom, common.legend = TRUE, legend = "right")
dev.off()

png(filename = file.path(figDir, "PCAforestComp_climate_byCzoneNR.png"), res = 300, 
    width = 14, height = 6, units = "in")
ggarrange(plot2, plot2Zoom, common.legend = TRUE, legend = "right")
dev.off()
```


### Now separating Foothills + Rocky Mountain from Boreal Forest + Canadian Shield
Including elevation, slope, presence/absence of understorey in PCA then fitting environmental vectors (mean and cv of temp, rain, rh, ffmc, fwi) and severity. Species are still separate

Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r FootVsBoreal_summaryData_PCA_prep, cache = FALSE}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR$|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR2", names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix
```

```{r Foot_summaryData_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## FOOTHILLS + ROCKY MOUNTAIN
## exclude also P. banksiana, because it has no data
colsA <- grep("SEV_CONT|bank|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|mean|cv|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData_PCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
# unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME)) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData_PCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR \nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## BOREAL FOREST + CANADIAN SHIELD
## exclude also P. banksiana, P. engelmann and A. lasiocarpa because they have no data
colsA <- grep("SEV_CONT|bank|enge|lasi|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|enge|lasi|NR1|mean|cv|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData_PCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
# unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME)) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData_PCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

## COMMUNITY COMPOSITION/STRUCTURE PCAs - simplified data AND collapsed species
Including elevation, slope, presence/absence of understorey and CTI, crown closure and height in PCA then fitting environmental vectors (mean and cv of fire weather indices) and severity. Species were collapsed as:

* Group fire resistant conifers
+ Pinu cont
+ Pinu bank
+ Pice enge
* Group fire sensitive conifers
+ Abie lasi	
+ Abie bals
+ Pice glau
* Group Larix
+ Lari lari	
* Group Pice mari
+ Pice mari	
* Group deciduous
+ Betu papy	
+ Popu trem	
+ Popu balb


Coloured by FIRE_NAME
```{r summaryData2_PCA_plotFIRE_NAME, fig.height=8, fig.show = "hold"}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- c(grep("FlamConif|Decid|Lari|Abie$", names(summaryABSK_AllData), value = TRUE),
          grep("isForest|NonForestValue|pixID", names(summaryABSK_AllData), value = TRUE),
          grep("SMR_ord|CROWN.*UPPER|HEIGHT.*UPPER|UNDERSTOREY", names(summaryABSK_AllData), value = TRUE),
          grep("SEV_CONT|FIRE_NAME|Czone|NR1", names(summaryABSK_AllData), value = TRUE),
          grep("^mean(DMC|DC|ISI|FFMC|BUI|FWI)", names(summaryABSK_AllData), value = TRUE),
          grep("^cv(DMC|DC|ISI|FFMC|BUI|FWI)", names(summaryABSK_AllData), value = TRUE),
          grep("PctSlope|Elevation|^CTI$", names(summaryABSK_AllData), value = TRUE))


summaryABSK_AllData2 <- na.omit(summaryABSK_AllData[, ..cols])
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix
summaryABSK_AllData2[, isForest := as.factor(as.character(isForest))] ## convert to factor for dudi.mix
summaryABSK_AllData2[, NonForestValue := as.factor(as.character(NonForestValue))] ## convert to factor for dudi.mix

## make a PCA without climate/severity/pixID
## exclude also P. banksiana, because it has no data
cols2 <- grep("SEV_CONT|bank|mean|cv|pixID|NonForestValue|isForest|FIRE_NAME|Czone|NR1", 
              cols, value = TRUE, invert = TRUE)
summDataPCA <- ade4::dudi.mix(as.data.frame(summaryABSK_AllData2[, ..cols2])
                              , nf = 3
                              , scannf = FALSE)

## superimpose severity and climate
cols3 <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2[, ..cols3]))

plotDataPCApoints <- data.table(summDataPCA$li)[, 1:2]
setnames(plotDataPCApoints, new = c("x", "y"))
plotDataPCApoints$isForest <- summaryABSK_AllData2$isForest
plotDataPCApoints$NonForestValue <- summaryABSK_AllData2$NonForestValue
plotDataPCApoints$FIRE_NAME <- summaryABSK_AllData2$FIRE_NAME
plotDataPCApoints$Czone1 <- summaryABSK_AllData2$Czone1
plotDataPCApoints$NR1 <- summaryABSK_AllData2$NR1

plotDataPCAvect <- summDataPCA$co[, 1:2]
plotDataPCAvect$labels <- rownames(plotDataPCAvect)
plotDataPCAvect <- data.table(plotDataPCAvect)
setnames(plotDataPCAvect, new = c("x", "y", "labels"))

plotDataEnvFitvect <- data.frame(env.fit$vectors$arrows)[, 1:2]
plotDataEnvFitvect$labels <- rownames(plotDataEnvFitvect)
plotDataEnvFitvect <- data.table(plotDataEnvFitvect)
setnames(plotDataEnvFitvect, new = c("x", "y", "labels"))

colLabs <- sort(unique(plotDataPCApoints$FIRE_NAME))
names(colLabs) <- as.numeric(as.factor(colLabs))

plot1 <- ggplot() +
  geom_point(data = plotDataPCApoints[isForest == 1], 
             aes(x = x, y = y, color = as.numeric(as.factor(FIRE_NAME))),
             alpha = 0.5, size = 1) +
  geom_segment(data = plotDataPCAvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1) +
  geom_text(data = plotDataPCAvect,
            aes(x = x, y = y, label = labels), 
            size = 5, vjust = 1.2, color = "black") +
  geom_segment(data = plotDataEnvFitvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1, linetype = "dashed") +
  geom_text(data = plotDataEnvFitvect,
            aes(x = x, y = y, label = labels),
            size = 5, vjust = 1.2, color = "black") +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) +
  scale_colour_distiller(palette = "Spectral", 
                         breaks = as.numeric(names(colLabs)), 
                         labels = colLabs, guide = "legend") +
  theme_pubr(base_size = 12) + 
  theme(legend.direction = "vertical") +
  labs(x = "PC 1", y = "PC 2", colour = "") +
  guides(colour = guide_legend(override.aes = list(size = 2)))

plot1Zoom <- plot1 +
  scale_x_continuous(limits = c(-1.1,1.1)) +
  scale_y_continuous(limits = c(-1.1,1.1))

ggarrange(plot1, plot1Zoom, common.legend = TRUE, legend = "right")

```

Coloured by Czone1 and NR1
```{r summaryData2_PCA_plotCzone1NR1, fig.height=8, fig.show = "hold"}
plot2 <- ggplot() +
  geom_point(data = plotDataPCApoints[isForest == 1], 
             aes(x = x, y = y, color = paste(Czone1, NR1, sep = " - ")),
             alpha = 0.5, size = 1) +
  geom_segment(data = plotDataPCAvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", 
               size = 1) +
  geom_text(data = plotDataPCAvect,
            aes(x = x, y = y, label = labels), 
            size = 5, vjust = 1.2, 
            color =  "black") +
  geom_segment(data = plotDataEnvFitvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color =  "black", linetype = "dashed", size = 1) +
  geom_text(data = plotDataEnvFitvect,
            aes(x = x, y = y, label = labels),
            size = 5, vjust = 1.2, 
            color =  "black") +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) +
  scale_colour_brewer(palette = "Dark2") +
  theme_pubr(base_size = 12) + 
  theme(legend.direction = "vertical") +
  labs(x = "PC 1", y = "PC 2", colour = "") +
  guides(colour = guide_legend(override.aes = list(size = 2)))

plot2Zoom <- plot2 +
  scale_x_continuous(limits = c(-1.1,1.1)) +
  scale_y_continuous(limits = c(-1.1,1.1))

ggarrange(plot2, plot2Zoom, common.legend = TRUE, legend = "right")
```

Coloured by SMR
```{r summaryData2_PCA_plotSMR, fig.height=8, fig.show = "hold"}
plotDataPCApoints <- data.table(summDataPCA$li)[, 1:2]
setnames(plotDataPCApoints, new = c("x", "y"))
plotDataPCApoints$isForest <- summaryABSK_AllData2$isForest
plotDataPCApoints$NonForestValue <- summaryABSK_AllData2$NonForestValue
plotDataPCApoints$FIRE_NAME <- summaryABSK_AllData2$FIRE_NAME
plotDataPCApoints$Czone1 <- summaryABSK_AllData2$Czone1
plotDataPCApoints$NR1 <- summaryABSK_AllData2$NR1
plotDataPCApoints$SMR_ord <- summaryABSK_AllData2$SMR_ord

plotDataPCAvect <- summDataPCA$co[, 1:2]
plotDataPCAvect$labels <- rownames(plotDataPCAvect)
plotDataPCAvect <- data.table(plotDataPCAvect)
setnames(plotDataPCAvect, new = c("x", "y", "labels"))

plotDataEnvFitvect <- data.frame(env.fit$vectors$arrows)[, 1:2]
plotDataEnvFitvect$labels <- rownames(plotDataEnvFitvect)
plotDataEnvFitvect <- data.table(plotDataEnvFitvect)
setnames(plotDataEnvFitvect, new = c("x", "y", "labels"))

colLabs <- sort(unique(plotDataPCApoints$SMR_ord))
names(colLabs) <- as.numeric(as.factor(colLabs))

plot3 <- ggplot() +
  geom_point(data = plotDataPCApoints[isForest == 1], 
             aes(x = x, y = y, color = SMR_ord),
             alpha = 0.5, size = 1) +
  geom_segment(data = plotDataPCAvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1) +
  geom_text(data = plotDataPCAvect,
            aes(x = x, y = y, label = labels), 
            size = 5, vjust = 1.2, color = "black") +
  geom_segment(data = plotDataEnvFitvect,
               aes(x = 0, y = 0, xend = x, yend = y),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black", size = 1, linetype = "dashed") +
  geom_text(data = plotDataEnvFitvect,
            aes(x = x, y = y, label = labels),
            size = 5, vjust = 1.2, color = "black") +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) +
  scale_colour_brewer(palette = "Spectral", 
                      breaks = as.numeric(names(colLabs)), 
                      labels = colLabs, guide = "legend") +
  theme_pubr(base_size = 12) + 
  theme(legend.direction = "vertical") +
  labs(x = "PC 1", y = "PC 2", colour = "SMR") +
  guides(colour = guide_legend(override.aes = list(size = 2)))

plot3Zoom <- plot3 +
  scale_x_continuous(limits = c(-1.1,1.1)) +
  scale_y_continuous(limits = c(-1.1,1.1))

ggarrange(plot3, plot3Zoom, common.legend = TRUE, legend = "right")
```


```{r savePlots, include = FALSE}
png(filename = file.path(figDir, "PCAforestCompSppcollapse_climate_byFire.png"), res = 300, 
    width = 14, height = 6, units = "in")
ggarrange(plot1, plot1Zoom, common.legend = TRUE, legend = "right")
dev.off()

png(filename = file.path(figDir, "PCAforestCompSppcollapse_climate_byCzoneNR.png"), res = 300, 
    width = 14, height = 6, units = "in")
ggarrange(plot2, plot2Zoom, common.legend = TRUE, legend = "right")
dev.off()

png(filename = file.path(figDir, "PCAforestCompSppcollapse_climate_bySMR.png"), res = 300, 
    width = 14, height = 6, units = "in")
ggarrange(plot3, plot3Zoom, common.legend = TRUE, legend = "right")
dev.off()
```


## COMMUNITY COMPOSITION/STRUCTURE + FIRE WEATHER PCAs - simplified data
PCAs calculated on variables of vegetation composition **and** average and covariance of temp, rain, rh, ffmc, fwi
Severity fitted **a posteriori**

```{r summaryData_VegWeatherPCA_plotClosure, cache = FALSE, fig.height=8, fig.show = "hold"}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE_Y|pixID|SMR$|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR1|NR2", 
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix

## exclude also P. banksiana, because it has no data
cols2 <- grep("SEV_CONT|bank|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- ade4::dudi.mix(as.data.frame(summaryABSK_AllData2[, ..cols2])
                              , nf = 3
                              , scannf = FALSE)

env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2[, SEV_CONT]))

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[["CROWN_CLOSURE_LOWER"]]))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
unique( as.factor(summaryABSK_AllData2[["CROWN_CLOSURE_LOWER"]])) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r summaryData_VegWeatherPCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR")
points(PCAplot, "sites", pch = 19, 
       col = summaryABSK_AllData2[["SMR_ord"]])
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
sort(unique(summaryABSK_AllData2[["SMR_ord"]])) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r summaryData_VegWeatherPCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[["FIRE_NAME"]]))
text(PCAplot, "species", arrows = TRUE)
plot(env.fit, labels = "SEVERITY")
# unique(as.factor(summaryABSK_AllData2[["FIRE_NAME"]])) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

### Now separating Foothills + Rocky Mountain from Boreal Forest + Canadian Shield
Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r FootVsBoreal_summaryData_VegWeatherPCA_prep, cache = FALSE}
## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("SEV_CLASS|Patch|FIRE_YEAR|pixID|SMR$|Aspect|DMC|DC|BUI|WS|JulDay|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre\\.|CTI|Curv|SCOS|SSIN|Surf|TRASP|DegSlope|SlopePosition|Prov|Long|Lat|SPEC|STAND|SITE|min|max|range|Czone|NR2", names(summaryABSK_AllData), invert = TRUE, value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix
```

```{r Foot_summaryData_VegWeatherPCA_plotClosure, fig.height=8, fig.show = "hold"}
## FOOTHILLS + ROCKY MOUNTAIN
## exclude also P. banksiana, because it has no data
colsA <- grep("SEV_CONT|bank|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|NR1", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsC][, SEV_CONT]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData_VegWeatherPCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
# unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME)) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData_VegWeatherPCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord)
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
sort(unique(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData_VegWeatherPCA_plotClosure, fig.height=8, fig.show = "hold"}
## BOREAL FOREST + CANADIAN SHIELD
## exclude also P. banksiana, P. engelmann and A. lasiocarpa because they have no data
colsA <- grep("SEV_CONT|bank|enge|lasi|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|bank|enge|lasi|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|NR1", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsC][, SEV_CONT]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData_VegWeatherPCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```

```{r Boreal_summaryData_VegWeatherPCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord)
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
sort(unique(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

## FIRE WEATHER PCA - simplified data
Average and covariance of temp, rain, rh, ffmc, fwi
Also including elevation, slope, and then fitting everity

```{r weather_summaryData_PCA_plotFireName, cache = FALSE, fig.height=8, fig.show = "hold"}
rm(summaryABSK_AllData2); amc::.gc()

## check if any spp has >100 cover - good
# cols <- grep("SEV|Patch|FIRE|pixID|SMR|Elevation|Aspect|Temp|RH|Rain|DMC|DC|BUI|FWI|WS|JulDay|FFMC|ISI|DIST|NON|LAYER|WET|ORIG|PROD|Duration|Area|fire|ID|Pre.|CTI|Curv|Slope|SCOS|SSIN|Surf|TRASP|Prov|Long|Lat|SPEC|STAND|SITE|CROWN|HEIGHT|UNDER", 
#              names(summaryABSK_AllData), invert = TRUE, value = TRUE)
# summary(summaryABSK_AllData[, ..cols])

## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("mean|cv|Elev|PctSlope|NR1|SEV_CONT|FIRE_NAME", 
             names(summaryABSK_AllData), value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))

## exclude also P. banksiana, because it has no data
cols2 <- grep("mean|cv|Elev|Slope", cols, value = TRUE)
summDataPCA <- ade4::dudi.mix(as.data.frame(summaryABSK_AllData2[, ..cols2])
                              , nf = 3
                              , scannf = FALSE)
env.fit <- envfit(summDataPCA, env = as.data.frame(summaryABSK_AllData2[, SEV_CONT]))

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2$FIRE_NAME))
text(PCAplot, "species", labels = rep("", length(cols2)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```

### Now separating Foothills + Rocky Mountain from Boreal Forest + Canadian Shield
Foothills and Rocky Mountain collapsed
Boreal Forest and Canadian Shield collapsed

```{r Foot_weather_summaryData_PCA_plotFireName, fig.height=8, fig.show = "hold"}
## FOOTHILLS + ROCKY MOUNTAIN
## exclude also P. banksiana, because it has no data
colsA <- grep("SEV_CONT|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|NR1", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsC][, SEV_CONT]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```


```{r Boreal_weather_summaryData_PCA_plotFireName, fig.height=8, fig.show = "hold"}
## BOREAL FOREST + CANADIAN SHIELD
## exclude also P. banksiana, P. engelmann and A. lasiocarpa because they have no data
colsA <- grep("SEV_CONT|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|NR1", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsC][, SEV_CONT]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = "SEVERITY")
```


### Correlations
Pearson and spearman correlations between all covariates and each covariate and severity

#### Vegetation and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and STAND_AGE removed as they are redundant

```{r vegTopoCorrPearson}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND_S|SITE|NON_|UNDERSTOREY|ORIGIN|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
```

```{r vegTopoCorrPearsonPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Pearson correlations between fire weather and topo covariates only
SurfAspectRatio adn DegSlope removed due to high correlation with PctSlope

```{r weatherTopoCorrPearson}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|DegSlope|SurfAspectRatio|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND_S|SITE|NON_|UNDERSTOREY|HEIGHT|CROWN|ORIGIN|STAND_AGE|pixID|Czone|^NR1|^NR2|Lari|Pice|Pinu|Abie|Popu|Betu",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
```

```{r weatherTopoCorrPearsonPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Spearman correlations Vegetation and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and STAND_AGE removed as they are redundant

```{r vegTopoCorrSpearman}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND_S|SITE|NON_|UNDERSTOREY|ORIGIN|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
```

```{r vegTopoCorrSpearmanPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Spearman correlations between fire weather and topo covariates only
SurfAspectRatio and DegSlope removed due to high correlation with PctSlope
Note that minRain is always 0, hence the lack of correlation value

```{r weatherTopoCorrSpearman}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|DegSlope|SurfAspectRatio|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR|STAND_S|SITE|NON_|UNDERSTOREY|HEIGHT|CROWN|ORIGIN|STAND_AGE|pixID|Czone|^NR1|^NR2|startJulDay|Lari|Pice|Pinu|Abie|Popu|Betu",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl$minRain <- NULL
tabl <- tabl[which(rownames(tabl) != "minRain"),]
```

```{r weatherTopoCorrSpearmanPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Pearson correlations between severity and vegetation and topo
Correlations are significant but very low

```{r sevVegTopoCorrPearson}
cols <- grep("SEV_CLASS|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND_S|SITE|NON_|UNDERSTOREY|ORIGIN|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

cols  <- c(cols[!cols %in% "SEV_CONT"], "SEV_CONT")   ## reorder
tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))

corrVals <- as.numeric(gsub("\\*", "", tabl[, "SEV_CONT"]))
tabl <- tabl[order(abs(corrVals), decreasing = TRUE),, drop = FALSE]
```

```{r sevVegTopoCorrPearsonPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Pearson correlations between severity and weather
Correlations are significant but very low (only slightly higher than those with veg)

```{r sevWeatherTopoCorrPearson}
cols <- grep("SEV_CONT|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), value = TRUE)

cols  <- c(cols[!cols %in% "SEV_CONT"], "SEV_CONT")   ## reorder
tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "pearson")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))

corrVals <- as.numeric(gsub("\\*", "", tabl[, "SEV_CONT"]))
tabl <- tabl[order(abs(corrVals), decreasing = TRUE),, drop = FALSE]
```

```{r sevWeatherTopoCorrPearsonPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Spearman correlations between severity and vegetation and topo
Correlations imporved in some cases and got worse in others.

```{r sevVegTopoCorrSpearman}
cols <- grep("SEV_CLASS|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND_S|SITE|NON_|UNDERSTOREY|ORIGIN|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

cols  <- c(cols[!cols %in% "SEV_CONT"], "SEV_CONT")   ## reorder
tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))

corrVals <- as.numeric(gsub("\\*", "", tabl[, "SEV_CONT"]))
tabl <- tabl[order(abs(corrVals), decreasing = TRUE),, drop = FALSE]
```

```{r sevVegTopoCorrSpearmanPrint, results="asis"}
print(xtable(tabl), type = "html")
```

### Spearman correlations between severity and weather
Correlations are significant but very low (only slightly higher than those with veg).
Note that minRain is always 0, hence the lack of correlation value

```{r sevWeatherTopoCorrSpearman}
cols <- grep("SEV_CONT|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), value = TRUE)

cols  <- c(cols[!cols %in% "SEV_CONT"], "SEV_CONT")   ## reorder
tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))
tabl <- tabl[which(rownames(tabl) != "minRain"),, drop = FALSE]

corrVals <- as.numeric(gsub("\\*", "", tabl[, "SEV_CONT"]))
tabl <- tabl[order(abs(corrVals), decreasing = TRUE),, drop = FALSE]
```

```{r sevWeatherTopoCorrSpearmanPrint, results = "asis"}
print(xtable(tabl), type = "html")
```


## RELATIONSHIPS BETWEEN BEST CORRELATED VARIABLES

### Plots
#### Severity and vegetation covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and ORIGIN removed as they are redundant

```{r vegSevPlots, fig.height=8, fig.show = "hold", cache=FALSE}
cols <- grep("SEV_CLASS|FIRE_|fireNum|Patch|Province|P_ID|gridID|Aspect|CTI|Curv|Slope|SCOS|Elev|SSIN|Surf|TRASP|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND_S|SITE|NON_|ORIGIN|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)
tablePlot <- melt(summaryABSK_AllData[, ..cols], id.vars = "SEV_CONT")
tablePlot[value == "TRUE", value := "1"]
tablePlot[value == "FALSE", value := "0"]
tablePlot[, value := as.numeric(value)]

plot1 <- ggplot(tablePlot, aes(x = value, y = SEV_CONT, group = value)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free")
```

```{r vegSevPlotsPrint, fig.height=8, fig.show = "hold"}
plot1
```

#### Severity and topo covariates only (only highest correlations)
```{r topoPairsPlot, fig.height=8, fig.show = "hold", cache=FALSE}
cols <- grep("Aspect|^CTI$|SurfReliefRatio|PctSlope|Elevation|SEV_CONT",
             names(summaryABSK_AllData), value = TRUE)

tablePlot <- melt(summaryABSK_AllData[, ..cols], id.vars = "SEV_CONT")
tablePlot$valueCuts <- cut(tablePlot$value, breaks = seq(min(tablePlot$value), max(tablePlot$value),
                                                         length.out = 10))
plot1 <- ggplot(tablePlot, aes(x = valueCuts, y = SEV_CONT)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free")
```

```{r topoPairsPlotPrint, fig.height=8, fig.show = "hold"}
plot1
```

#### Severity and fire weather covariates only (only highest correlations)

```{r weatherTopoPairsPlot, fig.height=8, fig.show = "hold", cache=FALSE}
cols <- grep("SEV_CONT|meanFFMC|cvDMC|cvBUI|minFFMC",
             names(summaryABSK_AllData), value = TRUE)

tablePlot <- melt(summaryABSK_AllData[, ..cols], id.vars = "SEV_CONT")
tablePlot$valueCuts <- cut(tablePlot$value, breaks = seq(min(tablePlot$value), max(tablePlot$value),
                                                         length.out = 10))
plot1 <- ggplot(tablePlot, aes(x = valueCuts, y = SEV_CONT)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free")
```

```{r weatherTopoPairsPlotPrint, fig.height=8, fig.show = "hold"}
plot1
```

Similar variables but using means
```{r weatherTopoPairsPlot2, fig.height=8, fig.show = "hold", cache=FALSE}
cols <- grep("SEV_CONT|meanDMC|meanBUI|meanFFMC|meanISI|meanFWI",
             names(summaryABSK_AllData), value = TRUE)

tablePlot <- melt(summaryABSK_AllData[, ..cols], id.vars = "SEV_CONT")

plot1 <- ggplot(tablePlot, aes(x = value, y = SEV_CONT)) +
  geom_point() +
  facet_wrap(~ variable, scales = "free")
```

```{r weatherTopoPairsPlot2Print, fig.height=8, fig.show = "hold"}
plot1
```


## SIMPLIFY VEGETATION ANALYSIS/DATA 
Make variable with non forested classes:
* NATURALLY_NON_VEG
* NON_FORESTED_VEG
* NON_FORESTED_ANTHRO

Group species as:
* Group fire resistant conifers
+ Pinu cont
+ Pinu bank
+ Pice enge
* Group fire sensitive conifers
+ Abie lasi	
+ Abie bals
+ Pice glau
* Group Larix
+ Lari lari	
* Group Pice mari
+ Pice mari	
* Group deciduous
+ Betu papy	
+ Popu trem	
+ Popu balb

```{r summaryDataVeg, cache = FALSE}
# summarise weather data - faster if not cached
summaryABSK_AllData[, FireResConif := sum(`Pice enge`,
                                          `Pinu cont`, `Pinu bank`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, FireSenConif := sum(`Pice glau`,
                                          `Abie bals`, `Abie lasi`, na.rm = TRUE), 
                    by = pixID] 
summaryABSK_AllData[, Decid := sum(`Popu trem`, `Popu balb`, `Betu papy`, na.rm = TRUE),
                    by = pixID] 

## collapse non-forested variables
if (dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_VEG)])[1] |
    dim(summaryABSK_AllData[!is.na(NATURALLY_NON_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1] |
    dim(summaryABSK_AllData[!is.na(NON_FORESTED_VEG) & !is.na(NON_FORESTED_ANTHRO)])[1])
  warning("> 1 non-forest class type present")

collapseCols <- function(x) {
  x <- unlist(x)
  names(x) <- NULL
  if(all(is.na(x)))
    "NA"
  else 
    x[!is.na(x)]
}

cols <- c("NATURALLY_NON_VEG", "NON_FORESTED_VEG", "NON_FORESTED_ANTHRO")
summaryABSK_AllData[, (cols) := lapply(.SD, as.character), .SDcols = cols]
summaryABSK_AllData[, NonForestValue := collapseCols(c(.SD)), .SDcols = cols, by = pixID]
```

### COMMUNITY COMPOSITION/STRUCTURE PCAs - FOOTHILLS AND BOREAL SEPARATED

```{r FootVsBoreal_summaryData2_PCA_prep, cache = FALSE}
## select community structure/composition variables for PCA + severity + topo + some climate variables
cols <- grep("UNDER|SEV_CONT|FIRE_NAME|Conif|Decid|Lari|Pice mari|HEIGHT_LOWER|CROWN_CLOSURE_LOWER|SMR_ord|NR1|mean|cv", names(summaryABSK_AllData), value = TRUE)

summaryABSK_AllData2 <- unique(na.omit(summaryABSK_AllData[, ..cols]))
summaryABSK_AllData2[, UNDERSTOREY := as.factor(as.character(UNDERSTOREY))] ## convert to factor for dudi.mix
```

```{r Foot_summaryData2_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## FOOTHILLS + ROCKY MOUNTAIN
## exclude also P. banksiana, because it has no data
colsA <- grep("SEV_CONT|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONTmean|cv|NR1|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Foothills|Rocky", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData2_PCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
# unique(as.factor(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$FIRE_NAME)) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Foot_summaryData2_PCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR \nFoothills + Rocky Mountain")
points(PCAplot, "sites", pch = 19, 
       col = summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord)
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
sort(unique(summaryABSK_AllData2[grep("Foothills|Rocky", NR1)]$SMR_ord)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData2_PCA_plotClosure, fig.height=8, fig.show = "hold"}
## BOREAL FOREST + CANADIAN SHIELD
## exclude also P. banksiana, P. engelmann and A. lasiocarpa because they have no data
colsA <- grep("SEV_CONT|FIRE_NAME", cols, value = TRUE, invert = TRUE)
colsB <- grep("SEV_CONT|NR1|mean|cv|FIRE_NAME", cols, value = TRUE, invert = TRUE)
summDataPCA <- as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsA][, ..colsB]) %>%
  ade4::dudi.mix(.
                 , nf = 3
                 , scannf = FALSE)

colsC <- grep("SEV_CONT|mean|cv|NR1", cols, value = TRUE)
colsD <- grep("SEV_CONT|mean|cv", cols, value = TRUE)
env.fit <-  as.data.frame(summaryABSK_AllData2[grep("Boreal|Canadian", NR1), ..colsC][, ..colsD]) %>%
  envfit(summDataPCA, env = .)

# dev()
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by crown closure\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$CROWN_CLOSURE_LOWER)) %>%
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData2_PCA_plotFireName, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by fire name\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME))
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
# unique(as.factor(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$FIRE_NAME)) %>%
#   legend(x = "bottomright", legend = ., col = ., pch = 19)
```

```{r Boreal_summaryData2_PCA_plotSMR, fig.height=8, fig.show = "hold"}
PCAplot <- ordiplot(summDataPCA, type = "none", main = "Coloured by SMR\nBoreal Forest + Canadian Shield")
points(PCAplot, "sites", pch = 19, 
       col = summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord)
text(PCAplot, "species", labels = rep("", nrow(summDataPCA$co)), arrows = TRUE)
text(PCAplot, "species", arrows = FALSE)
plot(env.fit, labels = colsD)
sort(unique(summaryABSK_AllData2[grep("Boreal|Canadian", NR1)]$SMR_ord)) %>
  legend(x = "bottomright", legend = ., col = ., pch = 19)
```


### Spearman correlations collapsed species and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and ORIGIN replacing STAND_AGE

```{r vegTopoCorrSpearman2}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range|Pice [g|e]|Pinu|Abie [b|l]|Popu|Betu|NonForest",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
```

```{r vegTopoCorrSpearman2Print, results="asis"}
print(xtable(tabl), type = "html")
```

### Spearman correlations between severity and collapsed species and topo covariates only
CROWN_CLOSURE_LOWER, HEIGHT_LOWER and ORIGIN replacing STAND_AGE
Correlation values are worse becase some groups are collapsing species that had opposite correlation signs with severity.
```{r vegSevTopoCorrSpearman2}
cols <- grep("SEV|FIRE_|fireNum|Patch|Province|P_ID|gridID|Long|Lat|WETLAND|MAPSHEET|Duration|Area|Pre.|LAYER|PROD|DIST|SMR$|STAND|SITE|NON_|UNDERSTOREY|LOWER|pixID|Czone|^NR1|^NR2|startJulDay|^mean|^min|^max|^cv|^range|Pice [g|e]|Pinu|Abie [b|l]|Popu|Betu|NonForest",
             names(summaryABSK_AllData), invert = TRUE, value = TRUE)

cols  <- c(cols[!cols %in% "SEV_CONT"], "SEV_CONT")   ## reorder
tabl <- as.matrix(summaryABSK_AllData[, ..cols]) %>%
  prepCorrTable(., method = "spearman")
tabl <- tabl[rownames(tabl) == "SEV_CONT",]
tabl <- as.data.frame(t(tabl))

corrVals <- as.numeric(gsub("\\*", "", tabl[, "SEV_CONT"]))
tabl <- tabl[order(abs(corrVals), decreasing = TRUE),, drop = FALSE]
```

```{r vegSevTopoCorrSpearman2Print, results="asis"}
print(xtable(tabl), type = "html")
```


### Plots
#### Severity and collapsed vegetation covariates only

```{r vegSevPlots2, fig.height=8, fig.show = "hold", cache=FALSE}
cols <- grep("Pice mari|Lari|Decid|Conif|SEV_CONT", names(summaryABSK_AllData), value = TRUE)
tablePlot <- melt(summaryABSK_AllData[, ..cols], id.vars = "SEV_CONT")
tablePlot[value == "TRUE", value := "1"]
tablePlot[value == "FALSE", value := "0"]
tablePlot[, value := as.numeric(value)]

plot1 <- ggplot(tablePlot, aes(x = value, y = SEV_CONT, group = value)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free")

plot2 <- ggplot(summaryABSK_AllData, aes(x = NonForestValue, y = SEV_CONT)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("FA" = "Infrastruc.",
                              "HG" = "Grassland (gram.)",
                              "IN" = "Industry", 
                              "RI" = "River", 
                              "SL" = "Low shrub",
                              "ST" = "Tall shrub",
                              "NA" = "Forested"))
```

Note that only 7 pixels have "Infrascturure", "River" or "Industry", and only 3230/417985 have a non-NA non forested class.

```{r vegSevPlots2Print, fig.height=8, fig.show = "hold"}
plot1
plot2
```
