---
title: "Dave Andison's fire data Severity models"
author: "Ceres Barros"
date: "02/04/2019"
date-modified: last-modified
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(getwd()))
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)
```

```{r setup2, include=FALSE}
if (!exists("pkgDir")) {
  pkgDir <- file.path(
    if (Sys.info()[["user"]] == "rstudio") "packages_docker" else "packages",
    version$platform,
    paste0(version$major, ".", strsplit(version$minor, "[.]")[[1]][1])
  )
  
  if (!dir.exists(pkgDir)) {
    dir.create(pkgDir, recursive = TRUE)
  }
  .libPaths(pkgDir)
}

install.packages("Require")

## no need to load these.
Require::Install(c(
  "carrier" ## for in_parallel with purrr::*
  "future",
  "future.apply",
  "mirai",   ## for in_parallel with purrr::*
))

Require::Require(
  c(
    "crayon",
    "data.table",
    "reshape2",
    "sf",
    "dplyr",
    "xlsx",
    "foreign",
    "fasterize",
    "prioritizr",
    "gamlss.inf",
    "gamlss.spatial", ## "CeresBarros/gamlss.spatial (==3.0-3.9000)"
    "gamlss",
    "ggplot2",
    "ggpubr",
    "ggrepel",
    "LandR.CS",
    "caret",
    "purrr",
    "reproducible",   ## "PredictiveEcology/reproducible@AI (==2.1.2.9001)",
    "raster",
    "terra",
    "tidyterra",
    "ToolsCB"         ## "CeresBarros/ToolsCB (>=0.0.9001)"
  ), install = FALSE
)

Require::Require("reproducible", install = FALSE) ## for some reasons it's not being loaded above
doCache = TRUE

figDir <- "analyses/Figs"
dir.create(figDir, showWarnings = FALSE)

options("reproducible.useNewDigestAlgorithm" = TRUE,
        "reproducible.cachePath" = "analyses/cache",
        "reproducible.useMemoise" = FALSE,
        "future.globals.maxSize" = +Inf)

source("R/R_tools/prepCorrTable.R")
source("R/R_tools/Useful_functions.R") ## may replace some functions in ToolsCB, this is ok
source("R/R_tools/Neighbourhood_functions.R") 
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("analyses/R_tools/summarizeABSK_AllData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE

figDir <- "analyses/Figs"
```

```{r dataprep, include = FALSE, cache = FALSE}
resolution = 30

summaryABSK_AllData <- Cache(
  dataPrepWrapper,
  resolution = resolution, 
  fireDataPath = "data/fires_Dave/fireSev",
  vegDataPath = "data/fires_Dave/prefireVeg",
  topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
  weatherDataPath = "data/fires_Dave/fireWeather",
  doCache = doCache,
  userTags = "summaryABSK_AllData",
  cacheId = "185c1b8db3c9788d",
  omitArgs = c("useCache", "doCache", "userTags"))

## /!\ the following variables have many NA's and should perhaps be excluded
## wetlands are indicated within `LandCover`, the wetland vars are too much for this analysis.
## "WETLAND_CLASS", "WETLAND_VEG_MOD", "WETLAND_LAND_MOD", "WETLAND_LOCAL_MOD"
## "Duration"
## "cvRain
## SMR_ord/wet -- probably nothing that can be done here and there are few NAs on this one.
```

```{r addNgbNoBurns, cache = FALSE, include = FALSE}
## make different buffer sizes around points
crsProj <- crs(vect("data/fires_Dave/fireSev/albertafires1_postfire.shp"), proj = TRUE)

SEV_PROPpoints <- summaryABSK_AllData[, .(pixID, FIRE_NAME, SEV_PROP, Lat, Long)]
SEV_PROPpoints <- st_as_sf(SEV_PROPpoints, coords = c("Long", "Lat"),
                           agr = "constant", crs = st_crs(crsProj))

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i - 1] * 2
}

ngbNoBurns <- reproducible::Cache(calculateNgbBurnsWrapper,
                                  dists = bufferDists,
                                  sevPoints = SEV_PROPpoints, 
                                  sevColID = "SEV_PROP",
                                  fireColID = "FIRE_NAME",
                                  resolution = resolution,
                                  parallel = TRUE,
                                  cores = 3,
                                  useCache = doCache,
                                  cacheId = "c7536dcb493ad453",
                                  userTags = "ngbBurns",
                                  omitArgs = c("useCache", "userTags", "cores")) 
ngbNoBurns <- dcast.data.table(ngbNoBurns, ... ~ bufferSize, value.var = "ngbPropBurns")

cols <- grep("pixID|FIRE_NAME", names(ngbNoBurns), invert = TRUE, value = TRUE)
setnames(ngbNoBurns, old = cols, new = paste0("buffer_", cols, "m"))

summaryABSK_AllData <- ngbNoBurns[summaryABSK_AllData, on = .(pixID, FIRE_NAME)]

rm(SEV_PROPpoints); gc(reset = TRUE)
```

```{r addNgbAvgProperties, cache = FALSE, include = FALSE}
## get projection
crsProj <- crs(vect("data/fires_Dave/fireSev/albertafires1_postfire.shp"), proj = TRUE)

## list of variables retained in model without spatial covariates (model8.4FULL in DAfires_sevModelsGAMLSS.Rmd)
vars <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "UNDERSTOREY", "FlamConif", "PctSlope", "SurfAspectRatio",
          "isForest")

## convert logical to binary. averages will be the proportion of 1s
dataPoints <- summaryABSK_AllData[, .SD, .SDcols = c("pixID", "FIRE_NAME", "Lat", "Long", vars)]
cols <- which(unlist(as.vector(dataPoints[, lapply(.SD, is.logical)])))
dataPoints <- dataPoints[, (cols) := lapply(.SD, as.integer), .SDcols = cols]
dataPoints <- st_as_sf(dataPoints, coords = c("Long", "Lat"),
                       agr = "constant", crs = st_crs(crsProj))

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i - 1] * 2
}

## note that for binary variables mean = proportion of 1s
ngbAvgs <- reproducible::Cache(calculateNgbAvgsWrapper,
                               dists = bufferDists,
                               dataPoints = dataPoints,
                               pointIDColID = "pixID",
                               fireColID = "FIRE_NAME",
                               resolution = resolution,
                               parallel = TRUE,
                               cores = 10,
                               # useCache = doCache,
                               useCache = FALSE,
                               # cacheId = "",
                               userTags = "ngbAvg",
                               omitArgs = c("useCache", "userTags", "cores", "parallel"))

# Only add the variables that are pertinent, or the table will become immense
# summaryABSK_AllData <- ngbNoBurns[summaryABSK_AllData, on = .(pixID, FIRE_NAME)]

rm(dataPoints); gc(reset = TRUE)

```



## Include neighbours
### Proportion of burnt neighbours
Start from last best model, and find the best buffer sizes to add to each parameter.
```{r burnNgbPropModels, cache = FALSE, eval=FALSE}
model8.4 <- readRDS("analyses/model8_4.rds")

upperFormula <- as.formula(paste(
  paste(as.character(model8.4$mu.formula)[c(1,3)], collapse = " "),
  "+",
  paste(grep("buffer", names(modelData), value = TRUE), collapse = " + ")))

modelNgbSevMu <- add1(model8.4, what = "mu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

modelNgbSev1 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + buffer_960m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev1$sigma.formula), collapse = " "),
  "+",
  paste(grep("buffer", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevSigma <- add1(modelNgbSev1, what = "sigma",
                         scope = upperFormula,
                         k = log(length(samp)),
                         parallel = "multicore", ncpus = 10)

modelNgbSev2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + buffer_960m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev2$nu.formula), collapse = " "),
  "+",
  paste(grep("buffer", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevNu <- add1(modelNgbSev2, what = "nu",
                      scope = upperFormula,
                      k = log(length(samp)),
                      parallel = "multicore", ncpus = 10)

modelNgbSev3 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + buffer_960m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet + buffer_30m +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI + SurfReliefRatio + SMR_wet +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))

upperFormula <- as.formula(paste(
  paste(as.character(modelNgbSev3$tau.formula), collapse = " "),
  "+",
  paste(grep("buffer", names(modelData), value = TRUE), collapse = " + ")))
modelNgbSevTau <- add1(modelNgbSev3, what = "tau",
                       scope = upperFormula,
                       k = log(length(samp)),
                       parallel = "multicore", ncpus = 10)
modelNgbSev4 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                         CTI + Lari + SMR_wet + buffer_960m +
                         re(random = ~1 | FIRE_NAME),
                       sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                         re(random = ~1 | FIRE_NAME),
                       nu.formula = ~FlamConif +
                         Decid + UNDERSTOREY + PctSlope * SurfAspectRatio +
                         PctSlope * SurfReliefRatio + SMR_wet + buffer_30m +
                         re(random = ~1 | FIRE_NAME),
                       tau.formula = ~isForest + isForest *
                         UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                         Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                         SurfReliefRatio + SMR_wet + buffer_30m +
                         re(random = ~1 | FIRE_NAME), family = BEINF,
                       data = na.omit(modelData), control = gamlss.control(c.crit = 0.1))
```

### Average predictors values within buffer rings
Start from last best model without neighbours, and find the best buffer rings and average covarates to add to each parameter.

#### HERE!!!!
ASVASV

#### fit model to full dataset (and simplify if need be)
```{r FullburnNgbPropModel, cache = FALSE, eval=FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")

modelNgbSev4FULL <- update(modelNgbSev4, data = na.omit(summaryABSK_AllData[, cols, with = FALSE]))
Rsq(modelNgbSev4FULL)
summary(modelNgbSev4FULL)

## remove non-significant effects:
modelNgbSev4FULL2 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                              CTI + Lari + SMR_wet + buffer_960m +
                              re(random = ~1 | FIRE_NAME),
                            sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                              re(random = ~1 | FIRE_NAME),
                            nu.formula = ~FlamConif + UNDERSTOREY + SurfReliefRatio +
                              PctSlope + SMR_wet + buffer_30m +
                              re(random = ~1 | FIRE_NAME),
                            tau.formula = ~isForest + isForest *
                              UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                              Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                              SurfReliefRatio + SMR_wet + buffer_30m +
                              re(random = ~1 | FIRE_NAME), family = BEINF,
                            data = na.omit(summaryABSK_AllData[, cols, with = FALSE]), 
                            control = gamlss.control(c.crit = 0.1))

summary(modelNgbSev4FULL2)

## add level = 0 to get marginal means.
modelNgbSev4FULL3 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                              CTI + Lari + SMR_wet + buffer_960m +
                              re(random = ~1 | FIRE_NAME, level = 0),
                            sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                              re(random = ~1 | FIRE_NAME, level = 0),
                            nu.formula = ~FlamConif + SurfReliefRatio + SMR_wet + buffer_30m +
                              re(random = ~1 | FIRE_NAME, level = 0),
                            tau.formula = ~isForest + isForest *
                              UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                              Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                              SurfReliefRatio + SMR_wet + buffer_30m +
                              re(random = ~1 | FIRE_NAME, level = 0), family = BEINF,
                            data = na.omit(summaryABSK_AllData[, cols, with = FALSE]), 
                            control = gamlss.control(c.crit = 0.1))

## try the above with Ian's custom function (given to us by Dr Stasinopolous)
modelNgbSev4FULL4 <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                              CTI + Lari + SMR_wet + buffer_960m +
                              re(random = ~1|FIRE_NAME, level = 0),
                            sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                              re(random = ~1|FIRE_NAME, level = 0),
                            nu.formula = ~FlamConif + SurfReliefRatio + SMR_wet + buffer_30m +
                              re(random = ~1|FIRE_NAME, level = 0),
                            tau.formula = ~isForest + isForest *
                              UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                              Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                              SurfReliefRatio + SMR_wet + buffer_30m +
                              re(random = ~1|FIRE_NAME, level = 0), family = BEINF,
                            data = na.omit(summaryABSK_AllData[, cols, with = FALSE]), 
                            control = gamlss.control(c.crit = 0.1))

sink("analyses/modelNgbSev4FULL4_SUMMARY.txt")
summary(modelNgbSev4FULL4)
sink()

tiff("analyses/modelNgbSev4FULL4_RESIDS.tif")
plot(modelNgbSev4FULL4)
graphics.off()

saveRDS(modelNgbSev4FULL4, "analyses/modelNgbSev4FULL4.rds")

GAIC(modelNgbSev4FULL,
     modelNgbSev4FULL2,
     modelNgbSev4FULL3,
     modelNgbSev4FULL4,
     k = log(nrow(na.omit(summaryABSK_AllData[, cols, with = FALSE]))))
#                         df      AIC
# modelNgbSev4FULL3 172.6552 486027.6
# modelNgbSev4FULL2 174.6581 486048.1
# modelNgbSev4FULL  178.6594 486093.9
```

#### fit model to full dataset with scaled covariates
```{r FullburnNgbPropModelScaledPreds, cache = FALSE, eval = FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "SurfReliefRatio", "isForest")
modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

modelNgbSev4FULL4_Scaled <- gamlss(formula = SEV_PROP ~ Decid * HEIGHT_LOWER +
                                     CTI + Lari + SMR_wet + buffer_960m +
                                     re(random = ~1|FIRE_NAME, level = 0),
                                   sigma.formula = ~Decid * UNDERSTOREY + SMR_wet + buffer_1920m +
                                     re(random = ~1|FIRE_NAME, level = 0),
                                   nu.formula = ~FlamConif + SurfReliefRatio + SMR_wet + buffer_30m +
                                     re(random = ~1|FIRE_NAME, level = 0),
                                   tau.formula = ~isForest + isForest *
                                     UNDERSTOREY + Decid * HEIGHT_LOWER + Decid * UNDERSTOREY +
                                     Lari * HEIGHT_LOWER + Lari * UNDERSTOREY + CTI +
                                     SurfReliefRatio + SMR_wet + buffer_30m +
                                     re(random = ~1|FIRE_NAME, level = 0), family = BEINF,
                                   data = na.omit(modelDataFULLScaled), 
                                   control = gamlss.control(c.crit = 0.1))
GAIC(modelNgbSev4FULL4,
     modelNgbSev4FULL4_Scaled,
     k = log(nrow(na.omit(modelDataFULLScaled))))

sink("analyses/modelNgbSev4FULL4Scaled_SUMMARY.txt")
summary(modelNgbSev4FULL4_Scaled)
sink()

tiff("analyses/modelNgbSev4FULL4Scaled_RESIDS.tif")
plot(modelNgbSev4FULL4_Scaled)
graphics.off()

saveRDS(modelNgbSev4FULL4_Scaled, "analyses/modelNgbSev4FULL4Scaled.rds")

```

#### confusion matrix with full model
```{r confMatrixFullModelScaled, cache = FALSE, eval = FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "SurfReliefRatio", "isForest", "pixID",
          "SEV_CLASS")

modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")

predictionsDT <- na.omit(modelDataFULLScaled)[, .(SEV_PROP, SEV_CLASS)]
predData <- predictAll(modelNgbSev4FULL4_Scaled, type = "response")
predictionsDT[, predSEV_PROP := calcMeanBEINF(predData$mu, predData$nu, predData$tau)]

rm(predData)
## convert to classes, using the quantiles corresponding to the observed class proportions
## accumulate proportions to get probabilities
quantProbs <- cumsum(table(predictionsDT$SEV_CLASS)/nrow(predictionsDT))
classRanges <- c(0, quantile(predictionsDT$predSEV_PROP, probs = quantProbs))

predictionsDT[, predSEV_CLASS := cut(predSEV_PROP, breaks = classRanges,
                                     include.lowest = TRUE, right = FALSE)]

## convert to numbered factor (subtracting one, because classes are 0-5)
predictionsDT[, predSEV_CLASS := as.numeric(predSEV_CLASS)-1]
classes <- as.character(sort(unique(predictionsDT$SEV_CLASS)))
predictionsDT[, `:=`(SEV_CLASS = factor(SEV_CLASS, levels = classes),
                     predSEV_CLASS = factor(predSEV_CLASS, levels = classes))]

confMatrix <- caret::confusionMatrix(data = predictionsDT$predSEV_CLASS, reference = predictionsDT$SEV_CLASS)

sink("analyses/modelNgbSev4FULL4Scaled_CONFMATRX.txt")
confMatrix
sink()

tiff("analyses/modelNgbSev4FULL4Scaled_obsVSpred.tif", width = 7, height = 5, units = "in", res = 300)
ggplot(predictionsDT, aes(x = SEV_PROP, y = predSEV_PROP)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", colour = "gray30") +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(0,1)) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  annotate("text", x = 0.1, y = 0.9, size = 5,
           label = paste("r =", round(cor.test(predictionsDT$SEV_PROP, predictionsDT$predSEV_PROP)$estimate, 2))) +
  labs(x = "Observed severity", y = "Expected mean severity")
graphics.off()
```


#### Model validation

##### On `modelNgbSev4FULL4` 

Using k-fold cross validation on `modelNgbSev4FULL4` (which was fitted on the full data), with the data partitioned in 4 chunks.
```{r crossValidationBurnNgbProp, message=FALSE, eval=FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
modelNgbSev4FULL4 <- readRDS("analyses/modelNgbSev4FULL4.rds")

crossValidResults3 <- Cache(crossValidFunction,
                            fullDT = summaryABSK_AllData,
                            origData = na.omit(summaryABSK_AllData[, cols, with = FALSE]),
                            statsModel = modelNgbSev4FULL4,
                            k = 10, idCol = "pixID",
                            classVar = "SEV_CLASS",
                            level = 0, 
                            cacheObj1 = summaryABSK_AllData$pixID,
                            cacheObj2 = na.omit(summaryABSK_AllData[, cols, with = FALSE])$pixID,
                            parallel = FALSE,
                            workers = 2,
                            # useCache = "overwrite",   ## cacheId = 85b6d427b43ff53d
                            # showSimilar = TRUE,
                            userTags = c("validation", "modelNgbSev4FULL4"),
                            omitArgs = c("fullDT", "origData", "userTags", "parallel", "workers"))

# crossValidResults3 <- loadFromCache(cacheId="85b6d427b43ff53d")
## collapse results into data.tables
allValidMetrics3 <- rbindlist(lapply(crossValidResults3, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics3[, sampID := 1:length(crossValidResults3)]

allConfMatrices3 <- rbindlist(lapply(1:length(crossValidResults3), FUN = function(x) {
  DT <- as.data.table(crossValidResults3[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))
## add identifier
allConfMatrices3[, compareID := paste(Prediction, Reference, sep = "_")]

## average confusion matrix 
avgAllConfMatrices3 <- allConfMatrices3[, list(avgN = mean(N)), by = compareID]
avgAllConfMatrices3 <- unique(allConfMatrices3[, .(Prediction, Reference, compareID)])[avgAllConfMatrices3, on = "compareID"]
avgAllConfMatrices3[, compareID := NULL]
avgAllConfMatrices3 <- dcast(avgAllConfMatrices3, Prediction ~ Reference)

allRsquared3 <- rbindlist(sapply(crossValidResults3, FUN = function(x) as.data.table(t(x$Rsquared)),
                                 simplify = FALSE), idcol = "sampID")

allRsquared3GAMLSS <- rbindlist(sapply(crossValidResults3, FUN = function(x) as.data.table(t(x$RsqGAMLSS)),
                                       simplify = FALSE), idcol = "sampID")

allTGD <- rbindlist(sapply(crossValidResults3, FUN = function(x) as.data.table(t(x$TGD)),
                           simplify = FALSE), idcol = "sampID")

allCoefs <- rbindlist(sapply(crossValidResults3, FUN = function(x) {
  mu <- as.data.table(t(x$coefs$mu))
  sigma <- as.data.table(t(x$coefs$sigma))
  nu <- as.data.table(t(x$coefs$nu))
  tau <- as.data.table(t(x$coefs$tau))
  coefsDT <- data.table(rbindlist(list(mu, sigma, nu, tau), fill = TRUE), 
                        param = c("mu", "sigma", "nu", "tau"))
  coefsDT
},
simplify = FALSE), idcol = "sampID")

## save
sink("analyses/modelNgbSev4FULL4_crossValidOutputs.txt")
print("Validation metrics")
allValidMetrics3
print("******************")
print("Averaged confusion matrices")
avgAllConfMatrices3
print("******************")
print("R squared")
allRsquared3
print("******************")
print("R squared GAMLSS")
allRsquared3GAMLSS
print("******************")
print("TGD")
allTGD
print("******************")
print("Coefficients mean")
allCoefs[, lapply(.SD, mean), by = .(param),
         .SDcols = setdiff(names(allCoefs), c("sampID", "param"))]
print("Coefficients SD")
allCoefs[, lapply(.SD, mean), by = .(param),
         .SDcols = setdiff(names(allCoefs), c("sampID", "param"))]
sink()
```

##### On `modelNgbSev4FULL4_Scaled` with standardized predictors 
```{r crossValidationBurnNgbPropScaled, message=FALSE, eval=FALSE}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest", "pixID",
          "SEV_CLASS")

modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "PctSlope", "SurfAspectRatio", "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")

crossValidResults4 <- Cache(crossValidFunction,
                            fullDT = na.omit(modelDataFULLScaled),
                            origData = na.omit(modelDataFULLScaled),
                            statsModel = modelNgbSev4FULL4_Scaled,
                            k = 10, 
                            classVar = "SEV_CLASS",
                            sampleGroup = "FIRE_NAME",
                            idCol = "pixID",
                            level = 0, 
                            spec = 2,
                            parallel = FALSE,
                            .cacheExtra = list(residuals(modelNgbSev4FULL4_Scaled),
                                               a = modelDataFULLScaled$pixID,
                                               b = na.omit(modelDataFULLScaled)$pixID),
                            # useCache = "overwrite", 
                            # showSimilar = TRUE,
                            userTags = c("validation", "modelNgbSev4FULL4_Scaled"),
                            omitArgs = c("fullDT", "origData", "statsModel",
                                         "userTags", "parallel", "workers"))
rm(trainData); gc(reset = TRUE)
# crossValidResults4 <- loadFromCache(cacheId="85b6d427b43ff53d")
## collapse results into data.tables
allValidMetrics4 <- rbindlist(lapply(crossValidResults4, FUN = function(x) as.data.table(t(x$validMetrics))))
allValidMetrics4[, sampID := 1:length(crossValidResults4)]

allConfMatrices4 <- rbindlist(lapply(1:length(crossValidResults4), FUN = function(x) {
  DT <- as.data.table(crossValidResults4[[x]]$confMatrix$table)
  DT[, sampID := x]
  DT
}
))
## add identifier
allConfMatrices4[, compareID := paste(Prediction, Reference, sep = "_")]

## average confusion matrix 
avgAllConfMatrices4 <- allConfMatrices4[, list(avgN = mean(N)), by = compareID]
avgAllConfMatrices4 <- unique(allConfMatrices4[, .(Prediction, Reference, compareID)])[avgAllConfMatrices4, on = "compareID"]
avgAllConfMatrices4[, compareID := NULL]
avgAllConfMatrices4 <- dcast(avgAllConfMatrices4, Prediction ~ Reference)

allCoefs <- rbindlist(sapply(crossValidResults4, FUN = function(x) {
  mu <- as.data.table(t(x$coefs$mu))
  sigma <- as.data.table(t(x$coefs$sigma))
  nu <- as.data.table(t(x$coefs$nu))
  tau <- as.data.table(t(x$coefs$tau))
  coefsDT <- data.table(rbindlist(list(mu, sigma, nu, tau), fill = TRUE), 
                        param = c("mu", "sigma", "nu", "tau"))
  coefsDT
},
simplify = FALSE), idcol = "sampID")

## save
sink("analyses/modelNgbSev4FULL4Scaled_crossValidOutputs.txt")
print("Validation metrics")
allValidMetrics4
print("******************")
print("Averaged confusion matrices")
avgAllConfMatrices4
print("******************")
print("Coefficients mean")
allCoefs[, lapply(.SD, mean), by = .(param),
         .SDcols = setdiff(names(allCoefs), c("sampID", "param"))]
print("Coefficients SD")
allCoefs[, lapply(.SD, mean), by = .(param),
         .SDcols = setdiff(names(allCoefs), c("sampID", "param"))]
sink()
```

Plot cross-validation results against model metrics
```{r crossValidPlots_ScaledModel}
allRMSE4 <- rbindlist(sapply(crossValidResults4, FUN = function(x) {
  as.data.table(x$validMetrics[["RMSE"]])
},
simplify = FALSE), idcol = "sampID")

allRsquared4GAMLSS <- rbindlist(sapply(crossValidResults4, FUN = function(x) {
  as.data.table(x$validMetrics[["RsqGAMLSS"]])
},
simplify = FALSE), idcol = "sampID")

allTGD4 <- rbindlist(sapply(crossValidResults4, FUN = function(x) {
  as.data.table(x$validMetrics[["TGD"]])
},
simplify = FALSE), idcol = "sampID")

allRsquared4GAMLSS[, fullModel := Rsq(modelNgbSev4FULL4_Scaled)]

rmseDT <- data.table(pred = calcMeanBEINF(mu = predict(modelNgbSev4FULL4_Scaled, type = "response", what = "mu"),
                                          nu = predict(modelNgbSev4FULL4_Scaled, type = "response", what = "nu"),
                                          tau = predict(modelNgbSev4FULL4_Scaled, type = "response", what = "tau")),
                     obs = modelNgbSev4FULL4_Scaled$y)
allRMSE4[, fullModel := RMSE(rmseDT$pred, rmseDT$obs)]
## NOTE: don't compare TGD between full model and folds. get full deviance from model object
allTGD4[, fullModel := modelNgbSev4FULL4_Scaled$G.deviance]

plotData <- rbindlist(list(
  Rsq = allRsquared4GAMLSS
  , TGD = allTGD4
  # , RMSE = allRMSE4
),
idcol = "metric", fill = TRUE)

crossValidplot <- ggplot(plotData, aes(y = V1, x = as.factor(metric))) +
  geom_boxplot() +
  geom_point(aes(y = fullModel),
             position = position_dodge(width = 0), size = 4, 
             colour = "darkred", show.legend = FALSE) +
  theme_pubr(base_size = 16) +
  theme(strip.text = element_blank()) +
  scale_x_discrete(labels = c("Rsq" = "R-squared", "TGD" = "Global deviance")) +
  labs(y = "value", x = "") +
  facet_wrap(~ metric, scales = "free")

png(file.path(figDir, "crossValidPlot_DAfires_sevModels.png"), height = 5, width = 8,
    units = "in", res = 300)
crossValidplot
dev.off()
```

## Inspecting the final model and figures
Probabilities of 0s and 1s in function of covariates:
$$xi0 = \nu = \frac{p0}{1 + p0 + p1}$$
$$xi1 = \tau = \frac{p1}{1 + p0 + p1}$$
where *xi0* and *xi1* are the `gamlssInf0to1` equivalents of $\nu$ and $\tau$.
See above for variance (Var(Y)) of the zero- and one-inflated Beta regression.

### General effects - model with scaled variables.
Only significant covariates are plotted
```{r modelResults, include = FALSE, eval=FALSE}
modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")
## plot the probabilities at 0 and 1
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "SurfReliefRatio", "isForest", "pixID",
          "SEV_CLASS")

modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

## need to provide "newdata" to get marginal means
plotData <-  na.omit(modelDataFULLScaled)
predData <- predictAll(modelNgbSev4FULL4_Scaled, type = "response", newdata = plotData[, cols, with = FALSE])

plotData[, `:=`(muPred = predData$mu, 
                sigmaPred = predData$sigma, 
                nuPred = predData$nu,
                tauPred = predData$tau)]
plotData[ , `:=`(p0 = calcP0BEINF(nuPred, tauPred),
                 p1 = calcP1BEINF(nuPred, tauPred),
                 fittedVals = calcMeanBEINF(muPred, nuPred, tauPred),
                 varY = calcVarBEINF(muPred, sigmaPred, nuPred, tauPred))]
rm(predData)
```

#### Plots of probabilities at 0 and 1 vs. covariates
Note that because $\nu$ and $\tau$ influence both *p0* and *p1* we need to plot them against all $\nu$ and $\tau$ covariates.
```{r p0p1Plots}
## 0 probs - nu covariates
p0plot1 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p0)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p0 ~ FlamConif\n GAM smoother",
       x = "% cover")

p0plot2 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p0)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ roughness (SRR)\n GAM smoother",
       x = "% roughness")

p0plot3 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (SMR)\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

p0plot4 <- ggplot(plotData, 
                  aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ neigh. burns (30m)\n GAM smoother",
       x = "prop. neigh. burnt")

## 0 probs - tau covariates
p0plot5 <- ggplot(plotData, aes(x = as.integer(isForest), y = p0,
                                linetype = as.factor(UNDERSTOREY))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(breaks = c(0,1), labels = c("0" = "not forest", "1" = "forest")) +
  theme_pubr() + 
  labs(title = "p0 ~ isForest * understorey\n(GLM logit smoother)",
       x = "", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData[, HEIGHT_LOWER2 := cut(HEIGHT_LOWER, 
                                breaks = c(0, 5, 10, 20, max(plotData$HEIGHT_LOWER)))]
p0plot6 <- ggplot(plotData, 
                  aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p0,
                      linetype = as.factor(UNDERSTOREY),
                      colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Broadleaf * Height\n GAM smoother",
       x = "% cover", colour = "height", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


p0plot7 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p0,
                                linetype = as.factor(UNDERSTOREY),
                                colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Lari * Height\n GAM smoother",
       x = "% cover", colour = "height", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


p0plot8 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (CTI)\n GAM smoother",
       x = "CTI")

png(file.path(figDir, "p0Plots_DAfires_sevModels.png"), height = 15, width = 15,
    units = "in", res = 300)
ggarrange(p0plot1, p0plot2, p0plot3, 
          p0plot4, p0plot5, p0plot6,
          p0plot7, p0plot8,
          font.label = list(size = 16),
          nrow = 3, ncol = 3)
dev.off()

## 1 probs - tau covariates
p1plot1 <- ggplot(plotData, aes(x = as.integer(isForest), y = p1,
                                linetype = as.factor(UNDERSTOREY))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(breaks = c(0,1), labels = c("0" = "not forest", "1" = "forest")) +
  theme_pubr() + 
  labs(title = "p1 ~ isForest * understorey\n(GLM logit smoother)",
       x = "", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

p1plot2 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p1,
                                linetype = as.factor(UNDERSTOREY),
                                colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Broadleaf * Height\n GAM smoother",
       x = "% cover", colour = "height", linetype = "understorey")  +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


p1plot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p1,
                                linetype = as.factor(UNDERSTOREY),
                                colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Lari * Height\n GAM smoother",
       x = "% cover", colour = "height", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


p1plot4 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (CTI)\n GAM smoother",
       x = "CTI")

p1plot5 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ roughness (SRR)\n GAM smoother",
       x = "% roughness")

p1plot6 <- ggplot(plotData, 
                  aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ neigh. burns (30m)\n GAM smoother",
       x = "prop. neigh. burnt")

p1plot7 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = p1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (SMR)\n(GLM logit smoother)",
       x = "soil moisture (SMR)")

## 1 probs - nu covariates
p1plot8  <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p1)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p1 ~ FlamConif\n GAM smoother",
       x = "% cover")

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "p1Plots_DAfires_sevModels.png"), height = 15, width = 15,
    units = "in", res = 300)
ggarrange(
  p1plot2, #Decid
  p1plot8, #Flamconif
  p1plot3, #Lari
  p1plot7, #SMR
  p1plot1, #isForest
  p1plot4, #CTI
  p1plot5, #SRR
  p1plot6, #buffer_30m
  font.label = list(size = 16),
  nrow = 3, ncol = 3)
dev.off()
```

#### Plot of the expected mean vs covariates
Note that because $\mu$, $\nu$ and $\tau$ influence the expected mean, it has to be plot against all $\mu$, $\nu$ and $\tau$ covariates.
```{r fittedMeanPlots}
meanplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = fittedVals,
                                  colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height \n GAM smoother",
       x = "% cover", y = "mean sev.", colour = "height")

meanplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)\n GAM smoother",
       x = "CTI", y = "mean sev.")

meanplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari\n GAM smoother",
       x = "% cover", y = "mean sev.")

meanplot4 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)\n(GLM logit smoother)",
       x = "soil moisture (SMR)", y = "mean sev.")

meanplot5 <- ggplot(plotData, 
                    aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)\n GAM smoother",
       x = "prop. neigh. burnt", y = "mean sev.")

## effects via nu and/or tau:
meanplot6 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "Severity ~ FlamConif\n GAM smoother",
       x = "% cover", y = "mean sev.")

meanplot7 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ roughness (SRR)\n GAM smoother",
       x = "% roughness", y = "mean sev.")

meanplot8 <- ggplot(plotData, 
                    aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = fittedVals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (30m)\n GAM smoother",
       x = "prop. neigh. burnt", y = "mean sev.")

meanplot9 <- ggplot(plotData, aes(x = as.integer(isForest), y = fittedVals,
                                  linetype = as.factor(UNDERSTOREY))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(breaks = c(0,1), labels = c("0" = "not forest", "1" = "forest")) +
  theme_pubr() + 
  labs(title = "Severity ~ isForest * understorey\n(GLM logit smoother)",
       x = "", linetype = "understorey", y = "mean sev.") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "meanSevPlots_DAfires_sevModels.png"), height = 15, width = 15,
    units = "in", res = 300)
ggarrange(
  meanplot1, #Decid
  meanplot6, #Flamconif
  meanplot3, #Lari
  meanplot4, #SMR
  meanplot9, #isForest
  meanplot2, #CTI
  meanplot7, #SRR 
  meanplot8, #buffer_30m
  meanplot5, #buffer_960m
  nrow = 3, ncol = 3, 
  font.label = list(size = 16))
dev.off()
```

#### Plot of fitted mean (mu) vs covariates
```{r fittedMuPlots}
muplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = muPred,
                                colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height \n GAM smoother",
       x = "% cover", y = expression(mu), colour = "height")

muplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = muPred)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)\n GAM smoother",
       x = "CTI", y = expression(mu))

muplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = muPred)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari\n GAM smoother",
       x = "% cover", y = expression(mu))

muplot4 <- ggplot(plotData, aes(x = as.integer(SMR_wet), y = muPred)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)\n(GLM logit smoother)",
       x = "soil moisture (SMR)", y = expression(mu))

muplot5 <- ggplot(plotData, 
                  aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = muPred)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)\n GAM smoother",
       x = "prop. neigh. burnt", y = expression(mu))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "muSevPlots_DAfires_sevModels.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(muplot1, #Decid
          muplot3, #Lari
          muplot4, #SMR
          muplot2, #CTI
          muplot5, #buffer_960m
          nrow = 3, ncol = 2, 
          font.label = list(size = 16))
dev.off()
```

#### Plot of the expected variance vs covariates
Note that because all parameters influence the variance, we need to plot it against all model covariates.
```{r VarYPlots}
## effects via sigma
## this plot also includes height (an effect via mu)
varplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = varY,
                                 linetype = as.factor(UNDERSTOREY),
                                 colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Broadleaf * understorey\n GAM smoother",
       y = "Expected variance", 
       x = "% cover", colour = "height", 
       linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## SMR is also affects variance via nu
varplot2 <-  ggplot(plotData, aes(x = as.integer(SMR_wet), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (SMR)\n(GLM logit smoother)",
       y = "Expected variance", x = "soil moisture (SMR)")

varplot3 <- ggplot(plotData, 
                   aes(x = buffer_1920m * attr(modelDataFULLScaled$buffer_1920m, "scaled:scale") + attr(modelDataFULLScaled$buffer_1920m, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (1920m)\n GAM smoother",
       y = "Expected variance", x = "prop. neigh. burnt")

varplot4 <- ggplot(plotData, 
                   aes(x = buffer_1920m * attr(modelDataFULLScaled$buffer_1920m, "scaled:scale") + attr(modelDataFULLScaled$buffer_1920m, "scaled:center"), y = varY)) +
  geom_point(aes(colour = as.integer(FIRE_NAME)), alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  geom_text_repel(data = plotData[, list(mean(buffer_1920m), mean(varY)), by = "FIRE_NAME"],
                  mapping = aes(x = V1, y = V2, label = FIRE_NAME, group = FIRE_NAME),
                  point.padding = 0.5,
                  segment.color = "grey50", 
                  force = 50) +
  scale_color_distiller(palette = "Spectral") +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (1920m)\n GAM smoother",
       y = "Expected variance", x = "prop. neigh. burnt")

## effects via mu:
varplot5 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  # geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) + ## weird curve
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (CTI)\n GAM smoother",
       x = "CTI", y = "Expected variance")

## includes effects of understorey and eight via tau
varplot6 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = varY,
                                 linetype = as.factor(UNDERSTOREY),
                                 colour = HEIGHT_LOWER2)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Lari * Height\n GAM smoother",
       x = "Expected variance", colour = "height", linetype = "understorey")

varplot7 <- ggplot(plotData, 
                   aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (960m)\n GAM smoother",
       x = "prop. neigh. burnt", y = "Expected variance")

## effects via nu:
varplot8 <- ggplot(plotData, 
                   aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ FlamConif\n GAM smoother",
       x = "% cover", y = "Expected variance")

varplot9 <- ggplot(plotData, 
                   aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ roughness (SRR)\n GAM smoother",
       x = "% roughness", y = "Expected variance")

varplot10 <- ggplot(plotData, 
                    aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = varY)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4, bs = "cs")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (30m)\n GAM smoother",
       x = "prop. neigh. burnt", y = "Expected variance")

varplot11 <- ggplot(plotData, 
                    aes(x = as.integer(isForest), y = varY, 
                        linetype = as.factor(UNDERSTOREY))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  scale_x_continuous(breaks = c(0,1), labels = c("0" = "not forest", "1" = "forest")) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ isForest * understorey\n(GLM logit smoother)",
       x = "", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "varSevPlots_DAfires_sevModels.png"), height = 20, width = 15,
    units = "in", res = 300)
ggarrange(
  varplot1, #Decid
  varplot2, #SMR
  varplot3, #buffer_1920m
  varplot4, #buffer_1920m (with fireID)
  varplot5, #CTI
  varplot6, #Lari
  varplot7, #buffer_960
  varplot8, #Flamconif
  varplot9, #SRR
  varplot10, #buffer_30m
  varplot11, #isForest
  nrow = 4, ncol = 3,
  font.label = list(size = 16))
dev.off()
```

### Partial effects plots
```{r PEFwrapper, include = FALSE}
## wrapper function
PEFwrapper <- function(obj, DT, term, fixed.at = list()) {
  DT2 <- copy(DT)
  PEFFunctions <- getPEF.own(obj, parameter = "all", term = term, fixed.at = fixed.at,
                             type = "response", how = "mean", plot = FALSE, level = 0)
  
  DT2$muPred <- PEFFunctions$mu(DT2[[term]])
  DT2$sigmaPred <- PEFFunctions$sigma(DT2[[term]])
  DT2$nuPred <- PEFFunctions$nu(DT2[[term]])
  DT2$tauPred <- PEFFunctions$tau(DT2[[term]])
  DT2[, `:=`(fittedVals = calcMeanBEINF(muPred, nuPred, tauPred),
             p0 = calcP0BEINF(nuPred, tauPred),
             p1 = calcP1BEINF(nuPred, tauPred),
             varY = calcVarBEINF(muPred, sigmaPred, nuPred, tauPred))]
  return(DT2)
}
```

#### Partial effects on probabilities at 0 and 1 - forest pixels
Showing partial effects in forested pixels only, except when isForest was significant
Note that because $\nu$ and $\tau$ influence both *p0* and *p1* we need to plot them against all $\nu$ and $\tau$ covariates.
```{r p0p1PEFPlots}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
## 0 probs - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot1 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p0 ~ FlamConif",
       x = "% cover", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot2 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ roughness (SRR)",
       x = "% roughness", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot3 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = p0)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (SMR)",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot4 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## 0 probs - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot5 <- ggplot(plotData, aes(x = as.factor(isForest), y = p0)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ isForest",
       x = "", shape = "understorey")

vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p0PEFplot6 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Broadleaf * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p0PEFplot7 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Lari * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0PEFplot8 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (CTI)",
       x = "CTI", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


legPlot <- get_legend(p0PEFplot6 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

png(file.path(figDir, "p0PEFPlots_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
## override scale for clarity
## order follows veg, moisture, forest, topo and neighbourhood
ggarrange(
  p0PEFplot6 + scale_y_continuous(limits = c(0, 0.10)), #Decid
  p0PEFplot1 + scale_y_continuous(limits = c(0, 0.10)), #FlamConif
  p0PEFplot7 + scale_y_continuous(limits = c(0, 0.10)), #Lari
  p0PEFplot3 + scale_y_continuous(limits = c(0, 0.10)), #SMR
  p0PEFplot5 + scale_y_continuous(limits = c(0, 0.10)), #isForest
  p0PEFplot8 + scale_y_continuous(limits = c(0, 0.10)), #CTI
  p0PEFplot2 + scale_y_continuous(limits = c(0, 0.10)), #SurfReliefRatio
  p0PEFplot4, #buffer_30m
  as_ggplot(legPlot),
  nrow = 3, ncol = 3,
  font.label = list(size = 16),
  legend = "none"
)
dev.off()

## 1 probs - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot1 <- ggplot(plotData, aes(x = as.factor(isForest), y = p1)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ isForest",
       x = "", linetype = "understorey") +
  guides(shape = guide_legend(override.aes = list(colour = "black")))

vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p1PEFplot2 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Broadleaf * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p1PEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Lari * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot4 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (CTI)",
       x = "CTI", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot5 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ roughness (SRR)",
       x = "% roughness", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot6 <- ggplot(plotData, 
                     aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot7 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = p1)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (SMR)",
       x = "", shape = "understorey")


## 1 probs - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1PEFplot8 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p1 ~ FlamConif",
       x = "% cover", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

legPlot <- get_legend(p1PEFplot2 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

png(file.path(figDir, "p1PEFPlots_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
## override scale for clarity
## order follows veg, moisture, forest, topo and neighbourhood
ggarrange(
  p1PEFplot2, #decid
  p1PEFplot8, #flamconif 
  p1PEFplot3, #lari
  p1PEFplot7, #SMR
  p1PEFplot1, #isForest
  p1PEFplot4, #CTI
  p1PEFplot5, #SRR
  p1PEFplot6, #buffer_30m
  as_ggplot(legPlot),
  nrow = 3, ncol = 3,
  font.label = list(size = 16),
  legend = "none")
dev.off()
```

#### Partial effects on expected mean - forest pixels
Note that the "expected mean" includes the effects of 0s and 1s
Showing partial effects in forested pixels only, except when isForest was significant
Note that because $\mu$, $\nu$ and $\tau$ influence the expected mean, it has to be plot against all $\mu$, $\nu$ and $\tau$ covariates.
```{r fittedMeanPEFPPlots}
## mean - mu covariates
vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

meanPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height",
       x = "% cover", y = "mean sev.", 
       linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)",
       x = "CTI", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

meanPEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari",
       x = "% cover", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot4 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = fittedVals)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)", y = "mean sev.",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot5 <- ggplot(plotData, 
                       aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## these affect via nu/tau
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot6 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "Severity ~ FlamConif",
       x = "% cover", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot7 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ roughness (SRR)",
       x = "% roughness", y = "mean sev.", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot8 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanPEFplot9 <- ggplot(plotData, aes(x = as.factor(isForest), y = fittedVals)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ isForest",
       x = "", y = "mean sev.", linetype = "understorey") +
  guides(shape = guide_legend(override.aes = list(colour = "black")))


legPlot <- get_legend(meanPEFplot1 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "meanSevPEFPlots_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
ggarrange(meanPEFplot1 + scale_y_continuous(limits = c(0.6, 1)), #decid
          meanPEFplot6 + scale_y_continuous(limits = c(0.6, 1)), #flamconif
          meanPEFplot3 + scale_y_continuous(limits = c(0.6, 1)), #lari
          meanPEFplot4 + scale_y_continuous(limits = c(0.6, 1)), #SMR 
          meanPEFplot9 + scale_y_continuous(limits = c(0.6, 1)), #isForest
          meanPEFplot2 + scale_y_continuous(limits = c(0.6, 1)), #CTI
          meanPEFplot7 + scale_y_continuous(limits = c(0.6, 1)), #SRR
          meanPEFplot8,  #buffer_30m
          meanPEFplot5 + scale_y_continuous(limits = c(0.6, 1)), #buffer_960m
          nrow = 3, ncol = 3, 
          legend.grob = legPlot,
          font.label = list(size = 16), 
          legend = "bottom")
dev.off()
```

#### Partial effects on mu - forest pixels
Note that "mu" is the fitted mean of values in (0,1)
Showing partial effects in forested pixels only, except when isForest was significant
```{r fittedMuPEFPPlots}
## mu
vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

muPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height",
       x = "% cover", y = expression(mu), 
       linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muPEFplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)",
       x = "CTI", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muPEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari",
       x = "% cover", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muPEFplot4 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = muPred)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)", y = expression(mu), 
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muPEFplot5 <- ggplot(plotData, 
                     aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

legPlot <- get_legend(muPEFplot1 + 
                        theme(legend.box = "vertical"))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "muSevPEFPlots_DAfires_sevModels.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(muPEFplot1 + scale_y_continuous(limits = c(0, 0.6)), #decid
          muPEFplot3 + scale_y_continuous(limits = c(0, 0.6)), #lari
          muPEFplot4 + scale_y_continuous(limits = c(0, 0.6)), #SMR
          muPEFplot2 + scale_y_continuous(limits = c(0, 0.6)), #CTI
          muPEFplot5 + scale_y_continuous(limits = c(0, 0.6)), #buffer_960m
          as_ggplot(legPlot),
          nrow = 3, ncol = 2, 
          font.label = list(size = 16), 
          legend = "none")
dev.off()
```

#### Partial effects on variance - forest pixels
Showing partial effects in forested pixels only, except when isForest was significant
Note that because $\mu$ and $\sigma$ both influence the variance, we need to plot it against all $\mu$ and $\sigma$ covariates.
```{r VarYPEFPlots}
## variance - sigma covariates
## this plot includes the effect of height from mu and tau
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

varPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Broadleaf * Height",
       y = "Expected variance",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## the following also includes effect via nu
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot2 <-  ggplot(plotData, aes(x = as.factor(SMR_wet), y = varY)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (SMR)",
       y = "Expected variance",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_1920m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_1920m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot3 <- ggplot(plotData, 
                      aes(x = buffer_1920m * attr(modelDataFULLScaled$buffer_1920m, "scaled:scale") + attr(modelDataFULLScaled$buffer_1920m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), size = 1) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (1920m)",
       y = "Expected variance", 
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - mu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot4 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (CTI)",
       x = "CTI", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## includes effect of height from tau
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = TRUE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

varPEFplot5 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Lari * Height", y = "Expected variance",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot6 <- ggplot(plotData, 
                      aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot7 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "Var(Severity) ~ FlamConif",
       x = "% cover", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot8 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ roughness (SRR)",
       x = "% roughness", y = "Expected variance", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = TRUE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot9 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", y = "Expected variance",
       linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varPEFplot10 <- ggplot(plotData, aes(x = as.factor(isForest), y = varY)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ isForest",
       x = "", y = "Expected variance", shape = "understorey")

legPlot <- get_legend(varPEFplot1 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "varSevPEFPlots_DAfires_sevModels1.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(
  varPEFplot1 + scale_y_continuous(limits = c(0, 0.25)), #Decid
  varPEFplot7 + scale_y_continuous(limits = c(0, 0.25)), #FlamConif
  varPEFplot5 + scale_y_continuous(limits = c(0, 0.25)), #Lari
  varPEFplot2 + scale_y_continuous(limits = c(0, 0.25)), #smr
  varPEFplot10 + scale_y_continuous(limits = c(0, 0.25)), #isForest
  as_ggplot(legPlot),
  nrow = 3, ncol = 2,  
  common.legend = TRUE, legend = "none",
  font.label = list(size = 16)
)
dev.off()

legPlot <- get_legend(varPEFplot4 + 
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))
png(file.path(figDir, "varSevPEFPlots_DAfires_sevModels2.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(
  varPEFplot4 + scale_y_continuous(limits = c(0, 0.25)), #cti
  varPEFplot8 + scale_y_continuous(limits = c(0, 0.25)), #SurfReliefRatio
  varPEFplot9 + scale_y_continuous(limits = c(0, 0.25)), #buffer_30m
  varPEFplot6 + scale_y_continuous(limits = c(0, 0.25)), #buffer_960m
  varPEFplot3 + scale_y_continuous(limits = c(0, 0.25)), #buffer_1920m
  as_ggplot(legPlot),
  nrow = 3, ncol = 2,  
  common.legend = TRUE, legend = "none",
  font.label = list(size = 16)
)
dev.off()
```

#### Partial effects on probabilities at 0 and 1 - non-forest pixels
Showing partial effects in non-forested pixels only, except when isForest was significant
Note that because $\nu$ and $\tau$ influence both *p0* and *p1* we need to plot them against all $\nu$ and $\tau$ covariates.
```{r p0p1PEFPlotsNonForest}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest")
## 0 probs - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot1 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p0 ~ FlamConif",
       x = "% cover", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot2 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ roughness (SRR)",
       x = "% roughness", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot3 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = p0)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (SMR)",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot4 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## 0 probs - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot5 <- ggplot(plotData, aes(x = as.factor(isForest), y = p0)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ isForest",
       x = "", linetype = "understorey") +
  guides(shape = guide_legend(override.aes = list(colour = "black")))

vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p0NFPEFplot6 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Broadleaf * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p0NFPEFplot7 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ Lari * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p0NFPEFplot8 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p0)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p0 ~ moisture (CTI)",
       x = "CTI", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

legPlot <- get_legend(p0NFPEFplot6 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## override scale for clarity
## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "p0PEFPlotsNotForest_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
ggarrange(
  p0NFPEFplot6 + scale_y_continuous(limits = c(0, 0.10)), #Decid
  p0NFPEFplot1 + scale_y_continuous(limits = c(0, 0.10)), #Flamconif
  p0NFPEFplot7 + scale_y_continuous(limits = c(0, 0.10)), #Lari
  p0NFPEFplot3 + scale_y_continuous(limits = c(0, 0.10)), #SMR
  p0NFPEFplot5 + scale_y_continuous(limits = c(0, 0.10)), #isForest
  p0NFPEFplot8 + scale_y_continuous(limits = c(0, 0.10)), #CTI
  p0NFPEFplot2 + scale_y_continuous(limits = c(0, 0.10)), #SRR
  p0NFPEFplot4, #buffer_30m
  as_ggplot(legPlot),
  nrow = 3, ncol = 3,
  font.label = list(size = 16),
  legend = "none")
dev.off()

## 1 probs - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot1 <- ggplot(plotData, aes(x = as.factor(isForest), y = p1)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ isForest",
       x = "", linetype = "understorey") +
  guides(shape = guide_legend(override.aes = list(colour = "black")))

vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p1NFPEFplot2 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Broadleaf * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

p1NFPEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ Lari * Height",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot4 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (CTI)",
       x = "CTI", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot5 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ roughness (SRR)",
       x = "% roughness", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot6 <- ggplot(plotData, 
                       aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot7 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = p1)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "p1 ~ moisture (SMR)",
       x = "", shape = "understorey")


## 1 probs - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
p1NFPEFplot8 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = p1)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "p1 ~ FlamConif",
       x = "% cover", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

legPlot <- get_legend(p1NFPEFplot2 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "p1PEFPlotsNotForest_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
ggarrange(
  p1NFPEFplot2, #Decid
  p1NFPEFplot8, #Flamconif 
  p1NFPEFplot3, #Lari
  p1NFPEFplot7, #SMR
  p1NFPEFplot1, #isForest
  p1NFPEFplot4, #CTI
  p1NFPEFplot5, #SRR
  p1NFPEFplot6, #buffer_30m
  as_ggplot(legPlot),
  nrow = 3, ncol = 3,
  font.label = list(size = 16),
  legend = "none")
dev.off()
```

#### Partial effects on mean - non-forest pixels
Showing partial effects in non-forested pixels only, except when isForest was significant
Note that because $\mu$, $\nu$ and $\tau$ influence the expected mean, it has to be plot against all $\mu$, $\nu$ and $\tau$ covariates.
```{r fittedMeanPEFPPlotsNonForest}
## mean
vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

meanNFPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height",
       x = "% cover", y = "mean sev.", 
       linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)",
       x = "CTI", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

meanNFPEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari",
       x = "% cover", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot4 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = fittedVals)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)", y = "mean sev.",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot5 <- ggplot(plotData, 
                         aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$fittedVals), max(plotData$fittedVals))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## these affect via nu/tau
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot6 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "Severity ~ FlamConif",
       x = "% cover", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot7 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ roughness (SRR)",
       x = "% roughness", y = "mean sev.", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot8 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = fittedVals)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", y = "mean sev.", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
meanNFPEFplot9 <- ggplot(plotData, aes(x = as.factor(isForest), y = fittedVals)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ isForest",
       x = "", y = "mean sev.", linetype = "understorey") +
  guides(shape = guide_legend(override.aes = list(colour = "black")))


legPlot <- get_legend(meanNFPEFplot1 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "meanSevPEFPlotsNotForest_DAfires_sevModels.png"), height = 13, width = 13,
    units = "in", res = 300)
ggarrange(meanNFPEFplot1 + scale_y_continuous(limits = c(0.5, 1)), #Decid
          meanNFPEFplot6 + scale_y_continuous(limits = c(0.5, 1)), #Flamconif
          meanNFPEFplot3 + scale_y_continuous(limits = c(0.5, 1)), #Lari
          meanNFPEFplot4 + scale_y_continuous(limits = c(0.5, 1)), #SMR  
          meanNFPEFplot9 + scale_y_continuous(limits = c(0.5, 1)), #isForest
          meanNFPEFplot2 + scale_y_continuous(limits = c(0.5, 1)), #CTI
          meanNFPEFplot7 + scale_y_continuous(limits = c(0.5, 1)), #SRR
          meanNFPEFplot8, #buffer_30m
          meanNFPEFplot5 + scale_y_continuous(limits = c(0.6, 1)), #buffer_960m
          nrow = 3, ncol = 3, 
          legend.grob = legPlot,
          font.label = list(size = 16), 
          legend = "bottom")
dev.off()
```

#### Partial effects on mu - non-forest pixels
Note that "mu" is the fitted mean of values in (0,1)
Showing partial effects in non-forested pixels only, except when isForest was significant
```{r fittedMuPEFPPlotsNonForest}
## mean
vars <- expand.grid(UNDERSTOREY = c(FALSE, TRUE), HEIGH_LOWER = quantile(na.omit(modelDataFULLScaled$HEIGHT_LOWER), probs = c(0.05, 0.50, 0.95)))
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

muNFPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Broadleaf * Height",
       x = "% cover", y = expression(mu), 
       linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muNFPEFplot2 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (CTI)",
       x = "CTI", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muNFPEFplot3 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ Lari",
       x = "% cover", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muNFPEFplot4 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = muPred)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Severity ~ moisture (SMR)", y = expression(mu),
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
muNFPEFplot5 <- ggplot(plotData, 
                       aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = muPred)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$muPred), max(plotData$muPred))) +
  theme_pubr() + 
  labs(title = "Severity ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = expression(mu), linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

legPlot <- get_legend(muNFPEFplot1 + 
                        theme(legend.box = "vertical"))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "muSevPEFPlotsNotForest_DAfires_sevModels.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(muNFPEFplot1 + scale_y_continuous(limits = c(0, 0.6)), #Decid
          muNFPEFplot3 + scale_y_continuous(limits = c(0, 0.6)), #Lari
          muNFPEFplot4 + scale_y_continuous(limits = c(0, 0.6)), #SMR
          muNFPEFplot2 + scale_y_continuous(limits = c(0, 0.6)), #CTI
          muNFPEFplot5 + scale_y_continuous(limits = c(0, 0.6)), #buffer_960m
          as_ggplot(legPlot),
          nrow = 3, ncol = 2, 
          font.label = list(size = 16), 
          legend = "none")
dev.off()
```

#### Partial effects on variance - non-forest pixels
Showing partial effects in non-forested pixels only, except when isForest was significant
Note that because $\mu$ and $\sigma$ both influence the variance, we need to plot it against all $\mu$ and $\sigma$ covariates.
```{r VarYPEFPlotsNonForest}
## variance - sigma covariates
## this plot includes the effect of height from mu
plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Decid", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y,  "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

varNFPEFplot1 <- ggplot(plotData, aes(x = Decid * attr(modelDataFULLScaled$Decid, "scaled:scale") + attr(modelDataFULLScaled$Decid, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Broadleaf * Height",
       y = "Expected variance",
       x = "% cover", linetype = "understorey", colour = "height") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))


plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SMR_wet",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot2 <- ggplot(plotData, aes(x = as.factor(SMR_wet), y = varY)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not wet/aquatic", "TRUE" = "wet/aquatic")) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (SMR)", y = "Expected variance",
       x = "", shape = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_1920m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_1920m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot3 <- ggplot(plotData, 
                        aes(x = buffer_1920m * attr(modelDataFULLScaled$buffer_1920m, "scaled:scale") + attr(modelDataFULLScaled$buffer_1920m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (1920m)",
       y = "Expected variance", 
       x = "prop. neigh. burnt", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - mu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "CTI",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot4 <- ggplot(plotData, aes(x = CTI * attr(modelDataFULLScaled$CTI, "scaled:scale") + attr(modelDataFULLScaled$CTI, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ moisture (CTI)",
       x = "CTI", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- mapply(FUN = function(x, y) {
  DT <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "Lari", 
                   fixed.at = list("UNDERSTOREY" = x, "HEIGHT_LOWER" = y, "isForest" = FALSE))
  DT$UNDERSTOREY <- x
  DT$HEIGHT_LOWER <- y
  DT
},
x = vars$UNDERSTOREY, y = vars$HEIGH_LOWER, SIMPLIFY = FALSE)
plotData <- rbindlist(plotData)

varNFPEFplot5 <- ggplot(plotData, aes(x = Lari * attr(modelDataFULLScaled$Lari, "scaled:scale") + attr(modelDataFULLScaled$Lari, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY), 
                colour = as.factor(HEIGHT_LOWER * attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:scale") + attr(modelDataFULLScaled$HEIGHT_LOWER, "scaled:center"))), lineend = "round", size = 1.5) +  
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ Lari",
       x = "% cover", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m", 
                       fixed.at = list("UNDERSTOREY" = FALSE,  "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_960m",
                        fixed.at = list("UNDERSTOREY" = TRUE,  "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot6 <- ggplot(plotData, 
                        aes(x = buffer_960m * attr(modelDataFULLScaled$buffer_960m, "scaled:scale") + attr(modelDataFULLScaled$buffer_960m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(min(plotData$varY), max(plotData$varY))) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (960m)",
       x = "prop. neigh. burnt", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - nu covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "FlamConif",
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot7 <- ggplot(plotData, aes(x = FlamConif * attr(modelDataFULLScaled$FlamConif, "scaled:scale") + attr(modelDataFULLScaled$FlamConif, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() +
  labs(title = "Var(Severity) ~ FlamConif",
       x = "% cover", y = "Expected variance", linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "SurfReliefRatio", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot8 <- ggplot(plotData, aes(x = SurfReliefRatio * attr(modelDataFULLScaled$SurfReliefRatio, "scaled:scale") + attr(modelDataFULLScaled$SurfReliefRatio, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ roughness (SRR)",
       x = "% roughness", y = "Expected variance", linetype = "understorey")

plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                       fixed.at = list("UNDERSTOREY" = FALSE, "isForest" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "buffer_30m", 
                        fixed.at = list("UNDERSTOREY" = TRUE, "isForest" = FALSE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot9 <- ggplot(plotData, aes(x = buffer_30m * attr(modelDataFULLScaled$buffer_30m, "scaled:scale") + attr(modelDataFULLScaled$buffer_30m, "scaled:center"), y = varY)) +
  geom_line(aes(linetype = as.factor(UNDERSTOREY)), lineend = "round", size = 1.5) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ neigh. burns (30m)",
       x = "prop. neigh. burnt", y = "Expected variance",
       linetype = "understorey") +
  guides(linetype = guide_legend(override.aes = list(colour = "black")))

## variance - tau covariates
plotData <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest", 
                       fixed.at = list("UNDERSTOREY" = FALSE))
plotData2 <- PEFwrapper(modelNgbSev4FULL4_Scaled, modelDataFULLScaled, term = "isForest",
                        fixed.at = list("UNDERSTOREY" = TRUE))
plotData$UNDERSTOREY <- FALSE
plotData2$UNDERSTOREY <- TRUE
plotData <- rbind(plotData, plotData2)
varNFPEFplot10 <- ggplot(plotData, aes(x = as.factor(isForest), y = varY)) +
  geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
  scale_x_discrete(labels = c("FALSE" = "not forest", "TRUE" = "forest")) +
  scale_y_continuous(limits = c(0,1)) +
  theme_pubr() + 
  labs(title = "Var(Severity) ~ isForest",
       x = "", y = "Expected variance", shape = "understorey")

legPlot <- get_legend(varNFPEFplot1 + 
                        geom_point(aes(shape = as.factor(UNDERSTOREY)), size = 2) +
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))

## order follows veg, moisture, forest, topo and neighbourhood
png(file.path(figDir, "varSevPEFPlotsNotForest_DAfires_sevModels1.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(
  varNFPEFplot1 + scale_y_continuous(limits = c(0, 0.25)), #Decid
  varNFPEFplot7 + scale_y_continuous(limits = c(0, 0.25)), #FlamConif
  varNFPEFplot5 + scale_y_continuous(limits = c(0, 0.25)), #Lari
  varNFPEFplot2 + scale_y_continuous(limits = c(0, 0.25)), #smr
  varNFPEFplot10 + scale_y_continuous(limits = c(0, 0.25)), #isForest
  as_ggplot(legPlot),
  nrow = 3, ncol = 2,  
  common.legend = TRUE, legend = "none",
  font.label = list(size = 16)
)
dev.off()

legPlot <- get_legend(varNFPEFplot4 + 
                        theme(legend.box = "vertical") +
                        labs(shape = "understorey") +
                        guides(linetype = guide_legend(order = 1),
                               shape = guide_legend(order = 2)))
png(file.path(figDir, "varSevPEFPlotsNotForest_DAfires_sevModels2.png"), height = 10, width = 10,
    units = "in", res = 300)
ggarrange(
  varNFPEFplot4 + scale_y_continuous(limits = c(0, 0.25)), #cti
  varNFPEFplot8 + scale_y_continuous(limits = c(0, 0.25)), #SurfReliefRatio
  varNFPEFplot9 + scale_y_continuous(limits = c(0, 0.25)), #buffer_30m
  varNFPEFplot6 + scale_y_continuous(limits = c(0, 0.25)), #buffer_960m
  varNFPEFplot3 + scale_y_continuous(limits = c(0, 0.25)), #buffer_1920m
  as_ggplot(legPlot),
  nrow = 3, ncol = 2,  
  common.legend = TRUE, legend = "none",
  font.label = list(size = 16)
)
dev.off()
```


#### Maps of observed and predicted severity
```{r sevMaps}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "SurfReliefRatio", "isForest", "pixID",
          "SEV_CLASS", "Lat", "Long")

modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")

predictionsDT <- na.omit(modelDataFULLScaled)[, .(SEV_PROP, SEV_CLASS, pixID, FIRE_NAME, Lat, Long)]
predData <- predictAll(modelNgbSev4FULL4_Scaled, type = "response")
predictionsDT[, `:=`(predSEV_PROP = calcMeanBEINF(predData$mu, predData$nu, predData$tau),
                     p1 = calcP1BEINF(predData$nu, predData$tau),
                     varY = calcVarBEINF(predData$mu, predData$sigma, predData$nu, predData$tau))]
rm(predData, modelDataFULLScaled)

## make severity rasters
firePoints <- predictionsDT[, .(pixID, SEV_PROP, predSEV_PROP, p1, varY, FIRE_NAME, Lat, Long)]
firePoints <- st_as_sf(firePoints, coords = c("Long", "Lat"),
                       agr = "constant", crs = crsProj) |> 
  vect()

## select 2 fires with high agreement between predictions and observations
## correlation between pred and obs severity by fire.
fires2mapHighAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
                                         by = FIRE_NAME][order(V1)] |>
                             tail(2))$FIRE_NAME |>
  as.character()

fires2mapLowAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
                                        by = FIRE_NAME][order(V1)] |>
                            head(2))$FIRE_NAME |>
  as.character()


plotFun <- function(firePoints, var, fireID, varTitle) {
  fireID <- sub("_(obs|pred|p1|varY)", "", fireID)
  ggplot(firePoints[firePoints$FIRE_NAME == fireID]) +
    geom_spatvector(aes_string(colour = var)) +
    scale_color_distiller(palette = "RdYlBu", direction = -1,
                          limits = c(0,1)) +
    theme(axis.text.x = element_text(hjust = 1, angle = 20)) +
    labs(colour = "", x = "Longitude", y = "Latitude", 
         title = varTitle)
}

sevPlotsHighAgrement <- Map(f = plotFun,
                            fireID = as.vector(outer(fires2mapHighAgreement, 
                                                     c("_obs", "_pred", "_p1", "_varY"), paste0)),
                            var = rep(c("SEV_PROP", "predSEV_PROP", "p1", "varY"), each = 2),
                            varTitle = rep(c("Observed", "Predicted mean sev.", 
                                             "Predicted prob. of max. sev.", "Variance"),
                                           each = 2),
                            MoreArgs = list(firePoints = firePoints))

sevPlotsLowAgrement <- Map(f = plotFun,
                           fireID = as.vector(outer(fires2mapLowAgreement, 
                                                    c("_obs", "_pred", "_p1", "_varY"), paste0)),
                           var = rep(c("SEV_PROP", "predSEV_PROP", "p1", "varY"), each = 2),
                           varTitle = rep(c("Observed", "Predicted mean sev.", 
                                            "Predicted prob. of max. sev.", "Variance"),
                                          each = 2),
                           MoreArgs = list(firePoints = firePoints))

png(file.path(figDir, "severityMaps_highAgreement.png"), height = 8, width = 16,
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsHighAgrement$BUFFALO_HEAD_obs + theme(axis.title.x = element_blank()), 
  sevPlotsHighAgrement$BUFFALO_HEAD_pred + theme(axis.title = element_blank()), 
  sevPlotsHighAgrement$BUFFALO_HEAD_p1 + theme(axis.title = element_blank()),
  sevPlotsHighAgrement$BUFFALO_HEAD_varY + theme(axis.title = element_blank()) + scale_color_distiller(palette = "RdYlBu", direction = -1),
  ## line 2
  sevPlotsHighAgrement$LIVOCK_obs, 
  sevPlotsHighAgrement$LIVOCK_pred + theme(axis.title.y = element_blank()), 
  sevPlotsHighAgrement$LIVOCK_p1 + theme(axis.title.y = element_blank()), 
  sevPlotsHighAgrement$LIVOCK_varY + theme(axis.title.y = element_blank()) + scale_color_distiller(palette = "RdYlBu", direction = -1), 
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "", "", "b)", "", "", ""), 
  ncol = 4, nrow = 2
)
dev.off()

png(file.path(figDir, "severityMaps_lowAgreement.png"), height = 8, width = 12,   
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsLowAgrement$`CONTEST #2_obs` + theme(axis.title.x = element_blank()),
  sevPlotsLowAgrement$`CONTEST #2_pred` + theme(axis.title = element_blank()), 
  sevPlotsLowAgrement$`CONTEST #2_p1` + theme(axis.title = element_blank()),
  sevPlotsLowAgrement$`CONTEST #2_varY` + theme(axis.title = element_blank()) + scale_color_distiller(palette = "RdYlBu", direction = -1),
  ## line 2
  sevPlotsLowAgrement$WOODMAN_obs, 
  sevPlotsLowAgrement$WOODMAN_pred + theme(axis.title.y = element_blank()), 
  sevPlotsLowAgrement$WOODMAN_p1 + theme(axis.title.y = element_blank()),
  sevPlotsLowAgrement$WOODMAN_varY + theme(axis.title = element_blank()) + scale_color_distiller(palette = "RdYlBu", direction = -1),
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "", "", "b)", "", "", ""),
  ncol = 4, nrow = 2
)
dev.off()
```


```{r saveLoadRdata, include = FALSE, eval = FALSE}
# save(list = c("plotData", grep("model", ls(), value = TRUE)),
#      # file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_20m_goodSample_Nov1.RData")
#      # file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_30m_goodSample_Nov1.RData")
#      file = "/media/data/cbarros/LandscapesInMotion/analyses/modelsGAMLSS_0-3Days_30m_goodSample_NgbSevApr21.RData")

## save to GDrive
# library(googledrive)
# save.image(file = "/mnt/storage/cbarros/LandscapesInMotion/analyses/LIM.RData")
# drive_auth(use_oob = TRUE)
# drive_upload("~/Downloads/LIM.RData", path = as_id("1jSrwbU83NVBosWFTxnv2JB2FVkiR5OLD"))

# load("analyses/modelsGAMLSS_0-3Days_goodSample_Nov1.RData")
```


## Accuracy values of similar studies

Fang et al. 2018 and Ferster et al. 2016 model severity using BRTs and multinomial logistic models, respectively. In Fang et al. 2018 only 3 severity classes were modelled. Ferster et al. 2016 used the same data as here.

Note that the contingency matrix in Fang et al. 2018 was transposed to match the format expected by `caret::confusionMatrix`.

```{r accuracyFangAndFerster}
FangEtAl2018_yr2000 <- data.frame(low = c(50, 26, 7),
                                  moderate = c(261, 496, 142),
                                  high = c(42, 369, 1184), 
                                  row.names = c("low", "moderate", "high")) |>
  as.matrix() |>
  as.table()


FangEtAl2018_yr2010 <- data.frame(low = c(196, 263, 94),
                                  moderate = c(95, 276, 115),
                                  high = c(33, 431, 662), 
                                  row.names = c("low", "moderate", "high")) |>
  as.matrix() |>
  as.table()


confusionMatrix(FangEtAl2018_yr2000)
confusionMatrix(FangEtAl2018_yr2010)
```

