---
title: "Dave Andison's fire data Severity models - using Gradient-Boosted Trees"
author: "Ceres Barros"
date: "09/09/2025"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(getwd()))
knitr::opts_chunk$set(echo = FALSE, cache = 2, warning = FALSE)
```

```{r setup2, include=FALSE}
if (!exists("pkgDir")) {
  pkgDir <- file.path(
    if (Sys.info()[["user"]] == "rstudio") "packages_docker" else "packages",
    version$platform,
    paste0(version$major, ".", strsplit(version$minor, "[.]")[[1]][1])
  )
  
  if (!dir.exists(pkgDir)) {
    dir.create(pkgDir, recursive = TRUE)
  }
  .libPaths(pkgDir)
}

install.packages("Require")

Require::Require("xgboost (==3.0.5.1)", repos = "https://dmlc.r-universe.dev") ## not working "SSL connect error", same with install.packages('xgboost', repos = c('https://dmlc.r-universe.dev', 'https://cloud.r-project.org'))
# if (!"xgboost" %in% rownames(installed.packages()) && packageVersion("xgboost") != "3.0.5.1") {
#   install.packages("C:/Users/cbarros/Downloads/xgboost_3.0.5.1.tar.gz")  ## needed to manually download
# }
Require::Install("party")  ## for gpboost

Require::Require(
  c(
    # "caret",
    "crayon",
    "data.table",
    "dplyr",
    "foreign",
    "fasterize",
    "future",
    "future.apply",
    "gamlss",
    "gamlss.inf", 
    "gpboost",
    "CeresBarros/gpboost.utils (>=0.1.0.9001)",
    "ggplot2",
    "ggpubr",
    "ggrepel",
    "Hmisc",
    "LandR.CS",
    "lightgbm",
    "prioritizr",
    "purrr",
    "pROC",
    "raster",
    "PredictiveEcology/reproducible@AI",
    "reshape2",
    "sf",
    "shapr",  
    "PredictiveEcology/SHAPforxgboost (>=0.1.3.9001)",
    "terra",
    "tidyterra",
    "CeresBarros/ToolsCB (>=0.0.9001)",
    # "xgboost (>=3.0.5.1)",  ## above
    "xlsx"
  )
)

devtools::load_all("C:/Users/cbarros/GitHub/caret/pkg/caret/")

options("reproducible.cachePath" = "analyses/cache",
        "reproducible.useDBI" = TRUE,
        "future.globals.maxSize" = +Inf)

source("R/R_tools/Useful_functions.R") ## may replace some functions in ToolsCB, this is ok
source("R/R_tools/CASFRIrelated_functions.R")
source("R/R_tools/prepFireWeather.R")
source("R/R_tools/joinSevVegTopoWeatherData.R")
source("analyses/R_tools/summarizeABSK_AllData.R")
source("analyses/R_tools/DAfires_expAnalyses_dataPrep.R")

doCache = TRUE

figDir <- "analyses/Figs"
dir.create(figDir, showWarnings = FALSE)
```

```{r dataprep, include = FALSE, cache = FALSE}
resolution = 30
summaryABSK_AllData <- Cache(
  dataPrepWrapper,
  resolution = resolution, 
  fireDataPath = "data/fires_Dave/fireSev",
  vegDataPath = "data/fires_Dave/prefireVeg",
  topoDataPath = "data/fires_Dave/DEM/Grid30Intersect",
  weatherDataPath = "data/fires_Dave/fireWeather",
  doCache = doCache,
  useCache = doCache,
  userTags = "summaryABSK_AllData",
  cacheId = "0a70f8b25dc39895",   ## overwride. oct 22 2025, Cache is being silly about "new" userTags
  omitArgs = c("useCache", "doCache", "userTags"))
```

```{r addNgbNoBurns, cache = FALSE, include = FALSE}
## make different buffer sizes around points
crsProj <- crs(vect("data/fires_Dave/fireSev/albertafires1_postfire.shp"), proj = TRUE)

SEV_PROPpoints <- summaryABSK_AllData[, .(pixID, FIRE_NAME, SEV_PROP, Lat, Long)]
SEV_PROPpoints <- st_as_sf(SEV_PROPpoints, coords = c("Long", "Lat"),
                           agr = "constant", crs = st_crs(crsProj))

bufferDists <- 30
for (i in 2:7) {
  bufferDists[i] <- bufferDists[i - 1] * 2
}

ngbNoBurns <- reproducible::Cache(calculateNgbBurnsWrapper,
                                  dists = bufferDists,
                                  sevPoints = SEV_PROPpoints, 
                                  sevColID = "SEV_PROP",
                                  fireColID = "FIRE_NAME",
                                  resolution = resolution,
                                  parallel = TRUE,
                                  cores = 3,
                                  useCache = doCache,
                                  cacheId = "9a366f60217632b1",
                                  userTags = "ngbBurns",
                                  omitArgs = c("useCache", "userTags", "cores")) 
ngbNoBurns <- dcast.data.table(ngbNoBurns, ... ~ bufferSize, value.var = "ngbPropBurns")

cols <- grep("pixID|FIRE_NAME", names(ngbNoBurns), invert = TRUE, value = TRUE)
setnames(ngbNoBurns, old = cols, new = paste0("buffer_", cols, "m"))

summaryABSK_AllData <- ngbNoBurns[summaryABSK_AllData, on = .(pixID, FIRE_NAME)]

rm(SEV_PROPpoints); gc(reset = TRUE)
```

## Maps of fire perimeters, ecozones and nat. regions
```{r}
firesABSK <- Cache(cleanAndBindFireData,
                   files = c("albertafires1_postfire", "albertafires2_postfire", "saskatchewanfires_postfire"),
                   fireDataPath = "data/fires_Dave/fireSev",
                   cachePath = "analyses/cache",
                   userTags = "allFireData",
                   cacheId = "allFireData",
                   omitArgs = c("cachePath", "userTags"))
firesABSK <- vect(firesABSK)

NRAlberta <- prepInputs(targetFile = "AB_Natural_Sub_Regions.shp",
                        archive = "AB_Natural_Sub_Regions.zip",
                        alsoExtract = "similar",
                        url = "https://drive.google.com/file/d/1hW6zy0CpUBdk-K2IAjzW4INjVl1J4aLJ",
                        fun = "sf::st_read",
                        destinationPath = "data/maps",
                        cachePath = "analyses/cache",
                        userTags = "NRAlberta")

NRAlberta <- Cache(prepInputs,
                   targetFile = "Natural_Regions_Subregions_of_Alberta.shp",
                   archive = asPath("natural_regions_subregions_of_alberta.zip"),
                   url = "https://www.albertaparks.ca/media/429607/natural_regions_subregions_of_alberta.zip",
                   alsoExtract = "similar",
                   fun = "sf::st_read",
                   destinationPath = "R/SpaDES/inputs/",
                   cachePath = "R/SpaDES/data/cache",
                   userTags = c("prepInputsNatSubRegionsAB_SA"),
                   cacheId = "652465e92ab1f7e2",
                   omitArgs = c("cachePath", "userTags", "destinationPath", "alsoExtract"))

NRAlberta <- vect(NRAlberta)

## dissolve by region
NRAlberta <- terra::aggregate(NRAlberta[, "NRNAME"], by = "NRNAME", dissolve = TRUE)
NRAlberta <- project(NRAlberta, crs(firesABSK, proj = TRUE))

extentFires <- ext(rbind(firesABSK, NRAlberta))
extentFires <- vect(extentFires)
crs(extentFires) <- crs(firesABSK, proj = TRUE)

EcozonesCan <- prepInputs(targetFile = "ecozones.shp",
                          url = "https://sis.agr.gc.ca/cansis/nsdb/ecostrat/zone/ecozone_shp.zip",
                          archive = "ecozone_shp.zip",
                          fun = "sf::st_read",
                          destinationPath = "data/maps",
                          cachePath = "analyses/cache",
                          to = extentFires, 
                          userTags = "EcozonesCan")
EcozonesCan[EcozonesCan$ZONE_NAME == "Boreal PLain", "ZONE_NAME"] <- "Boreal Plain"
EcozonesCan <- vect(EcozonesCan["ZONE_NAME"])

## crop layers to a slightly larger area
CA_admin <- vect("data/CA_admin/gpr_000a11a_e.shp")
CA_admin <- project(CA_admin, crs(firesABSK, proj = TRUE))
CA_admin <- crop(CA_admin, extentFires)
NRAlberta <- crop(NRAlberta, extentFires)
EcozonesCan <- crop(EcozonesCan, extentFires)
rm(extentFires); gc(reset = TRUE)
```

```{r}
## Maps
if (FALSE) {
  opts <- par(bg = "white")
  plot(EcozonesCan, "ZONE_NAME", col = RColorBrewer::brewer.pal(name = "Pastel2", n = 7),
       border = NA, mar = c(3.1, 3.1, 2.1, 7.5), sort = TRUE, main = "")
  plot(CA_admin, legend = FALSE, add = TRUE)
  plot(NRAlberta, "NRNAME", 
       col = alpha("white", 0.1), sort = TRUE, legend = FALSE,
       border = RColorBrewer::brewer.pal(name = "Set1", n = 6), lwd = 2,
       alpha = 0.3, mar = c(3.1, 3.1, 2.1, 7.5), add = TRUE)
  legend("bottomright", bty = "n",
         cex = 0.85,
         inset = c(-0.2, 0),     # Negative inset moves legend outside
         legend = sort(unique(NRAlberta$NRNAME)),
         fill = alpha("white", 0),
         border = RColorBrewer::brewer.pal(name = "Set1", n = 6),
         xpd = TRUE,                 # Allow drawing outside plot region
         title = "")
  text(CA_admin[CA_admin$PRENAME %in% c("Alberta", "Saskatchewan")],
       labels = "PRENAME", adj = c(0.5, 6))
  plot(firesABSK[firesABSK$FIRE_NAME %in% summaryABSK_AllData$FIRE_NAME], legend = FALSE, add = TRUE)
  dev.print(png, filename = file.path(figDir, "firesEcozonesNatRAB.png"), height = 7, width = 8,
            units = "in", res = 300, bg = "white")
  
  
  plot(EcozonesCan, "ZONE_NAME", col = RColorBrewer::brewer.pal(name = "Pastel2", n = 7),
       border = NA, mar = c(3.1, 3.1, 2.1, 7.5), main = "")
  plot(CA_admin, legend = FALSE, add = TRUE)
  text(CA_admin[CA_admin$PRENAME %in% c("Alberta", "Saskatchewan")],
       labels = "PRENAME", adj = c(0.5, 6))
  plot(firesABSK[firesABSK$FIRE_NAME %in% summaryABSK_AllData$FIRE_NAME], legend = FALSE, add = TRUE)
  dev.print(png, filename = file.path(figDir, "firesEcozones.png"), height = 7, width = 8,
            units = "in", res = 300, bg = "white")
  par(opts)
}

## severity
if (FALSE) {
  ggplot(summaryABSK_AllData, aes(x = SEV_CONT)) +
    geom_histogram(fill = "cadetblue", alpha = 0.5) +
    theme_bw()
  
    ggplot(summaryABSK_AllData, aes(x = SEV_CLASS)) +
    geom_histogram(fill = "cadetblue", alpha = 0.5) +
    theme_bw()
  
plot(firesABSK[firesABSK$FIRE_NAME == unique(summaryABSK_AllData$FIRE_NAME)[5]], 
     "SEV_CONT", col = map.pal("inferno", 8)[c(2, 4:8)])


ggplot(summaryABSK_AllData, aes(x = meanFFMC, y = SEV_CONT)) +
  geom_point() +
  theme_bw()

}


```

## Summary of preliminary analyses

Variables selected for models (when using all fire weather days):
* Climate:
+ meanFFMC	-0.16***
+ cvDMC	0.16***
+ cvBUI	0.15***
+ minFFMC	-0.14***
+ rangeFFMC	0.14*** - highly correlated with minFFMC
+ cvFFMC	0.13*** - highly correlated with minFFMC
+ minDMC	-0.13*** - highly correlated with cvDMC
+ minBUI	-0.13*** - - highly correlated with min/cvDMC

* Topography
+ PctSlope	0.08***
+ SurfAspectRatio	0.08*** - **SHOULD BE CALLED SURFACE AREA RATIO** - 
+ Aspect	0.07***
+ CTI	-0.07*** ([Compound Topographic Index (CTI) (Moore et al. 1991) or "soil wetness" transformation can be used to models aspects of hydrologic systems. CTI is strongly correlated with soil moisture](https://wikispaces.psu.edu/display/AnthSpace/Compound+Topographic+Index))
+ SurfReliefRatio	0.04*** ([Describes rugosity. Ratio between relief and length in a particular window](https://en.wikipedia.org/wiki/Relief_ratio))
+ Elevation	-0.03***

* Vegetation - Spearmand correlations with severity
+ Group flammable conifers 0.10 ***
+ Pinu bank	0.15***
+ Pice mari	0.06***
+ Pice glau	-0.04***
+ Pinu cont	-0.04***
+ Pice enge	0.01***
+ Group Larix
+ Lari lari	-0.05***
+ Group Abies 0.00***
+ Abie lasi	0.00
+ Abie bals	0.01***
+ Group deciduous 0.03***
+ Betu papy	-0.03***
+ Popu trem	0.03***
+ Popu balb	0.02***

+ UNDERSTOREY
+ ORIGIN_UPPER 0.10***

* Landcover type
+ isForest

## Gradient boosted trees with XGBoost

### Fit and test model with k-fold cross-validation

*Note: we don't need to scale data for XGBoost*
See [here](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.1365-2656.2008.01390.x)

**Oct 30th 2025**: I tried using a Tweedie distribution. Residuals were just as bad 
and there were actually decreases in model fit.
```{r}
## variables for analysis when considering all weather (first 3 days) - dropping range
## due to lack of data in some fires, dropping:
## crown closure, SMR (which has a proxy- CTI), cvRain, ORIGIN_*
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "FIRE_NAME|pixID|SMR_ord|SMR_wet|buffer", sep = "|"), 
             names(summaryABSK_AllData),
             value = TRUE)
cols <- grep("^range", cols, invert = TRUE, value = TRUE)

modelData <- summaryABSK_AllData[, ..cols]
# cols2scale <- setdiff(cols, 
#                       c("pixID", "FIRE_NAME", "Czone1", "NR1", "SMR_ord", 
#                         "UNDERSTOREY", "isForest", "SEV_PROP", "SMR_wet"))
# modelData[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

## drop FIRENAME for now, as I'm not sure how to include it as a random effects
cols <- setdiff(cols, c("FIRE_NAME", "pixID"))
modelData <- modelData[, ..cols]
digestOfData <- .robustDigest(modelData)

modelGBT <- runXGBOOST(modelData, digestOfData, nFolds = 10
                       , SHAPthresh = 0.25  ## higher/lower threshold decreases AUC
                       , figDir = file.path(figDir, "XGBoost")
)
```

### RMSE and AUC

```{r}
RMSEs <- sapply(modelGBT, function(mm) {
  tail(attr(mm$mod,"evaluation_log"), 1)$eval_rmse
})

## AUC is inappropriate. This is not a classification problem
## yes, we could check accuracy of classifying 0s and 1s, but then we'd
## lose the middle values's accuracies.
AUCs <- sapply(modelGBT, function(mm) {
  tail(attr(mm$mod,"evaluation_log"), 1)$eval_auc
})

mean(AUCs)
```

### Confidence matrix
```{r}
## predict from each fold and average final confusion matrix
## runXGBOOST outputs predictions and observations and residuals in valData
outConfMats <- modelGBT |>
  purrr::map(\(x) xgboostConfMat(mod = x,
                                 classes = as.character(sort(unique(summaryABSK_AllData[["SEV_CLASS"]]))),
                                 classMap = unique(summaryABSK_AllData[, .SD, .SDcols = c("SEV_CLASS", "SEV_PROP")])
  ))

## make a list, then stack into 3-D array
confMats <- sapply(outConfMats, FUN = function(x) {
  DT <- as.data.table(x$table)
  DT <- dcast.data.table(DT, Prediction ~ Reference)
  cols <- names(DT)
  DT[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]
  as.matrix(DT, rownames = 1)
}, simplify = "array")

outConfMatsStats <- lapply(outConfMats, function(x) {
  DT <- as.data.table(x$byClass, keep.rownames = TRUE)
  setnames(DT, "rn", "Class")
  DT
}) |>
  rbindlist(idcol = "fold", use.names = TRUE)

```

### Feature importance

```{r}
figDir2 <- file.path(figDir, "XGBoost")
dir.create(figDir2, showWarnings = FALSE)
all_meanSHAP <- purrr::map(modelGBT,
                           function(modFold) {
                             as.data.table(as.list(modFold$shap_long))
                           }) |> 
  rbindlist(use.names = TRUE, idcol = "Fold")

meanSHAP <- all_meanSHAP[, .(Fold, variable, mean_value)]
meanSHAP <- meanSHAP[, list(mean_value = mean(mean_value)), by = variable]

varOrder <- meanSHAP[order(mean_value), variable]

ggplot(meanSHAP, aes(x = variable, y = mean_value)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = varOrder) +
  theme_pubr() +
  labs(x = "Predictor", y = "SHAP value") +
  coord_flip()
ggsave(file.path(figDir2, "Cross-fold_featImpSHAP.png"),
       width = 7, height = 8.5)


## all pairwise combos:
varcombos <- expand.grid(unique(shap_long$variable), unique(shap_long$variable))
varcombos <- apply(varcombos, 1, FUN = function(X) {
  X <- sort(X)
  names(X) <- NULL
  data.table(Var1 = X[1], Var2 = X[2])
}) |>
  rbindlist() |>
  unique()

varcombos <- varcombos[Var1 != Var2, ]
varcombos <- varcombos[!(grepl("buffer", Var1) & grepl("buffer", Var2))]  ## weird to plot buffer interactions

allPlots <- Map(f = function(x, y) {
  gtitle <- paste("SHAP for", x, "|", y)
  plot1 <- shap.plot.dependence(data_long = shap_long,
                                x = x,
                                y = y,
                                dilute = 20) +
    ggtitle(gtitle)
  plot1
}
, x = varcombos$Var1
, y = varcombos$Var2
)

DT <- data.table(plotIDs = 1:length(allPlots))
DT[, splits := cut(plotIDs, breaks = round(.N/6))]

figDir2 <- file.path(figDir,"XGBoost")
dir.create(figDir2, recursive = TRUE, showWarnings = FALSE)
Map(f = function(ids) {
  fileStr <- paste(head(ids$plotIDs, 1), tail(ids$plotIDs, 1), sep = "_")
  fileName <- paste0("SHAP_partialDep_", fileStr, ".png")
  plot1 <- ggarrange(plotlist = allPlots[ids$plotIDs])
  ggsave(file.path(figDir2, fileName), plot = plot1, 
         width = 12, height = 7.23,
         units = "in")
},
ids = split(DT, f = ~ splits))


shap_long <- modelGBT$Fold01$shap_long
shap.plot.dependence(data_long = shap_long,
                        x = 'HEIGHT_LOWER',
                        y = 'Decid',
                        dilute = 20) +
  ggtitle("SHAP for Decidous cover | Height")


## significant from GAMLSS
shap.plot.dependence(data_long = modelGBT$Fold01$shap_long,
                     x = 'UNDERSTOREYTRUE',
                     y = 'Decid',
                     dilute = 20) +
  ggtitle("SHAP for Decidous cover | Understorey presence")

## dropped:
# shap.plot.dependence(data_long = shap_long,
#                      x = 'UNDERSTOREYTRUE',
#                      y = 'isForestTRUE',
#                      dilute = 20) +
#   ggtitle("SHAP for Decidous cover | Height")

shap.plot.dependence(data_long = shap_long,
                     x = 'HEIGHT_LOWER',
                     y = 'Lari',
                     dilute = 20) +
  ggtitle("SHAP for Larch cover | Height")

shap.plot.dependence(data_long = shap_long,
                     x = 'UNDERSTOREYTRUE',
                     y = 'Lari',
                     dilute = 20) +
  ggtitle("SHAP for Larch cover | Understorey presence")

## other features
shap.plot.dependence(data_long = shap_long,
                     x = 'FlamConif',
                     dilute = 10) +
  ggtitle("SHAP for Flammable conifer cover")

shap.plot.dependence(data_long = shap_long,
                     x = 'HEIGHT_LOWER',
                     dilute = 10) +
  ggtitle("SHAP for Height")


shap.plot.dependence(data_long = shap_long,
                     x = 'minISI',
                     dilute = 20) +
  ggtitle("SHAP for min ISI over first 3 days of fire")
shap.plot.dependence(data_long = shap_long,
                     x = 'maxDMC',
                     dilute = 20) +
  ggtitle("SHAP for max DMC over first 3 days of fire")
shap.plot.dependence(data_long = shap_long,
                     x = 'cvBUI',
                     dilute = 20) +
  ggtitle("SHAP for cv BUI over first 3 days of fire")

shap.plot.dependence(data_long = shap_long,
                     x = 'buffer_30m',
                     dilute = 20) +
  ggtitle("SHAP for proportion of burnt neigh. - 30m radius")
shap.plot.dependence(data_long = shap_long,
                     x = 'buffer_60m',
                     dilute = 20) +
  ggtitle("SHAP for proportion of burnt neigh. - 60m radius")
shap.plot.dependence(data_long = shap_long,
                     x = 'buffer_1920m',
                     dilute = 20) +
  ggtitle("SHAP for proportion of burnt neigh. - 1920m radius")
```

### Feature selection

Constraining feature interactions to within "feature types" (i.e. climate, 
vegetation, topographic) yielded worse models.

```{r constrain_interactions}
interaction_constraints <- list(
  grep("^min|^max|^mean|^cv|^SMR|^buffer", colnamesPred, value = TRUE), ## interactions between climate vars, and climate and burnt neighbours
  grep("^Abie|^FlamConif|^Lari|^Decid|^UNDERST|^HEIGHT|^buffer", colnamesPred, value = TRUE), ## interactions between veg vars, and veg and burnt neighbours
  grep("^Elev|^Aspect|^CTI|^Surf|^PctSlope|^buffer", colnamesPred, value = TRUE) ## interactions between topo vars, and topo and burnt neighbours
)
modelGBT2 <- runXGBOOST(modelData, digestOfData, nFolds = 5
                        , SHAPthresh = 0.25  ## higher/lower threshold decreases AUC
                        , interaction_constraints = interaction_constraints   ## this decreased model performance
)
```

Constraining features and feature interactions to those in GAMLSS model also yielded
wore models (somewhere between the unconstrained XGBoost and GAMLSS)
```{r constrain_interactions}
## get all GAMLSS model covariates and interactions
modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")
vars <- sapply(coefAll(modelNgbSev4FULL4_Scaled), function(x) {
  names(x)[!grepl("Intercept|random", names(x))]
}) |>
  unlist()
vars <- unique(vars)
vars <- gsub("TRUE", "", vars)
ints <- vars[grepl(":", vars)] |>
  strsplit(split = ":") |>
  lapply(FUN = sort) |>
  unique()

vars <- unlist(strsplit(vars, split = ":")) |>
  unique()

modelData <- summaryABSK_AllData[, .SD, .SDcols = c("SEV_PROP", vars)]
digestOfData <- .robustDigest(modelData)
modelGBT3 <- runXGBOOST(modelData, digestOfData
                        , nFolds = 10
                        , SHAPthresh = 0 ## higher/lower threshold decreases AUC
                        , interaction_constraints = ints
)
```

### Maps of predicted severity
TODO: predict with full model, or combine CV preds

```{r}
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "FIRE_NAME|pixID|SMR_ord|SMR_wet|buffer", sep = "|"), 
             names(summaryABSK_AllData),
             value = TRUE)
cols <- grep("^range|FIRE_NAME", cols, invert = TRUE, value = TRUE)

modelData <- summaryABSK_AllData[, ..cols]

predictionsDT <- summaryABSK_AllData[, .(SEV_PROP, pixID, FIRE_NAME, Lat, Long)]
modelData2pred <- model.matrix(~ . + 0, data = modelData) |>
  as.data.table()

predData <- data.table(predSEV_PROP = predict(modelGBT$Fold01$mod, 
                                              modelData2pred, 
                                              type = "response"),
                       pixID = modelData2pred$pixID)
predictionsDT <- predData[predictionsDT, on = "pixID"]

rm(predData, modelData2pred); gc(reset = TRUE)

## make severity rasters
firePoints <- predictionsDT[, .(pixID, SEV_PROP, predSEV_PROP, FIRE_NAME, Lat, Long)]
firePoints <- st_as_sf(firePoints, coords = c("Long", "Lat"),
                       agr = "constant", crs = crsProj) |> 
  vect()

## select 2 fires with high agreement between predictions and observations
## correlation between pred and obs severity by fire.
# fires2mapHighAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
#                                          by = FIRE_NAME][order(V1)] |>
#                              tail(2))$FIRE_NAME |>
#   as.character()
# 
# fires2mapLowAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
#                                         by = FIRE_NAME][order(V1)] |>
#                             head(2))$FIRE_NAME |>
#   as.character()
## use same as GAMLSS for comparison
fires2mapHighAgreement <- c("BUFFALO_HEAD", "LIVOCK")
fires2mapLowAgreement <- c("CONTEST #2", "WOODMAN")

sevPlotsHighAgrement <- Map(f = plotFun,
                            fireID = as.vector(outer(fires2mapHighAgreement, 
                                                     c("_obs", "_pred"), paste0)),
                            var = rep(c("SEV_PROP", "predSEV_PROP"), each = 2),
                            varTitle = rep(c("Observed", "Predicted mean sev."),
                                           each = 2),
                            MoreArgs = list(firePoints = firePoints))

sevPlotsLowAgrement <- Map(f = plotFun,
                           fireID = as.vector(outer(fires2mapLowAgreement, 
                                                    c("_obs", "_pred"), paste0)),
                           var = rep(c("SEV_PROP", "predSEV_PROP"), each = 2),
                           varTitle = rep(c("Observed", "Predicted mean sev."),
                                          each = 2),
                           MoreArgs = list(firePoints = firePoints))

figDir2 <- file.path(figDir,"XGBoost")
png(file.path(figDir2, "severityMaps_highAgreement.png"), height = 8, width = 8,
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsHighAgrement$BUFFALO_HEAD_obs + theme(axis.title.x = element_blank()), 
  sevPlotsHighAgrement$BUFFALO_HEAD_pred + theme(axis.title = element_blank()),
  ## line 2
  sevPlotsHighAgrement$LIVOCK_obs, 
  sevPlotsHighAgrement$LIVOCK_pred + theme(axis.title.y = element_blank()), 
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "b)", ""), 
  ncol = 2, nrow = 2
)
dev.off()

png(file.path(figDir2, "severityMaps_lowAgreement.png"), height = 8, width = 6,   
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsLowAgrement$`CONTEST #2_obs` + theme(axis.title.x = element_blank()),
  sevPlotsLowAgrement$`CONTEST #2_pred` + theme(axis.title = element_blank()), 
  ## line 2
  sevPlotsLowAgrement$WOODMAN_obs, 
  sevPlotsLowAgrement$WOODMAN_pred + theme(axis.title.y = element_blank()), 
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "b)", ""),
  ncol = 2, nrow = 2
)
dev.off()
```

### Residuals
```{r}
valData <- modelGBT$Fold01$valData
plotResiduals(valData)
```



## Gradient boosted trees with GPBoost
Gaussian process and mixed effects models with gradient boosting

Allows modelling a zero-inflated gamma, which could be a good choice if we model
% survival, instead of mortality -- the severity data is mostly one-inflated and this
would flip it to mostly 0 inflated

### Fit and test model with k-fold cross-validation

```{r}
cols <- grep(paste("SEV_PROP|FFMC$|DC$|DMC$|ISI$|BUI$|FWI$",
                   "FlamConif|Abie$|Lari$|Decid",
                   "isForest|HEIGHT_LOWER|UNDERSTOREY",
                   "^CTI$|Aspect|Elevation|Surf.*Ratio|PctSlope",
                   "FIRE_NAME|pixID|SMR_ord|SMR_wet|buffer", sep = "|"), 
             names(summaryABSK_AllData),
             value = TRUE)
cols <- grep("^range", cols, invert = TRUE, value = TRUE)

modelData <- na.omit(summaryABSK_AllData[, ..cols])
modelData[, SURV_PROP := abs(SEV_PROP-1)]  ## create a survival variable

if (FALSE) {
  ggplot(modelData) +
    geom_density(aes(x = SEV_PROP), colour = "cadetblue", fill = "cadetblue", alpha = 0.5) +
    geom_density(aes(x = SURV_PROP), colour = "coral", fill = "coral", alpha = 0.5) +
    labs(x = "Severity (blue) and survival (coral)")
}

## select covariates, convert to numeric (these are specified as categorical later)
predvars <- setdiff(names(modelData), 
                    c("FIRE_NAME", "SEV_PROP", "SURV_PROP", "pixID"))
modelData[, (predvars) := lapply(.SD, as.numeric), .SDcols = predvars]

digestOfData <- .robustDigest(modelData)
## TODO: RE-RUN PARAMETER OPTIMIZATION AFTER PRESENTATION and export RMSE's

modelGPB <- runGPBOOST(dat = modelData
                       , digestOfData 
                       , nFolds = 10
                       , colnamesGrp = "FIRE_NAME"
                       , colnamesPred = predvars
                       , colnamesCat = c("SMR_wet", "isForest", "UNDERSTOREY", "SMR_ord")
                       , SHAPthresh = 0) #|>
## TODO: RE-RUN ALL INTERNAL CACHES SO THAT RAW DATA IS SAVED and this Cache works
# Cache(omitArgs = "dat",
# .functionName = "runGPBOOST")  ## models can't be cached. files are being saved and output is the filename

```

### RMSE and AUC
TODO:
```{r}
# RMSEs <- sapply(modelGPB$Fold01$modelGPBcv$best_score, function(mm) {
#   tail(attr(mm$mod,"evaluation_log"), 1)$eval_rmse
# })

## AUC is inappropriate. This is not a classification problem
## yes, we could check accuracy of classifying 0s and 1s, but then we'd
## lose the middle values's accuracies.
# AUCs <- sapply(modelGBT, function(mm) {
#   tail(attr(mm$mod,"evaluation_log"), 1)$eval_auc
# })

# mean(AUCs)
```

### Confidence matrix
```{r}
## predict from each fold and average final confusion matrix
## runXGBOOST outputs predictions and observations and residuals in valData
outConfMats <- modelGPB |>
  purrr::map(\(x) gpboostConfMat(
    mod = x,
    classes = as.character(sort(unique(summaryABSK_AllData[["SEV_CLASS"]]))),
    classMap = unique(summaryABSK_AllData[, .SD, .SDcols = c("SEV_CLASS", "SEV_PROP")])
  )
  )

## make a list, then stack into 3-D array
confMats <- sapply(outConfMats, FUN = function(x) {
  x <- x$confMatrix
  DT <- as.data.table(x$table)
  DT <- dcast.data.table(DT, Prediction ~ Reference)
  cols <- names(DT)
  DT[, (cols) := lapply(.SD, as.numeric), .SDcols = cols]
  as.matrix(DT, rownames = 1)
}, simplify = "array")

outConfMatsStatsGPB <- lapply(outConfMats, function(x) {
  DT <- as.data.table(x$byClass, keep.rownames = TRUE)
  setnames(DT, "rn", "Class")
  DT
}) |>
  rbindlist(idcol = "fold", use.names = TRUE)

```


### Feature importance
```{r}
figDir2 <- file.path(figDir, "GPBoost")
dir.create(figDir2, showWarnings = FALSE)
all_meanSHAP <- purrr::map(modelGPB,
                           function(modFold) {
                             as.data.table(as.list(modFold$shap_values$mean_shap_score))
                           }) |> 
  rbindlist(use.names = TRUE, idcol = "Fold")

meanSHAP <- colMeans(all_meanSHAP[, -"Fold"])
meanSHAP <- as.data.table(melt(meanSHAP), keep.rownames = TRUE)
varOrder <- meanSHAP[order(value), rn]

ggplot(meanSHAP, aes(x = rn, y = value)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = varOrder) +
  theme_pubr() +
  labs(x = "Predictor", y = "SHAP value") +
  coord_flip()
ggsave(file.path(figDir2, "Cross-fold_featImpSHAP.png"),
       width = 7, height = 8.5)

## the full model needs to be fit -- at least I don't see how to do this with 
## the CV outputs
## TODO: cache full model
gp_model <- quote(GPModel(group_data = modelData[["FIRE_NAME"]],
                    likelihood = "zero_inflated_gamma"))
gp_train <- gpb.Dataset(as.matrix(modelData[, ..predvars]),
                        label = modelData[["SURV_PROP"]])

## get parameters from one of the folds (they are always the same)
params <- modelGPB$Fold01$modelGPBfit$modelGPBfit$params
modelGPBFULL <- gpboost(data = gp_train
                        , gp_model = eval(gp_model)
                        , categorical_feature = c("SMR_wet", "isForest", "UNDERSTOREY", "SMR_ord")  ## as above
                        , nrounds = 200
                        , params = params
                        , early_stopping_rounds = 20
) |>
  Cache(omitArgs = c("data", "gp_model"),
        .cacheExtra = c(digestOfData, gp_model),
        .functionName = "gpboost_FULL")

lapply(predvars, function(varname) {
  png(file.path(figDir2, paste0("PDplot_", varname, ".png")))
  gpb.plot.partial.dependence(modelGPBFULL
                              , data = as.matrix(modelData[, ..predvars])
                              , variable = varname
                              , type = "l"
                              , ylab = "survival")
  graphics.off()
  
})


## significant interactions from GAMLSS
png(file.path(figDir2, "PDplot_Decid_UNDERSTOREY.png"))
gpb.plot.part.dep.interact(modelGPBFULL
                           , data = as.matrix(modelData[, ..predvars])
                           , variables = c("Decid", "UNDERSTOREY")
                           , zlab = "survival")
graphics.off()

png(file.path(figDir2, "PDplot_Decid_HEIGHT.png"))
gpb.plot.part.dep.interact(modelGPBFULL
                           , data = as.matrix(modelData[, ..predvars])
                           , variables = c("Decid", "HEIGHT_LOWER")
                           , zlab = "survival")
graphics.off()

png(file.path(figDir2, "PDplot_isForest_UNDERSTOREY.png"))
gpb.plot.part.dep.interact(modelGPBFULL
                           , data = as.matrix(modelData[, ..predvars])
                           , variables = c("isForest", "UNDERSTOREY")
                           , zlab = "survival")
graphics.off()

png(file.path(figDir2, "PDplot_Lari_HEIGHT.png"))
gpb.plot.part.dep.interact(modelGPBFULL
                           , data = as.matrix(modelData[, ..predvars])
                           , variables = c("Lari", "HEIGHT_LOWER")
                           , zlab = "survival")
graphics.off()

png(file.path(figDir2, "PDplot_Lari_UNDERSTOREY.png"))
gpb.plot.part.dep.interact(modelGPBFULL
                           , data = as.matrix(modelData[, ..predvars])
                           , variables = c("Lari", "UNDERSTOREY")
                           , zlab = "survival")
graphics.off()

```

### Feature selection
```{r}
## TODO: test with interaction constraints
```


### Between-covariate relationships
```{r}
## cvDMC correlations > 0.7
plotData <- melt.data.table(summaryABSK_AllData, id.vars = "cvDMC",
                 measure.vars = c("meanDMC", "minDMC", "minBUI",
                                  "cvBUI", "minFFMC", "meanFWI",
                                  'minFWI', 'cvFWI'))
ggplot(plotData, aes(x = cvDMC, y = value)) +
  geom_point() +
  facet_wrap(~ variable)

plotData <- melt.data.table(summaryABSK_AllData, id.vars = "cvBUI",
                 measure.vars = c("meanDMC", "minDMC", "minBUI",
                                  "cvDMC", "minFFMC", "meanFWI",
                                  'minFWI', 'cvFWI'))
ggplot(plotData, aes(x = cvBUI, y = value)) +
  geom_point() +
  facet_wrap(~ variable)
```


### Maps of predicted severity (back transformed)
```{r}
predictionsDT <- summaryABSK_AllData[, .(SEV_PROP, pixID, FIRE_NAME, Lat, Long)]

predData <- data.table(predSURV_PROP = predict(modelGPBFULL, data = as.matrix(modelData[, .SD, .SDcols = predvars]), 
                                              group_data_pred = modelData[["FIRE_NAME"]])$response_mean,
                       pixID = modelData$pixID)
predictionsDT <- predData[predictionsDT, on = "pixID"]
predictionsDT[, predSEV_PROP := 1-predSURV_PROP]
rm(predData); gc(reset = TRUE)

## make severity rasters
firePoints <- predictionsDT[, .(pixID, SEV_PROP, predSEV_PROP, FIRE_NAME, Lat, Long)]
firePoints <- st_as_sf(firePoints, coords = c("Long", "Lat"),
                       agr = "constant", crs = crsProj) |> 
  vect()

## select 2 fires with high agreement between predictions and observations
## correlation between pred and obs severity by fire.
# fires2mapHighAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
#                                          by = FIRE_NAME][order(V1)] |>
#                              tail(2))$FIRE_NAME |>
#   as.character()
# 
# fires2mapLowAgreement <- (predictionsDT[, cor.test(SEV_PROP, predSEV_PROP, method = "pearson")$estimate,
#                                         by = FIRE_NAME][order(V1)] |>
#                             head(2))$FIRE_NAME |>
#   as.character()
## use same as GAMLSS for comparison
fires2mapHighAgreement <- c("BUFFALO_HEAD", "LIVOCK")
fires2mapLowAgreement <- c("CONTEST #2", "WOODMAN")

sevPlotsHighAgrement <- Map(f = plotFun,
                            fireID = as.vector(outer(fires2mapHighAgreement, 
                                                     c("_obs", "_pred"), paste0)),
                            var = rep(c("SEV_PROP", "predSEV_PROP"), each = 2),
                            varTitle = rep(c("Observed", "Predicted mean sev."),
                                           each = 2),
                            MoreArgs = list(firePoints = firePoints))

sevPlotsLowAgrement <- Map(f = plotFun,
                           fireID = as.vector(outer(fires2mapLowAgreement, 
                                                    c("_obs", "_pred"), paste0)),
                           var = rep(c("SEV_PROP", "predSEV_PROP"), each = 2),
                           varTitle = rep(c("Observed", "Predicted mean sev."),
                                          each = 2),
                           MoreArgs = list(firePoints = firePoints))

png(file.path(figDir2, "severityMaps_highAgreement.png"), height = 8, width = 8,
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsHighAgrement$BUFFALO_HEAD_obs + theme(axis.title.x = element_blank()), 
  sevPlotsHighAgrement$BUFFALO_HEAD_pred + theme(axis.title = element_blank()),
  ## line 2
  sevPlotsHighAgrement$LIVOCK_obs, 
  sevPlotsHighAgrement$LIVOCK_pred + theme(axis.title.y = element_blank()), 
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "b)", ""), 
  ncol = 2, nrow = 2
)
dev.off()

png(file.path(figDir2, "severityMaps_lowAgreement.png"), height = 8, width = 6,   
    units = "in", res = 300)
ggarrange(
  ## line 1
  sevPlotsLowAgrement$`CONTEST #2_obs` + theme(axis.title.x = element_blank()),
  sevPlotsLowAgrement$`CONTEST #2_pred` + theme(axis.title = element_blank()), 
  ## line 2
  sevPlotsLowAgrement$WOODMAN_obs, 
  sevPlotsLowAgrement$WOODMAN_pred + theme(axis.title.y = element_blank()), 
  align = "h",
  legend = "right", common.legend = FALSE, 
  labels = c("a)", "", "b)", ""),
  ncol = 2, nrow = 2
)
dev.off()


```

### Residuals
```{r}
valData <- modelGPB$Fold01$valData
plotResiduals(valData)
```


## Compare with zero-one-inflated beta regression

### SHAP analyses
```{r}
cols <- c("SEV_PROP", "Decid", "HEIGHT_LOWER", "CTI", "Lari", "SMR_wet",
          "buffer_960m", "FIRE_NAME", "UNDERSTOREY", "buffer_1920m", "buffer_30m",
          "FlamConif", "PctSlope", "SurfAspectRatio", "SurfReliefRatio", "isForest", "pixID",
          "SEV_CLASS")

modelDataFULLScaled <- summaryABSK_AllData[, ..cols]
cols2scale <- c("Decid", "HEIGHT_LOWER", "CTI", "Lari", "buffer_960m",
                "buffer_1920m", "buffer_30m", "FlamConif", 
                "PctSlope", "SurfAspectRatio", "SurfReliefRatio")
modelDataFULLScaled[, (cols2scale) := lapply(.SD, scale), .SDcols = cols2scale]

modelNgbSev4FULL4_Scaled <- readRDS("analyses/modelNgbSev4FULL4Scaled.rds")
modelDataFULLScaled <- na.omit(modelDataFULLScaled)

p0 <- mean(modelDataFULLScaled$SEV_PROP)

testIDs <- createFolds(modelDataFULLScaled$SEV_PROP, k = 10)$Fold01
trainIDs <- setdiff(1:nrow(modelDataFULLScaled), testIDs)

## get only the predictor columns
cols <- .modelspecsfunGAMLSS(modelNgbSev4FULL4_Scaled)$labels

## much too slow...
explainOut <- explain(
  model = modelNgbSev4FULL4_Scaled,
  x_explain = modelDataFULLScaled[trainIDs, ..cols],
  x_train = modelDataFULLScaled[testIDs, ..cols],
  approach = "ctree",
  phi0 = p0,
  seed = 1,
  predict_model = .predfunGAMLSS,
  get_model_specs = .modelspecsfunGAMLSS
) |> 
  Cache(omitArgs = c("model", "x_explain", "x_train"),
        .functionName = "explain_gamlss")
```

### Get GAMLSS validation metrics and average confidence matrix
```{r}
allLines <- readLines("analyses/modelNgbSev4FULL4Scaled_crossValidOutputs.txt")

## Get GAMLSS validation metrics
startLine <- grep("[[:alpha:]]", allLines)[1]+1
endLine <- grep("\\*{3,}", allLines)[1]-1
vmLines <- allLines[startLine:endLine]
vmLines <- sub("^\\s*\\d+:\\s*", "", vmLines)
## rm col classes
vmLines <- vmLines[-2]

vmFile <- tempfile()
writeLines(vmLines, vmFile)
gamlssValidMetrics <- fread(vmFile)

## Get GAMLSS conf matrix metrics
cacheLines <- showCache(userTags = c("crossValidFunction"))[tagKey == "function" & tagValue == "crossValidFunction"]
gamlssCVout <- cacheLines[order(createdDate, decreasing = TRUE)][1, cacheId] |>
  loadFromCache(cacheId = _)

outConfMatsStatsGamlss <- lapply(gamlssCVout, function(x) {
  DT <- as.data.table(x$confMatrix$byClass, keep.rownames = TRUE)
  setnames(DT, "rn", "Class")
  DT
}) |>
  rbindlist(idcol = "fold", use.names = TRUE)

## Get GAMLSS avg conf matrix
startLine <- grep("Averaged confusion matrices", allLines)[1]+2
cmLines <- allLines[startLine:length(allLines)]
endLine <- grep("\\*{3,}", cmLines)[1]-1
cmLines <- cmLines[1:endLine]
cmLines <- sub("^\\s*\\d+:\\s*", "", cmLines)
## rm col classes
cmLines <- cmLines[-2]

cmFile <- tempfile()
writeLines(cmLines, cmFile)
gamlssConfMatrix <- fread(cmFile)
```

### RMSE

```{r}
message("XGBoost unrestricted -- RMSE, averaged across folds:\n", round(mean(RMSEs), 2))
message("XGBoost restricted features -- RMSE, averaged across folds:\n", round(mean(RMSEs_modelGBT3), 2))
message("GAMLSS -- RMSE, averaged across folds:\n", round(mean(gamlssValidMetrics$RMSE), 2))
```


### Sensitivity and specificity
We can't compare the confusion matrix values between the two models 
directly, we can compare avg sensitivities/specificities per class

```{r}
avgSSxgboost <- outConfMatsStats[, list(Specificity = mean(Specificity),
                                        Sensitivity = mean(Sensitivity)), 
                                 by = Class]

avgSSxgboost_modelGBT3 <- outConfMatsStats_modelGBT3[, list(Specificity = mean(Specificity),
                                                            Sensitivity = mean(Sensitivity)), 
                                                     by = Class]

avgSSgpboost <- outConfMatsStatsGPB[, list(Specificity = mean(Specificity),
                                           Sensitivity = mean(Sensitivity)), 
                                    by = Class]

avgSSgamlss <- outConfMatsStatsGamlss[, list(Specificity = mean(Specificity),
                                             Sensitivity = mean(Sensitivity)), 
                                      by = Class]

```



